### 2017-02-22 doc/paper.tex
Added paper so far

\documentclass{article}
\usepackage{times}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{natbib}
\usepackage{algorithm}
\usepackage{algorithmic}
\newcommand{\theHalgorithm}{\arabic{algorithm}}
\usepackage{hyperref}

\usepackage{mathtools}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{cleveref}
\usepackage{cancel}

% Break in equations...
% \allowdisplaybreaks

% Commonly used sets
\newcommand{\Z}{\mathbb{Z}}   % Integers
\newcommand{\R}{\mathbb{R}}   % Real numbers
\newcommand{\N}{\mathbb{N}}   % Natural numbers

% Symbols
\newcommand{\es}{\varnothing}  % Empty set
\newcommand{\e}{\varepsilon}   % Epsilon
\newcommand{\sub}{\subseteq}   % Subset
\renewcommand{\d}{\partial}    % Partial
\renewcommand{\th}{\theta}     % Theta
\renewcommand{\O}{\mathcal{O}} % Landau's symbol

% Operators
\renewcommand{\Re}{\operatorname{Re}}
\renewcommand{\Im}{\operatorname{Im}}
\renewcommand{\max}{\operatorname{max}}
\newcommand{\argmax}{\operatorname{argmax}}
\newcommand{\argmin}{\operatorname{argmin}}
\newcommand{\tr}{\operatorname{tr}}
\newcommand{\sign}{\operatorname{sign}}
\newcommand{\rank}{\operatorname{rank}}
\newcommand{\diag}{\operatorname{diag}}
\newcommand{\card}{\#}
\newcommand{\comp}{\circ}
\newcommand{\had}{\circ}
\newcommand{\chol}{\operatorname{chol}}
\newcommand{\ind}{1}
\newcommand{\KL}{\operatorname{D}_{\text{KL}}}

% Special math commands
\renewcommand{\ss}[1]{_\mathit{#1}}       % Subscripts without spacing
\newcommand{\id}[1]{\, \mathrm{d} #1}     % Straight 'd' in integral
\newcommand{\sce}{\text{\sc{e}}}          % Scientific notation
\newcommand{\cond}{\, | \,}               % Conditioning
\renewcommand{\ll}{\left}
\newcommand{\rr}{\right}
\newcommand{\la}{\langle}
\newcommand{\ra}{\rangle}
\newcommand{\phan}[1]{\hphantom{#1\;}}

% Cref
\newtheorem{model}{Model}
\crefname{model}{model}{models}
\Crefname{model}{Model}{Models}
\crefname{equation}{equation}{equations}
\Crefname{equation}{Equation}{Equations}

\usepackage[accepted]{icml2017}  % option: accepted
\icmltitlerunning{Learning Causally-Generated Stationary Time Series}

\begin{document}
\twocolumn[
    \icmltitle{Learning Causally-Generated Stationary Time Series}
    \icmlsetsymbol{equal}{*}
    \begin{icmlauthorlist}
    \icmlauthor{Wessel P.\ Bruinsma}{invenia}
    \icmlauthor{Rich E.\ Turner}{cam}
    \end{icmlauthorlist}
    \icmlaffiliation{cam}{University of Cambridge, Cambridge, United Kingdom}
    \icmlaffiliation{invenia}{Invenia Labs Limited, Cambridge, United Kingdom}
    \icmlcorrespondingauthor{Wessel Bruinsma}{wessel.bruinsma@invenialabs.co.uk}
    \icmlkeywords{ICML, machine learning, probabilistic modelling, approximate inference, Gaussian process, causality}
    \vskip 0.3in
]
\printAffiliationsAndNotice{}
% \printAffiliationsAndNotice{\icmlEqualContribution}
% \footnotetext{hi}

% A table:
% \begin{table}[t]
%     \caption{Classification accuracies for naive Bayes and flexible
%     Bayes on various data sets.}
%     \label{sample-table}
%     \vskip 0.15in
%     \centering
%     \begin{small},
%         \begin{sc}
%             \begin{tabular}{ll}
%             \hline
%             \abovespace\belowspace
%             Name & Filter \\
%             \hline
%             \abovespace
%             a & a \\
%             \belowspace
%             a & a \\
%             \hline
%             \end{tabular}
%         \end{sc}
%     \end{small}
%     \vskip -0.1in
% \end{table}

\begin{abstract}
We investigate the notion of causality in modelling stationary time series. Extending on the work by \citet{Tobar:2015:Learning_Stationary}, we present the Causal Gaussian Process Convolution Model (CGPCM). The CGPCM models the kernel of a Gaussian process nonparametrically and has an inductive bias towards causally-generated time series. We develop an algorithm to perform inference and apply the CGPCM to tasks involving real-world data.
\end{abstract}

\section{Introduction}
Classically, nature is considered to be \textit{causal}, meaning that at any point in time the output of any realisable system can only depend on past values of the input. We should aim to include this inductive bias---\textit{causality}---in modelling dynamical phenomena, since that would exclude any unrealisable system from the model prior.

A widely-used model for a stationary time series $f:\R \to \R$ is the Gaussian process, which constructs a prior over functions by assuming that any finite collection of function values $f(t_1),\ldots,f(t_n)$ is multivariate-Gaussian distributed. An important modelling decision in using a Gaussian process is the choice of covariance   $k_f(t,t')$ between any two function values $f(t)$ and $f(t')$---$k_f$ is also called the \textit{kernel} of $f$. This choice of kernel entirely\footnote{The mean of any function value is without loss of generality assumed to be zero. \cite{Rasmussen:2006:Gaussian_Processes}} determines the nature of $f$. Hence much work has been done in developing flexible kernels \cite{Duvenaud:2014:Automatic_Construction,Wilson:2013:Spectral_Mixture,Tobar:2015:Learning_Stationary,Tobar:2015:Inter-Domain_Inducing}; most notably, \citet{Tobar:2015:Learning_Stationary} present the Gaussian Process Convolution Model (GPCM), which parametrises the kernel with \textit{another} Gaussian process.

In this paper we revisit the GPCM and bias it towards signals generated by causal systems---\textit{causally-generated signals}---whilst retaining the flexibility of modelling the kernel nonparametrically. We then investigate a number of properties, develop an algorithm to perform inference, and finally apply it to tasks involving real-world data.

\section{Modelling Causally-Generated Stationary Time Series}
Consider modelling the stationary time series $f$. Since many dynamical systems in nature can accurately be described by an initial value problem or a linear system, we let $f$ be the solution of a time-invariant linear initial value problem with causal Green's function $h$ and forcing function $x$, or equivalently the system response of a time-invariant linear system with causal impulse response $h$ and excitation $x$. It then holds that
\begin{align} \label{eq:model}
    f(t) = \int^t h(t- \tau)x(\tau)\id{\tau}.
\end{align}
We denote integration from negative infinity or to positive infinity by respectively omitting the lower or upper limit of the integral.
Note that $f(t)$ depends on $x(\tau)$ only for $\tau \le t$; this reveals \cref{eq:model}'s causal nature.

In this paper we consider the case that $x$ is white noise; that is, informally denoted, $x \sim \mathcal{GP}(0,\delta(t-t'))$ where $\delta$ denotes the Dirac delta function. In that case \cref{eq:model} can be interpreted as a Gaussian process with zero mean and kernel
\begin{align}
    &k_{f\cond h}(t,t')\nonumber \\
    &\quad= \int^t h(t- \tau)\int^{t'}h(t'- \tau')
        \mathbb{E}[ x(\tau) x(\tau')]\id{\tau'} \id{\tau} \nonumber \\
    &\quad= \int^{t \land t'} h(t - \tau) h(t' - \tau) \id{\tau} \nonumber \\
    &\quad= \int_0 h(|t - t'| + \tau)h(\tau) \id{\tau} \label{eq:kernel} \\
    &\quad= k_{f\cond h}(t-t') \nonumber
\end{align}
where $t \land t'$ denotes the minimum of $t$ and $t'$. Alternatively, \cref{eq:model} can in that case be interpreted as the continuous-time generalisation of a causal moving-average filter.

Since $h$ consitutes an unknown function, we choose to model it using a Gaussian process; that is, we let $h \sim \mathcal{GP}(0,k_h(t,t'))$. In choosing $k_h(t,t')$, we require that $f$ has finite power, since every real-world signal has so as well. This can be achieved by letting $h$ decay to zero at infinity sufficiently quickly \cite{Tobar:2015:Learning_Stationary}: let $h\cond g = w g$ where $w(t)= \exp(- \alpha t^2)$ and $g \sim \mathcal{GP}(0,k_g(t-t'))$ has finite power; in that case,
\begin{align*}
    \mathbb{V}[f(t)]
    &= \int_{0}\exp(-2 \alpha t^2)\mathbb{V}[g(\tau)]\id{\tau} \\
    &\le k_g(0) \int_0 \exp(- 2 \alpha t^2) \id{\tau} < \infty,
\end{align*}
which could be infinite otherwise.
Equivalently, we let $k_h(t,t')=\exp(- \alpha (t^2 + t^{\prime 2}))k_g(t-t')$. To retain flexibility in the prior on $h$, we let $k_g$ be an exponentiated square. We thus have that $k_h(t,t')=\exp(- \alpha (t^2 + t^{\prime 2}) - \gamma(t-t')^2)$. Note that $\alpha$ determines $k_f$'s length scale and $\gamma$ its flexibility.

Through specifying a prior on $x$ and a prior on $h$ we have specified a prior on $f$. Further including a scale $\sigma_f$ to control $f$'s prior power, we call this prior on $f$ the Causal Gaussian Process Convolution Model (CGPCM). To recapitulate, the CGPCM is formulated as follows:

\begin{model}[CGCPM (First Formulation)] \label{mod:cgpcm}
    \begin{align*}
        x &\sim \mathcal{GP}(0,\delta(t-t')), \\
        h &\sim \mathcal{GP}(0, k_h(t,t')), \\
        f\cond h, x &= \sigma_f \int^t h(t- \tau)x(\tau)\id{\tau}.
    \end{align*}
\end{model}
\begin{model}[CGPCM (Second Formulation)] \label{mod:cgpcm2}
    \begin{align*}
        h &\sim \mathcal{GP}(0, k_h(t,t')), \\
        f \cond h &\sim \mathcal{GP}(0,  \sigma_f^2\int_0 h(|t-t'|+\tau)h(\tau)\id{\tau}).
    \end{align*}
\end{model}

\Cref{mod:cgpcm2} reveals the CGCPM as a Gaussian process in which the kernel, or equivalently the power spectral density (PSD), is modelled nonparametrically. \Cref{fig:interpolation} illustrates the generative process of the CGPCM: First, a filter $h$ is generated. Then, the kernel $k_{f\cond h}$ is constructed. Finally, a sample $f\cond h$ is drawn from $\mathcal{GP}(0,k_{f \cond h}(t-t'))$.

The CPGCM is similar to the latent force model presented by \citet{Alvarez:2009:Latent_Force_Models}: whereas we let $h$ be free-form and let $x$ be white noise, \citet{Alvarez:2009:Latent_Force_Models} specify $h$ deterministically and let $x$ be free-form.

\subsection{Sampling from the CGPCM}
Sampling from the CGPCM is difficult because the integrals in \cref{mod:cgpcm,mod:cgpcm2} depend on the entirety of $h$ and $x$. To resolve this issue, we follow \citet{Tobar:2015:Learning_Stationary} and approximate \cref{eq:kernel} using quadrature \cite{Minka:2000:Quadrature_GP} by conditioning on finitely many values $u=(h(t_{u,1}),\ldots,h(t_{u,n_u}))\sim \mathcal{N}(0,K_u)$ of $h$:
\begin{align*}
    k_{f\cond h}(r)
    &\approx \mathbb{E}[k_{f\cond h}(r)\cond u]
    = \int_{0} \mathbb{E}[h(|r| + \tau) h(\tau)\cond u] \id{\tau} \\
    &= \int_0 k_h(|r| + \tau, \tau) \id{\tau} \\
    &\phan{=}+ \tr M^u \int_{0} k_h(t_u, |r|+ \tau)k_h(\tau, t_u^T) \id{\tau}
\end{align*}
where $M^{u}=K_u^{-1}uu^T K_u^{-1}-K_u$. Unlike the GPCM, the kernel approximation $\mathbb{E}[k_{f\cond h}(r)\cond u]$ of the CGPCM will not be a mixture of Gaussians, but a more complicated expression instead.

\subsection{Roughness of Sample Paths}
We show that the CGCPM models both differentiable and nondifferentiable phenomena, the latter in various levels of roughness. This modelling capability holds false for the GPCM, whose sample paths are always differentiable.

Let $h$ be fixed and decay to zero at infinity. We claim that $f$'s sample paths are almost surely everywhere differentiable if $h(0)=0$, and almost surely nowhere differentiable if $h(0)\neq 0$. To show this, let
\begin{align*}
    g(r) = \int_0 h(r + \tau) h(\tau) \id{\tau}
\end{align*}
so that
\begin{align*}
    k_{f\cond h}(r)
    &= g(|r|) \\
    &= g(0) + g'(0)|r| + \frac{1}{2}g''(0)r^2 + \O(|r|^3).
\end{align*}
Then, according to theorem 2.6 and example 2.3 in section 2.3.1.2 in \cite{Lindgren:2006:Lectures_on_Stationary_Stochastic_Processes}, $f$'s sample paths are almost surely everywhere differentiable if $g'(0)=0$; otherwise, $f$ is not even mean square differentiable, in which case $f$'s sample paths are almost surely nowhere differentiable, according to theorem 5 in \cite{Cambanis:1973:On_Some_Continuity_and_Differentiability}. Finally, it holds that
\begin{align*}
    g'(0)
    &= \int_0 h'(\tau) h(\tau) \id{\tau} \\
    &= \ll[ h^2(\tau)\rr]_0 - \int_0 h(\tau) h'(\tau) \id{\tau},
\end{align*}
which shows that $g'(0) = -\frac{1}{2} h^2(0)$---we assumed that $h$ decays to zero at infinity. Hence the claim is shown.

In the case that $h(0)\neq 0$, we can locally approximate $f$ by a scaled Wiener process; this scale $\sigma^2$ then quantifies the roughness of the sample paths. Specifically, given that the variance of an increment of a Wiener process is equal the increment's length, we have that
\begin{align*}
    \sigma^2
    &= \lim_{\e \downarrow 0} \frac{\mathbb{V}[f(t+\e)-f(t)]}{\e} \\
    &= \lim_{\e \downarrow 0} \frac{2k_{f\cond h}(0) - 2k_{f\cond h}(\e)}{\e}
    = -2g'(0)
    = h^2(0).
\end{align*}

In summary, the CGPCM models differentiable phenoma if $h(0)=0$ and nondifferentiable phenoma if $h(0)\neq 0$. In the latter case $|h(0)|$ quantifies the level of roughness.

\Cref{fig:interpolation} shows the filter $h$, the kernel $k_{f\cond h}$, and a sample $f\cond h \sim \mathcal{GP}(0,k_{f\cond h}(t-t'))$ while the filter is interpolated from one that satisfies $h(0)=0$ to one that satisfies $|h(0)|>0$. Note that the sample appears smooth for $h(0)=0$ and becomes rougher as $|h(0)|$ increases.

\begin{figure*}[t]
    % \vskip 0.2in
    \centering
    \includegraphics[width=\linewidth]{../../GGPCM/output/interpolation2.pdf}
    \caption{Generative process of the CGPCM. Shows the filter $h$, the kernel $k_{f\cond h}$, and a sample $f\cond h \sim \mathcal{GP}(0,k_{f\cond h}(t-t'))$ while the filter is interpolated from one that satisfies $h(0)=0$ to one that satisfies $|h(0)|>0$.}
    \label{fig:interpolation}
    % \vskip -0.2in
\end{figure*}

\section{Inference}
Let $y(t)\cond \sim \mathcal{N}(f(t),\sigma^2)$ for all $t$ be a noisy version of $f$. Then, given some observations $e=(y(t_1),\ldots,y(t_n))$ of $y$, we wish to compute $p(f\cond e)$.

We learn $h$ and $x$ through inducing points $u=(h(t_{u,1}),\ldots,h(t_{u,n_u}))\sim \mathcal{N}(0,K_u)$ and $z=(s(t_{z,1}),\ldots,s(t_{z,n_z}))\sim \mathcal{N}(0,K_z)$ for respectively the processes $h$ and $s=T[x]$ where $T$ is some linear interdomain transformation \cite{Titsias:2009:Variational_Learning,Lazaro-Gredilla:2009:Inter-Domain_Gaussian_Processes_for_Sparse,Alvarez:2010:Efficient_Multioutput_Gaussian_Processes_Through,Tobar:2015:Learning_Stationary}. Concretely, let the \textit{mean-field} (MF) approximation $q(f,u,z)=p(f\cond u, z)q(u)q(z)$ of $p(f,u,z\cond e)$ be such that it is closest to $p(f,u,z\cond e)$ in Kullback-Leibler divergence.
% \begin{align*}
%     (q(u),q(z))&= \underset{(q(u),q(z))}{\argmin}\KL(q(f,u,z)\|p(f,u,z\cond e)).
% \end{align*}

Rearranging
\begin{align*}
    &\log p(e) - \KL(q(f,u,z)\|p(f,u,z\cond e)) \\
    &\quad= \ll\la \log \frac{p(e\cond f)\cancel{p(f\cond u, z)}p(u)p(z)}{\cancel{p(f\cond u, z)}q(u)q(z)} \rr\ra_{q(f,u,z)} \\
    &\quad= \underbrace{\la \log p(e\cond f) \ra_{q(f,u,z)}}_{\text{reconstruction cost}} \\
    &\quad\phan{=}- \underbrace{\ll(\ll\la\log\frac{q(u)}{p(u)}\rr\ra_{q(u)} + \ll\la\log\frac{q(z)}{p(z)}\rr\ra_{q(z)}\rr)}_{\text{divergence from prior}} \\
    &\quad=\mathcal{L}[q(u),q(z)]
\end{align*}
shows that can find $q(u)$ and $q(z)$ by maximising $\mathcal{L}$. Since $\log p(e)\ge\mathcal{L}$, $\mathcal{L}$ is called the \textit{evidence lower bound} (ELBO). Observe that maximisation of $\mathcal{L}$ attempts to explain the data well whilst not diverging too far from the model prior.

To optimise $\mathcal{L}$ with respect to $q(u)$ and $q(z)$ we set their respective variations $\delta \mathcal{L} / \delta q(u)$ and $\delta \mathcal{L} / \delta q(z)$ to zero; then solving for $q(u)$ and $q(z)$ yields that
\begin{align}
    q(u) &\propto p(u) \exp \la \log p(e\cond f) \ra_{p(f\cond u,z)q(z)}, \label{eq:qu} \\
    q(z) &\propto p(z) \exp \la \log p(e\cond f) \ra_{p(f\cond u,z)q(u)}, \label{eq:qz}
\end{align}
which are computed in \cref{app:computation_quqz}.
\Cref{app:moments_f} shows that the mean and variance of $p(f\cond u, z)$ are respectively linear and quadratic in both $u$ and $z$; therefore, since $p(e\cond f)$ is Gaussian, we can find a stationary point of $\mathcal{L}$ in which both $q(u)$ and $q(z)$ are Gaussian. Hence, to find $q(u)$ and $q(z)$, we can initialise $q(u)$ and $q(z)$ to some Gaussian and either iterate \cref{eq:qu,eq:qz} or maximise $\mathcal{L}$ directly using gradient-based optimisation. In the latter case we can include any hyperparameter in the optimisation.

We can, however, significantly speed up the optimisation by solving for either $q(u)$ or $q(z)$ analytically. Substituting the optimal form $q(z)$ back in $\mathcal{L}$ yields that
\begin{align}
    &\mathcal{L}^*[q(u)] \nonumber\\
    &\quad=\max_{q(z)} \mathcal{L}[q(u),q(z)]\nonumber\\
    &\quad= \log \int p(z)\exp \la \log p(e\cond f) \ra_{p(f\cond u, z)q(u)} \id{z} \nonumber\\
    &\quad\phan{=}-\ll\la\log\frac{q(u)}{p(u)}\rr\ra_{q(u)},\label{eq:saturated_elbo}
\end{align}
which is computed in \cref{app:saturated_elbo}. We optimise this saturated lower bound $\mathcal{L}^*$ to yield $q(u)$ and then obtain $q(z)$ through \cref{eq:qz}.

Variational mean-field (MF) approaches to inference, like the one just presented, are often computationally efficient, but they are known to suffer from certain biases \cite{MacKay:2002:Information_Theory_Learning,Turner:2011:Two_Problems_With_Variational_Expectation,Murphy:2012:Probabilistic_Perspective}. We therefore further refine the MF approximation to alleviate these biases.

To this end, let the \textit{structured mean-field} (SMF) approximation $q(f,u,z)=p(f\cond u, z)q(u,z)$ of $p(f,u,z\cond e)$ be such that it that is closest to $p(f,u,z\cond e)$ in Kullback-Leibler divergence. Note that the only assumption underlying the SMF approximation is sufficiency of $u$ and $z$ for respectively $h$ and $s$; hence, the SMF approximation will be close the true posterior in the case that there are sufficiently many $u$ and $z$.

Following an argument similar to the one previously presented, we derive the corresponding ELBO:
\begin{align*}
    &\mathcal{L}[q(u),q(z\cond u)] \\
    &\quad= \la \log p(e\cond f) \ra_{q(f,u,z)}- \ll\la\log\frac{q(u)q(z\cond u)}{p(u)p(z)}\rr\ra_{q(u)q(z\cond u)}.
\end{align*}
Again, setting the variations $\delta \mathcal{L} / \delta q(u)$ and $\delta \mathcal{L} / \delta q(z\cond u)$ to zero and solving for for respectively $q(u)$ and $q(z\cond u)$ yields that
\begin{align}
    q(u) &\propto p(u) \int p(z) \exp\la\log p(e\cond f)\ra_{p(f\cond u, z)}\id{z}, \label{eq:qu-smf} \\
    q(z\cond u) &\propto p(z)\exp\la \log p(e\cond f)\ra_{p(f\cond u, z)}, \label{eq:qz-smf}
\end{align}
which are computed in \cref{app:computation_quz}.
As opposed to \cref{eq:qu,eq:qz}, \cref{eq:qu-smf,eq:qz-smf} are uncoupled, but $q(u)$'s moments are now intractable. We can, however, evaluate $q(u)$. We therefore use elliptical slice sampling (ESS) \cite{Murray:2010:Elliptical_Slice_Sampling} to sample from $q(u)$ and use these samples to approximate $q(f, u, z)$. To help mixing the Markov chain, we initialise the sampler with the MF approximation.

\subsection{Interdomain Transformation}
The interdomain transformation $T$ remains yet unspecified. We let
\begin{align*}
    T[x](t)=\int r(t- \tau)x(\tau) \id{\tau}
\end{align*}
where $r$ is some filter. In choosing $r$, we should make sure that $s=T[x]$ has power at the majority of frequencies present in the signal that we aim to model.

For baseband signals, we follow \citet{Tobar:2015:Learning_Stationary} and let $r(t)=\exp(-\omega t^2)$. For signals that have power at all frequencies, such as nondifferentiable signals, we let $r(t)=\exp(- \omega t^2)H(t)$ where $H$ denotes the Heaviside step function; namely, the step at $t=0$ spreads power throughout the whole spectrum. Note that this choice of $T$ yields a \textit{causal interdomain transformation}.

\section{Experiments}
We used TensorFlow \cite{Abadi:2016:TensorFlow_A_System_for_Large-Scale} to implement the CGPCM.\footnote{The implementation can be found at \url{https://github.com/wesselb/cgpcm}.} The main issue in implementing the CGPCM is that computation of the matrices in \cref{app:moments_f} requires extensive evaluation of the bivariate normal cumulative density function (BNCDF) \cite{Genz:2004:Numerical_Computation_of_Rectangular_Bivariate}. Since the BNCDF is not available in TensorFlow, we used the implementation of the BNCDF from the R package \texttt{pbivnorm}\footnote{See \url{https://github.com/brentonk/pbivnorm}.} to implement the BNCDF and its gradient in TensorFlow. We furthermore implementated an algorithm that symbolically solves the integrals in \cref{app:moments_f} and constructs the corresponding matrices afterwards; namely, manually solving these expressions  otherwise would be tedious and error prone.

\subsection{Ornstein-Uhlenbeck Process}
The Ornstein-Uhlenbeck process is a noisy relaxation process that can be formulated as a Gaussian process with kernel of the form $k(r)=\exp(- \alpha |r|)$ \cite{Rasmussen:2006:Gaussian_Processes}. This corresponds to the particular case of the CGPCM where $h(t)=\sqrt{2 \alpha} \exp(- \alpha |t|)$.

{\color{red} Experiment.}

\subsection{RLC Series Circuit}
Consider the voltage response $f(t)$ of a series circuit consisting of a resistance $R$, capacitance $C$, and inductance $L$ excited by current $x(t)$ with unity PSD due to thermal noise of the resistance. Application of Kirchhoff's voltage law \cite{Irwin:2010:Basic_Engineering_Circuit_Analysis} yields that
\begin{align*}
     L f'(t) + R f(t)+\frac{1}{C}\int^t f(t)\id{t} = x(t).
\end{align*}
Assume that the circuit is critically damped and that $R=2$ and $L=1$. Then $f(t)$ is of the form of \cref{eq:model} where $h(t) = t \exp(-t)$. Therefore, according to \cref{mod:cgpcm2}, it holds that $f(t) \sim \mathcal{GP}(0,k_{f\cond h}(t-t'))$ where
\begin{align*}
    k_{f\cond h}(r) = \frac{1}{4}(1+|r|)\exp(-|r|).
\end{align*}

{\color{red} Experiment.}

\subsection{Fricative Consonant}
{\color{red} Experiment.}

\subsection{Head-Related Transfer Function}
{\color{red} Experiment.}

\section{Discussion}
\begin{itemize}
    \item Discussion from \cite{Tobar:2015:Learning_Stationary,Tobar:2015:Inter-Domain_Inducing}
    \item Improved interdomain transformation
    \item HCM to improve sampling
    \item Bayesian inference over the parameters
    \item Improved model solving
\end{itemize}

\appendix
\section{Moments of $f\cond u, h$}
\label{app:moments_f}
We solve for $\la f(t) \ra_{p(f\cond u,z)}$ and $\la f(t) f(t') \ra_{p(f\cond u,z)}$. First, we have that
\begin{align*}
    &\la f(t) \ra_{p(f\cond u,z)} \\
    &\quad= \int^t \la h(t - \tau)\ra_{p(f\cond u)} \la x(\tau) \ra_{p(x\cond z)} \id{\tau} \\
    &\quad= u^T K_u^{-1} \underbrace{\int^t k_h(t_u,t-\tau) k_{xs}(\tau, t_z^T) \id{ \tau}}_{A^{hx}(t)=A^{(xh)T}(t)} K_z^{-1} z.
\end{align*}
Second, it holds that

\begin{align*}
    &\la f(t) f(t') \ra_{p(f\cond u,z)} \\
    &\quad= \int^t\!\!\!\!\int^{t'} \la h(t- \tau) h(t' - \tau')\ra_{p(h\cond u)} \\
    &\quad\phan{=\int^t\!\!\!\!\int^{t'}} \la x(\tau) x(\tau') \ra_{p(x\cond z)}\id{\tau'}\id{\tau} \\
    &\quad= \int^t\!\!\!\!\int^{t'} ( k_h(t- \tau,t' - \tau') \\
    &\quad\phan{=\int^t\!\!\!\!\int^{t'} (}+ k_h(t- \tau, t_u^T) M^u k_h(t_u, t' - \tau')) \\
    &\quad\phan{=\int^t\!\!\!\!\int^{t'}}( k_x(\tau,\tau') + k_{xs}(\tau, t_z^T) M^z k_{sx}(t_z, \tau')) \id{\tau'}\id{\tau} \\
    &\quad= a(t,t') + \tr M^u A^{h}(t,t') + \tr M^z A^{x}(t,t') \\
    &\quad\phan{=} + \tr M^u A^{hx}(t) M^z A^{xh}(t')
\end{align*}
where $M^u=K_u^{-1}uu^T K_u^{-1}-K_u^{-1}$, $M^z=K_z^{-1}zz^TK_z^{-1}-K_z^{-1}$, and
\begin{align*}
    a(t,t')&=\int^t\!\!\!\!\int^{t'} k_h(t- \tau,t' - \tau') k_x(\tau,\tau') \id{\tau'}\id{\tau} \\
    &=\int^{t \land t'} k_h(t- \tau,t' - \tau)\id{\tau}, \\
    A^{h}(t,t')&=\int^t\!\!\!\!\int^{t'} k_h(t_{u}, t' - \tau') k_x(\tau,\tau') \\
    &\phan{=\int^t\!\!\!\!\int^{t'}}\, k_h(t- \tau, t_{u}^T) \id{\tau'}\id{\tau}\\
    &=\int^{t \land t'} k_h(t_{u}, t' - \tau)k_h(t- \tau, t_{u}^T) \id{\tau}  \\
    A^{x}(t,t')&=\int^t\!\!\!\!\int^{t'} k_{sx}(t_z, \tau') k_h(t- \tau,t'-\tau') \\
    &\phan{=\int^t\!\!\!\!\int^{t'}}\, k_{xs}(\tau, t_z^T) \id{\tau'}\id{\tau}.
\end{align*}
Rearranging, we arrive at $\la f(t) \ra = u^T  A^{hx}(t) z$ and
\begin{align*}
    &\la f(t) f(t') \ra - \la f(t) \ra \la f(t') \ra \\
    &\quad = b(t,t') + u^T B^{h}(t,t') u + z^T B^{x}(t,t') z
\end{align*}
where the expectation is over $p(f\cond K_u^{-1} u, K_z^{-1} z)$ and
\begin{align*}
    b(t,t) &= a(t,t') - \tr K_u^{-1} A^{h}(t,t') - \tr K_z^{-1}A^{x}(t,t') \\
    &\phan{=} + \tr K_u^{-1}A^{hx}(t)K_z^{-1}A^{xh}(t'), \\
    B^{h}(t,t') &= A^{h}(t,t') - A^{hx}(t)K_z^{-1} A^{xh}(t'), \\
    B^{x}(t,t') &= A^{x}(t,t') - A^{xh}(t)K_u^{-1} A^{hx}(t').
\end{align*}
Finally denote $a(t)=a(t,t)$ and do so for $A^h$, $A^x$, $A^{hx}$, $b$, $B^h$, and $B^x$ as well.

\subsection{Causal Interdomain Transformation}
In the case of the causal interdomain transformation, $A^x$ and $A^{hx}$ reduce to
\begin{align*}
   A^{hx}(t) &=\int^{t\land t_z^T} k_h(t_u,t-\tau) k_{xs}(\tau, t_z^T) \id{ \tau}, \\
   A^{x}(t,t')&=\int^{t \land t_z}\!\!\!\!\int^{t'\land t_z^T} k_{sx}(t_z, \tau') k_h(t- \tau,t'-\tau') \\
    &\phan{=\int^{t \land t_z}\!\!\!\!\int^{t'\land t_z^T}}\, k_{xs}(\tau, t_z^T) \id{\tau'}\id{\tau}.
\end{align*}

\section{MF Approximation: Computation of $q(u)$ and $q(z)$}
\label{app:computation_quqz}
Following \cref{app:moments_f}, we have that
\begin{align*}
    &\la \log p(e\cond f) \ra_{p(f\cond K_u^{-1} u, K_z^{-1} z)} \\
    &\quad= -\frac{n}{2}\log 2 \pi \sigma^2 - \frac{1}{2 \sigma^2} \la e^{2}(t) - 2 \sigma_f u^T A^{hx}(t) e(t) z\\
    &\quad\phan{=}  + \sigma_f^2 ( b(t) + u^T B^{h}(t) u+ z^T B^{x}(t) z + (u^T A^{hx}(t) z)^2 ) \ra_t
\end{align*}
where $\la\,\cdot\,\ra_t$ denotes summation with respect to $t$ over $t_1,\ldots,t_n$.
It follows that
\begin{align*}
    &\log p(K_u^{-1} u)  + \la \log p(e\cond f) \ra_{p(f\cond K_z^{-1}z, K_u^{-1}u)q(K_z^{-1}z)} \\
    &\quad= -\frac{1}{2}u^T\ll(K_u + \frac{\sigma_f^2}{\sigma^2} \la B^{h}(t) \rr.\\[-3\jot]
    &\quad\phan{= -\frac{1}{2}u^T}
        \underbrace{
            \phan{\ll(K_u + \frac{\sigma_f^2}{\sigma^2} \la \rr. } \!\!\!\!\!
            \ll. \vphantom{\frac{\sigma_f^2}{\sigma^2} }
                 + A^{hx}(t)  z z^T A^{xh}(t) \ra_{t,q(K_z^{-1} z)}
            \rr)
        }_{\Sigma_u^{-1}}u \\
    &\quad\phan{=}+ u^T \underbrace{\frac{\sigma_f}{\sigma^2}\la e(t) A^{hx}(t) z\ra_{t,q(K_z^{-1}z)}}_{\Sigma_u^{-1} \mu_u} \\
    &\quad\phan{=}  + \ll(\vphantom{\frac{\sigma_f^2}{2 \sigma^2}} {-\frac{n}{2}}\log 2 \pi \sigma^2-\frac{1}{2}\log|2 \pi K_u^{-1}|  \rr. \\
    &\quad\phan{=+}\underbrace{\phan{ \ll(\vphantom{\frac{\sigma_f^2}{2 \sigma^2}}\rr.}- \frac{\la e^2(t)\ra_t}{2 \sigma^2} - \ll.\frac{\sigma_f^2}{2 \sigma^2}\la b(t) - z^T B^{x}(t) z\ra_{t,q(K_z^{-1}z)}\rr)}_{\text{constant independent of $u$}} \\
    &\quad= \log \underbrace{\mathcal{N}(u; \mu_u, \Sigma_u)}_{q(K_u^{-1}u)}+ \frac{1}{2}\log|2 \pi \Sigma_u| \\
    &\quad\phan{=}+ \frac{1}{2}\mu_u^T \Sigma_u^{-1}\mu_u  + \text{constant independent of $u$}
\end{align*}
and $q(z)$ is derived similarly.

\section{MF Approximation: Saturated ELBO}
\label{app:saturated_elbo}
Following \cref{eq:saturated_elbo} and \cref{app:computation_quqz}, we have that
\begin{align*}
    &\mathcal{L}^*[q(K_u^{-1} u)] \\
    &\quad = - \frac{n}{2}\log 2 \pi \sigma^2 + \frac{1}{2}\log|K_z^{-1}||\Sigma_z| + \frac{1}{2}\mu_z^T \Sigma_z^{-1}\mu_z  \\
    &\quad \phan{=} - \frac{\la e^2(t)\ra_t}{2 \sigma^2}- \frac{\sigma_f^2}{2 \sigma^2}\la b(t) +  u^T B^{h}(t) u \ra_{t,q(K_u^{-1}u)} \\
    &\quad \phan{=}  + \ll\la \log \frac{p(K_u^{-1}u)}{q(K_u^{-1}u)} \rr\ra_{q(K_u^{-1}u)}.
\end{align*}

\section{SMF Approximation: Computation of $q(u)$ and $q(z\cond u)$}
\label{app:computation_quz}
The mean and variance of $q(z\cond u)$ are similar to that of $q(z)$ in \cref{app:computation_quqz} with only the difference being that the expectation with respect to $q(u)$ is omitted.

To compute $q(u)$, we again follow \cref{app:computation_quqz} and have that
\begin{align*}
    \Sigma &= K_z + \frac{\sigma_f^2}{\sigma^2} \la B^x(t) + A^{xh}(t) u u^TA^{hx}(t) \ra_t, \\
    \mu &= \frac{\sigma_f}{\sigma^2}\la e(t) A^{xh}(t)u\ra_{t}, \\
    \log q(K_u^{-1} u) &= \log p(K_u^{-1} u)- \frac{n}{2}\log 2 \pi \sigma^2 -\frac{1}{2}\log|K_z||\Sigma|\\
    &\phan{=}+ \frac{1}{2}\mu^T \Sigma^{-1} \mu- \frac{\la e^2(t)\ra_t}{2 \sigma^2} \\
    &\phan{=}  - \frac{\sigma_f^2}{2 \sigma^2} \la b(t) + u^T B^h(t)u \ra_t.
\end{align*}

\bibliography{bibliography}
\bibliographystyle{icml2017}

% \section{Remarks}
% For the GPCM it holds that
% \begin{align*}
%     \hat{k}_{f\cond h}(\omega) = \int f(t)\exp(- i \omega t)\id{t}=|\hat{h}(\omega)|^2,
% \end{align*}
% whereas for the CGCPM it holds that
% \begin{align*}
%     \hat{k}_{f\cond h}(\omega)
%     &= \frac{1}{4}\ll|\ll(\pi \delta(\omega) -i \frac{1}{\pi \omega}\rr)\ast\hat{h}(\omega)\rr|^2 \\
%     &= \frac{1}{4}\ll|\pi h(\omega) - i (\mathcal{H}\hat{h})(\omega)\rr|^2
% \end{align*}
% where $\mathcal{H}$ is the Hilbert transform. Can this be interpreted?

\end{document}

### 2017-02-22 doc/paper.tex
Added missing image
Bit of discussion

@@
     \centering
-    \includegraphics[width=\linewidth]{../../GGPCM/output/interpolation2.pdf}
+    \includegraphics[width=\linewidth]{resources/interpolation.pdf}
     \caption{Generative process of the CGPCM.
@@
 
-\subsection{Fricative Consonant}
+\subsection{Unvoiced Fricatives}
+Unvoiced fricatives are consonants produced by forcing air through a narrow constriction in the vocal tract, thereby creating a noisy, turbulant sound \cite{Taylor:2009:Text-To-Speech_Synthesis}.
+
 {\color{red} Experiment.}
@@
 \subsection{Head-Related Transfer Function}
+A head-related transfer function (HRTF) is a transfer function that describes how an ear receives sound from a certain point in space.
+
 {\color{red} Experiment.}
@@
 \section{Discussion}
-\begin{itemize}
-    \item Discussion from \cite{Tobar:2015:Learning_Stationary,Tobar:2015:Inter-Domain_Inducing}
-    \item Improved interdomain transformation
-    \item HCM to improve sampling
-    \item Bayesian inference over the parameters
-    \item Improved model solving
-\end{itemize}
+We have presented the CGPCM as a model for causally-generated stationary time series.
+Inference is performed by finding an initial approximation through a variational mean-field approach, and then refining this approximation through a variational structured mean-field approach combined with elliptical slice sampling.
+The GPCM has been validated {\color{red} discussion of results of experiments here.}
+
+In future research on the CGPCM, various directions can be explored: First, the model could be extented to multidimensional input and output spaces by following a construction similar to that by \citet{Bruinsma:2016:GGPCM}.
+Second, the application of more efficient sparse approximation for $h$ and $x$, such as the tree-structured approximation \cite{Bui:2014:Tree-Structured_Gaussian}, would scale the CGPCM to larger data sets.
+Third, to deal with signals that are band-limited, but not necessarily baseband, the use of an harmonic interdomain transformation \cite{Tobar:2015:Inter-Domain_Inducing} and an harmonic extension of the causal interdomain transformation could be explored.
+Fourth, in the low noise case, ESS has a hard time exploring $q(u)$; therefore, in that case, afternative sampling algorithms such as Hamiltonian Monte-Carlo \cite{Neal:2012:MCMC_Using_Hamiltonian_Dynamics} could be used.
+Finally, using an approach similar to that by \citet{Lazaro-Gredilla:2013:Variational_Inference_for_Mahalanobis_Distance}, Bayesian inference in the hyperparameters could be attempted.
 

### 2017-02-23 doc/paper.tex
Various updates

@@
 An important modelling decision in using a Gaussian process is the choice of covariance   $k_f(t,t')$ between any two function values $f(t)$ and $f(t')$---$k_f$ is also called the \textit{kernel} of $f$.
-This choice of kernel entirely\footnote{The mean of any function value is without loss of generality assumed to be zero.
+This choice of kernel entirely\footnote{The mean of the function is assumed to be zero without loss of generality.
 \cite{Rasmussen:2006:Gaussian_Processes}} determines the nature of $f$.
@@
 \end{align}
-We denote integration from negative infinity or to positive infinity by respectively omitting the lower or upper limit of the integral.
+We denote integration from negative infinity or to positive infinity by omitting the respective limit from the integral.
 Note that $f(t)$ depends on $x(\tau)$ only for $\tau \le t$; this reveals \cref{eq:model}'s causal nature.
@@
 In choosing $k_h(t,t')$, we require that $f$ has finite power, since every real-world signal has so as well.
-This can be achieved by letting $h$ decay to zero at infinity sufficiently quickly \cite{Tobar:2015:Learning_Stationary}: let $h\cond g = w g$ where $w(t)= \exp(- \alpha t^2)$ and $g \sim \mathcal{GP}(0,k_g(t-t'))$ has finite power; in that case,
+Following \citet{Tobar:2015:Learning_Stationary}, this can be achieved by letting $h$ decay to zero at infinity sufficiently quickly: let $h\cond g = w g$ where $w(t)= \exp(- \alpha t^2)$ and $g \sim \mathcal{GP}(0,k_g(t-t'))$ has finite power; in that case,
 \begin{align*}
@@
 
-The CPGCM is similar to the latent force model presented by \citet{Alvarez:2009:Latent_Force_Models}: whereas we let $h$ be free-form and let $x$ be white noise, \citet{Alvarez:2009:Latent_Force_Models} specify $h$ deterministically and let $x$ be free-form.
+The CPGCM is similar to the latent force model presented by \citet{Alvarez:2009:Latent_Force_Models}; however, whereas we let $h$ be free-form and let $x$ be white noise, \citet{Alvarez:2009:Latent_Force_Models} specify $h$ deterministically and let $x$ be free-form.
 
@@
 where $M^{u}=K_u^{-1}uu^T K_u^{-1}-K_u$.
-Unlike the GPCM, the kernel approximation $\mathbb{E}[k_{f\cond h}(r)\cond u]$ of the CGPCM will not be a mixture of Gaussians, but a more complicated expression instead.
+Unlike the GPCM, the kernel approximation $\mathbb{E}[k_{f\cond h}(r)\cond u]$ of the CGPCM will not simply be a mixture of Gaussians; it will be a more complicated expression instead.
 
@@
 \section{Inference}
-Let $y(t)\cond \sim \mathcal{N}(f(t),\sigma^2)$ for all $t$ be a noisy version of $f$. Then, given some observations $e=(y(t_1),\ldots,y(t_n))$ of $y$, we wish to compute $p(f\cond e)$.
+Let $y(t)\cond \sim \mathcal{N}(f(t),\sigma^2)$ for all $t$ be a noisy version of $f$. Given some observations $e=(y(t_1),\ldots,y(t_n))$ of $y$, we wish to compute $p(f,h,x\cond e)$.
 
-We learn $h$ and $x$ through inducing points $u=(h(t_{u,1}),\ldots,h(t_{u,n_u}))\sim \mathcal{N}(0,K_u)$ and $z=(s(t_{z,1}),\ldots,s(t_{z,n_z}))\sim \mathcal{N}(0,K_z)$ for respectively the processes $h$ and $s=T[x]$ where $T$ is some linear interdomain transformation \cite{Titsias:2009:Variational_Learning,Lazaro-Gredilla:2009:Inter-Domain_Gaussian_Processes_for_Sparse,Alvarez:2010:Efficient_Multioutput_Gaussian_Processes_Through,Tobar:2015:Learning_Stationary}.
-Concretely, let the \textit{mean-field} (MF) approximation $q(f,u,z)=p(f\cond u, z)q(u)q(z)$ of $p(f,u,z\cond e)$ be such that it is closest to $p(f,u,z\cond e)$ in Kullback-Leibler divergence.
-% \begin{align*}
-%     (q(u),q(z))&= \underset{(q(u),q(z))}{\argmin}\KL(q(f,u,z)\|p(f,u,z\cond e)).
-% \end{align*}
+We learn $h$ and $x$ through inducing points \cite{Titsias:2009:Variational_Learning} $u=(h(t_{u,1}),\ldots,h(t_{u,n_u}))\sim \mathcal{N}(0,K_u)$ and $z=(s(t_{z,1}),\ldots,s(t_{z,n_z}))\sim \mathcal{N}(0,K_z)$ for respectively the processes $h$ and $s=T[x]$ where
+\begin{align*}
+    T[x](t)=\int r(t- \tau)x(\tau) \id{\tau}
+\end{align*}
+is some interdomain transformation with filter $r$ \cite{Lazaro-Gredilla:2009:Inter-Domain_Gaussian_Processes_for_Sparse,Alvarez:2010:Efficient_Multioutput_Gaussian_Processes_Through,Tobar:2015:Learning_Stationary}.
+In choosing $r$, we make sure that $s=T[x]$ has power at the majority of frequencies present in the signal that we aim to model.
+We follow \citet{Tobar:2015:Learning_Stationary} and let $r(t)=\exp(-\omega t^2)$, which fits baseband signals.
 
-Rearranging
+Let the \textit{mean-field} (MF) approximation $q(f,u,z)=p(f\cond u, z)q(u)q(z)$ of $p(f,u,z\cond e)$ be such that it is closest to $p(f,u,z\cond e)$ in Kullback-Leibler divergence. Rearranging
 \begin{align*}
@@
 \end{align*}
-shows that can find $q(u)$ and $q(z)$ by maximising $\mathcal{L}$.
+shows that we can find $q(u)$ and $q(z)$ through maximising $\mathcal{L}$.
 Since $\log p(e)\ge\mathcal{L}$, $\mathcal{L}$ is called the \textit{evidence lower bound} (ELBO).
@@
 
-We can, however, significantly speed up the optimisation by solving for either $q(u)$ or $q(z)$ analytically. Substituting the optimal form $q(z)$ back in $\mathcal{L}$ yields that
+We can, however, significantly speed up the optimisation by solving for either $q(u)$ or $q(z)$ analytically. Substituting the optimal form of $q(z)$ back in $\mathcal{L}$ yields that
 \begin{align}
@@
 
-Following an argument similar to the one previously presented, we derive the corresponding ELBO:
+Following an argument similar, we again derive the corresponding ELBO:
 \begin{align*}
@@
 
-\subsection{Interdomain Transformation}
-The interdomain transformation $T$ remains yet unspecified. We let
-\begin{align*}
-    T[x](t)=\int r(t- \tau)x(\tau) \id{\tau}
-\end{align*}
-where $r$ is some filter. In choosing $r$, we should make sure that $s=T[x]$ has power at the majority of frequencies present in the signal that we aim to model.
-
-For baseband signals, we follow \citet{Tobar:2015:Learning_Stationary} and let $r(t)=\exp(-\omega t^2)$.
-For signals that have power at all frequencies, such as nondifferentiable signals, we let $r(t)=\exp(- \omega t^2)H(t)$ where $H$ denotes the Heaviside step function; namely, the step at $t=0$ spreads power throughout the whole spectrum.
-Note that this choice of $T$ yields a \textit{causal interdomain transformation}.
-
 \section{Experiments}
-We used TensorFlow \cite{Abadi:2016:TensorFlow_A_System_for_Large-Scale} to implement the CGPCM.\footnote{The implementation can be found at \url{https://github.com/wesselb/cgpcm}.} The main issue in implementing the CGPCM is that computation of the matrices in \cref{app:moments_f} requires extensive evaluation of the bivariate normal cumulative density function (BNCDF) \cite{Genz:2004:Numerical_Computation_of_Rectangular_Bivariate}.
+We used TensorFlow \cite{Abadi:2016:TensorFlow_A_System_for_Large-Scale} to implement the CGPCM.\footnote{The implementation can be found at \url{https://github.com/wesselb/cgpcm}.} The main issue in implementing the CGPCM is that computation of the matrices in \cref{app:moments_f} requires the bivariate normal cumulative density function (BNCDF) \cite{Genz:2004:Numerical_Computation_of_Rectangular_Bivariate}.
 Since the BNCDF is not available in TensorFlow, we used the implementation of the BNCDF from the R package \texttt{pbivnorm}\footnote{See \url{https://github.com/brentonk/pbivnorm}.} to implement the BNCDF and its gradient in TensorFlow.
@@
 
-\subsection{Ornstein-Uhlenbeck Process}
-The Ornstein-Uhlenbeck process is a noisy relaxation process that can be formulated as a Gaussian process with kernel of the form $k(r)=\exp(- \alpha |r|)$ \cite{Rasmussen:2006:Gaussian_Processes}.
-This corresponds to the particular case of the CGPCM where $h(t)=\sqrt{2 \alpha} \exp(- \alpha |t|)$.
+\subsection{Learning Known Kernels}
 
 {\color{red} Experiment.}
+
+% \subsection{Ornstein-Uhlenbeck Process}
+% The Ornstein-Uhlenbeck process is a noisy relaxation process that can be formulated as a Gaussian process with kernel of the form $k(r)=\exp(- \alpha |r|)$ \cite{Rasmussen:2006:Gaussian_Processes}.
+This corresponds to the particular case of the CGPCM where $h(t)=\sqrt{2 \alpha} \exp(- \alpha |t|)$.
 
@@
 We have presented the CGPCM as a model for causally-generated stationary time series.
-Inference is performed by finding an initial approximation through a variational mean-field approach, and then refining this approximation through a variational structured mean-field approach combined with elliptical slice sampling.
+Inference is performed by finding an initial approximation through a variational MF approach, and then refining this approximation through a variational SMF approach combined with ESS.
 The GPCM has been validated {\color{red} discussion of results of experiments here.}
@@
 Second, the application of more efficient sparse approximation for $h$ and $x$, such as the tree-structured approximation \cite{Bui:2014:Tree-Structured_Gaussian}, would scale the CGPCM to larger data sets.
-Third, to deal with signals that are band-limited, but not necessarily baseband, the use of an harmonic interdomain transformation \cite{Tobar:2015:Inter-Domain_Inducing} and an harmonic extension of the causal interdomain transformation could be explored.
+Third, to deal with signals that are band-limited, but not necessarily baseband, the use of an harmonic interdomain transformation \cite{Tobar:2015:Inter-Domain_Inducing} could be explored.
 Fourth, in the low noise case, ESS has a hard time exploring $q(u)$; therefore, in that case, afternative sampling algorithms such as Hamiltonian Monte-Carlo \cite{Neal:2012:MCMC_Using_Hamiltonian_Dynamics} could be used.
@@
 \end{align*}
-Finally denote $a(t)=a(t,t)$ and do so for $A^h$, $A^x$, $A^{hx}$, $b$, $B^h$, and $B^x$ as well.
+Finally, we denote $a(t)=a(t,t)$ and do so for $A^h$, $A^x$, $A^{hx}$, $b$, $B^h$, and $B^x$ as well.
 
-\subsection{Causal Interdomain Transformation}
-In the case of the causal interdomain transformation, $A^x$ and $A^{hx}$ reduce to
-\begin{align*}
-   A^{hx}(t) &=\int^{t\land t_z^T} k_h(t_u,t-\tau) k_{xs}(\tau, t_z^T) \id{ \tau}, \\
-   A^{x}(t,t')&=\int^{t \land t_z}\!\!\!\!\int^{t'\land t_z^T} k_{sx}(t_z, \tau') k_h(t- \tau,t'-\tau') \\
-    &\phan{=\int^{t \land t_z}\!\!\!\!\int^{t'\land t_z^T}}\, k_{xs}(\tau, t_z^T) \id{\tau'}\id{\tau}.
-\end{align*}
+% \subsection{Causal Interdomain Transformation}
+% In the case of the causal interdomain transformation, $A^{hx}$ and $A^x$ reduce to
+% \begin{align*}
+%    A^{hx}(t) &=\int^{t\land t_z^T} k_h(t_u,t-\tau) k_{xs}(\tau, t_z^T) \id{ \tau}, \\
+%    A^{x}(t,t')&=\int^{t \land t_z}\!\!\!\!\int^{t'\land t_z^T} k_{sx}(t_z, \tau') k_h(t- \tau,t'-\tau') \\
+%     &\phan{=\int^{t \land t_z}\!\!\!\!\int^{t'\land t_z^T}}\, k_{xs}(\tau, t_z^T) \id{\tau'}\id{\tau}.
+% \end{align*}
 

### 2017-02-23 doc/paper.tex
Some updates

@@
 
-\usepackage[accepted]{icml2017}  % option: accepted
+\usepackage{icml2017}  % option: accepted
 \icmltitlerunning{Learning Causally-Generated Stationary Time Series}
@@
 
-Let the \textit{mean-field} (MF) approximation $q(f,u,z)=p(f\cond u, z)q(u)q(z)$ of $p(f,u,z\cond e)$ be such that it is closest to $p(f,u,z\cond e)$ in Kullback-Leibler divergence. Rearranging
+Specifically, let the \textit{mean-field} (MF) approximation $q(f,u,z)=p(f\cond u, z)q(u)q(z)$ of $p(f,u,z\cond e)$ be such that it is closest to $p(f,u,z\cond e)$ in Kullback-Leibler divergence.
+Rearranging
 \begin{align*}
@@
 Variational mean-field (MF) approaches to inference, like the one just presented, are often computationally efficient, but they are known to suffer from certain biases \cite{MacKay:2002:Information_Theory_Learning,Turner:2011:Two_Problems_With_Variational_Expectation,Murphy:2012:Probabilistic_Perspective}.
-We therefore further refine the MF approximation to alleviate these biases.
+We further refine the MF approximation to alleviate these biases.
 
@@
 
-\subsection{RLC Series Circuit}
-Consider the voltage response $f(t)$ of a series circuit consisting of a resistance $R$, capacitance $C$, and inductance $L$ excited by current $x(t)$ with unity PSD due to thermal noise of the resistance.
+% \subsection{RLC Series Circuit}
+% Consider the voltage response $f(t)$ of a series circuit consisting of a resistance $R$, capacitance $C$, and inductance $L$ excited by current $x(t)$ with unity PSD due to thermal noise of the resistance.
 Application of Kirchhoff's voltage law \cite{Irwin:2010:Basic_Engineering_Circuit_Analysis} yields that
-\begin{align*}
-     L f'(t) + R f(t)+\frac{1}{C}\int^t f(t)\id{t} = x(t).
-\end{align*}
-Assume that the circuit is critically damped and that $R=2$ and $L=1$.
+% \begin{align*}
+%      L f'(t) + R f(t)+\frac{1}{C}\int^t f(t)\id{t} = x(t).
+% \end{align*}
+% Assume that the circuit is critically damped and that $R=2$ and $L=1$.
 Then $f(t)$ is of the form of \cref{eq:model} where $h(t) = t \exp(-t)$.
 Therefore, according to \cref{mod:cgpcm2}, it holds that $f(t) \sim \mathcal{GP}(0,k_{f\cond h}(t-t'))$ where
-\begin{align*}
-    k_{f\cond h}(r) = \frac{1}{4}(1+|r|)\exp(-|r|).
-\end{align*}
+% \begin{align*}
+%     k_{f\cond h}(r) = \frac{1}{4}(1+|r|)\exp(-|r|).
+% \end{align*}
 
-{\color{red} Experiment.}
+% {\color{red} Experiment.}
 
-\subsection{Unvoiced Fricatives}
-Unvoiced fricatives are consonants produced by forcing air through a narrow constriction in the vocal tract, thereby creating a noisy, turbulant sound \cite{Taylor:2009:Text-To-Speech_Synthesis}.
-
-{\color{red} Experiment.}
+% \subsection{Unvoiced Fricatives}
+% Unvoiced fricatives are consonants produced by forcing air through a narrow constriction in the vocal tract, thereby creating a noisy, turbulant sound \cite{Taylor:2009:Text-To-Speech_Synthesis}.
 
@@
 
+\subsection{Speech}
+
 {\color{red} Experiment.}
+
+{\color{red} Experiment.}
+
+\subsection{Population Growth}
 
@@
 Third, to deal with signals that are band-limited, but not necessarily baseband, the use of an harmonic interdomain transformation \cite{Tobar:2015:Inter-Domain_Inducing} could be explored.
+%???
 Fourth, in the low noise case, ESS has a hard time exploring $q(u)$; therefore, in that case, afternative sampling algorithms such as Hamiltonian Monte-Carlo \cite{Neal:2012:MCMC_Using_Hamiltonian_Dynamics} could be used.

### 2017-02-23 doc/paper.tex
Minor edit

@@
     \icmlauthor{Wessel P.\ Bruinsma}{invenia}
-    \icmlauthor{Rich E.\ Turner}{cam}
+    \icmlauthor{Richard E.\ Turner}{cam}
     \end{icmlauthorlist}
@@
 \section{Introduction}
-Classically, nature is considered to be \textit{causal}, meaning that at any point in time the output of any realisable system can only depend on past values of the input.
+Many natural phenomena are \textit{causal} meaning that at any point in time the output of any realisable system can only depend on past values of the input.
 We should aim to include this inductive bias---\textit{causality}---in modelling dynamical phenomena, since that would exclude any unrealisable system from the model prior.

### 2017-02-23 doc/paper.tex
Made a pass through the intro.

@@
 \section{Introduction}
-Many natural phenomena are \textit{causal} meaning that at any point in time the output of any realisable system can only depend on past values of the input.
-We should aim to include this inductive bias---\textit{causality}---in modelling dynamical phenomena, since that would exclude any unrealisable system from the model prior.
 
-A widely-used model for a stationary time series $f:\R \to \R$ is the Gaussian process, which constructs a prior over functions by assuming that any finite collection of function values $f(t_1),\ldots,f(t_n)$ is multivariate-Gaussian distributed.
-An important modelling decision in using a Gaussian process is the choice of covariance   $k_f(t,t')$ between any two function values $f(t)$ and $f(t')$---$k_f$ is also called the \textit{kernel} of $f$.
-This choice of kernel entirely\footnote{The mean of the function is assumed to be zero without loss of generality.
-\cite{Rasmussen:2006:Gaussian_Processes}} determines the nature of $f$.
-Hence much work has been done in developing flexible kernels \cite{Duvenaud:2014:Automatic_Construction,Wilson:2013:Spectral_Mixture,Tobar:2015:Learning_Stationary,Tobar:2015:Inter-Domain_Inducing}; most notably, \citet{Tobar:2015:Learning_Stationary} present the Gaussian Process Convolution Model (GPCM), which parametrises the kernel with \textit{another} Gaussian process.
+One of the major goals of statistical inference is to develop models of dynamical phenomena that can be used for prediction, extrapolation, and system identification.
+Two key characteristics of the physical systems underlying natural and manmade dynamical phenomena are that they are \textit{causal} and \textit{spectrally complex}.
+Causal systems, where at any point in time the output of the system can only depend on past values of the input, are the only type which are physically realisable.
+Spectrally complex systems, which have rich power-spectral densities, arise because many physical systems have numerous resonances spanning many time scales.
+When developing statistical models for such phenomena this prior knowledge---causality and spectral richness---should be leveraged in order to exclude any unrealisable system from the model prior whilst at the same time allowing the model to have the capacity to capture varied spectral content that may be only slowly revealed as more data are seen.
+The goal of this paper is to develop Gaussian Process models together with associated inference and learning schemes that serve this purpose.
 
-In this paper we revisit the GPCM and bias it towards signals generated by causal systems---\textit{causally-generated signals}---whilst retaining the flexibility of modelling the kernel nonparametrically.
-We then investigate a number of properties, develop an algorithm to perform inference, and finally apply it to tasks involving real-world data.
+Gaussian Processes are a widely-used model for a stationary time series that place a prior distribution over the latent function underlying the time series $f:\R \to \R$ by assuming that any finite collection of function values $f(t_1),\ldots,f(t_n)$ is multivariate-Gaussian distributed.
+The key modelling decision in using a Gaussian process is the choice of covariance  $k_f(t,t')$ between any two function values $f(t)$ and $f(t')$---$k_f$ is also called the \textit{kernel} of $f$.
+This choice of kernel\footnote{The mean of the function is assumed to be zero without loss of generality.
+\cite{Rasmussen:2006:Gaussian_Processes}} encodes prior information about the function $f$.
+Consequently, a large research effort has been devoted toward developing flexible kernels \cite{Duvenaud:2014:Automatic_Construction,Wilson:2013:Spectral_Mixture,Tobar:2015:Learning_Stationary,Tobar:2015:Inter-Domain_Inducing}.
+
+Of particular interest for the current application are GP models that use non-parametric models of the power-spectra of the Gaussian Process (doubly non-parametric models) as they have the capacity to flexibly model signals with arbitrary spectral complexity.
+One approach employs a Dirichlet Process Mixture Model for the power spectra \cite{Bayesian nonparametric kernel learning, Wilson}, but it is not clear whether it is possible to build causality into such a construction.
+An alternative approach induces a non-parametric power spectrum by placing a Gaussian Process prior over the system's impulse response function  \citet{Tobar:2015:Learning_Stationary}.
+Critically, the so-called Gaussian Process Convolution Model (GPCM), does not incorporate a causality constraint (the impulse response in question should be thought of as an acausal spatial response).
+In this paper we revisit the GPCM and explicitly build in a causal constraint whilst retaining the flexibility of modelling the kernel nonparametrically.
+In addition to these modelling contributions, we develop enhanced variational inference and learning schemes for the causal and acausal variants of the GPCM -- collapsed variational bounds and structured approximating distributions -- that significantly improve statistical accuracy.
+These modelling and inferential contributions are demonstrated on synthetic and real world signals.
 

### 2017-02-23 doc/paper.tex
Minor tweak.

@@
 Of particular interest for the current application are GP models that use non-parametric models of the power-spectra of the Gaussian Process (doubly non-parametric models) as they have the capacity to flexibly model signals with arbitrary spectral complexity.
-One approach employs a Dirichlet Process Mixture Model for the power spectra \cite{Bayesian nonparametric kernel learning, Wilson}, but it is not clear whether it is possible to build causality into such a construction.
-An alternative approach induces a non-parametric power spectrum by placing a Gaussian Process prior over the system's impulse response function  \citet{Tobar:2015:Learning_Stationary}.
+One approach employs a Dirichlet Process Mixture Model for the power spectra \cite{Bayesian nonparametric kernel learning, Wilson}, but it is not clear whether it is possible to build the causality constraint into such a construction.
+An alternative approach instead induces a non-parametric power spectrum by placing a Gaussian Process prior over the system's impulse response function  \citet{Tobar:2015:Learning_Stationary}.
 Critically, the so-called Gaussian Process Convolution Model (GPCM), does not incorporate a causality constraint (the impulse response in question should be thought of as an acausal spatial response).

### 2017-02-23 doc/paper.tex
Minor edits, typo's, consistency in style

--- 2017-02-23 doc/paper.tex
@@
 Causal systems, where at any point in time the output of the system can only depend on past values of the input, are the only type which are physically realisable.
-Spectrally complex systems, which have rich power-spectral densities, arise because many physical systems have numerous resonances spanning many time scales.
+Spectrally complex systems, which have rich power spectral densities, arise because many physical systems have numerous resonances spanning many time scales.
 When developing statistical models for such phenomena this prior knowledge---causality and spectral richness---should be leveraged in order to exclude any unrealisable system from the model prior whilst at the same time allowing the model to have the capacity to capture varied spectral content that may be only slowly revealed as more data are seen.
-The goal of this paper is to develop Gaussian Process models together with associated inference and learning schemes that serve this purpose.
+The goal of this paper is to develop Gaussian Process (GP) models together with associated inference and learning schemes that serve this purpose.
 
-Gaussian Processes are a widely-used model for a stationary time series that place a prior distribution over the latent function underlying the time series $f:\R \to \R$ by assuming that any finite collection of function values $f(t_1),\ldots,f(t_n)$ is multivariate-Gaussian distributed.
-The key modelling decision in using a Gaussian process is the choice of covariance  $k_f(t,t')$ between any two function values $f(t)$ and $f(t')$---$k_f$ is also called the \textit{kernel} of $f$.
+A widely-used model for stationary time series are Gaussian Processes, which place a prior distribution over the latent function underlying the time series $f:\R \to \R$ by assuming that any finite collection of function values $f(t_1),\ldots,f(t_n)$ is multivariate-Gaussian distributed.
+The key modelling decision in using a Gaussian Process is the choice of covariance $k_f(t,t')$ between any two function values $f(t)$ and $f(t')$---$k_f$ is also called the \textit{kernel} of $f$.
 This choice of kernel\footnote{The mean of the function is assumed to be zero without loss of generality.
 \cite{Rasmussen:2006:Gaussian_Processes}} encodes prior information about the function $f$.
-Consequently, a large research effort has been devoted toward developing flexible kernels \cite{Duvenaud:2014:Automatic_Construction,Wilson:2013:Spectral_Mixture,Tobar:2015:Learning_Stationary,Tobar:2015:Inter-Domain_Inducing}.
+Consequently, a large research effort has been devoted towards developing flexible kernels \cite{Duvenaud:2014:Automatic_Construction,Wilson:2013:Spectral_Mixture,Tobar:2015:Learning_Stationary,Tobar:2015:Inter-Domain_Inducing}.
 
-Of particular interest for the current application are GP models that use non-parametric models of the power-spectra of the Gaussian Process (doubly non-parametric models) as they have the capacity to flexibly model signals with arbitrary spectral complexity.
-One approach employs a Dirichlet Process Mixture Model for the power spectra \cite{Bayesian nonparametric kernel learning, Wilson}, but it is not clear whether it is possible to build causality into such a construction.
-An alternative approach induces a non-parametric power spectrum by placing a Gaussian Process prior over the system's impulse response function  \citet{Tobar:2015:Learning_Stationary}.
-Critically, the so-called Gaussian Process Convolution Model (GPCM), does not incorporate a causality constraint (the impulse response in question should be thought of as an acausal spatial response).
-In this paper we revisit the GPCM and explicitly build in a causal constraint whilst retaining the flexibility of modelling the kernel nonparametrically.
-In addition to these modelling contributions, we develop enhanced variational inference and learning schemes for the causal and acausal variants of the GPCM -- collapsed variational bounds and structured approximating distributions -- that significantly improve statistical accuracy.
-These modelling and inferential contributions are demonstrated on synthetic and real world signals.
+Of particular interest for the current application are GP models that use nonparametric models of the power spectra of the GPs---doubly non-parametric models---as they have the capacity to flexibly model signals with arbitrary spectral complexity.
+One approach employs a Dirichlet Process Mixture Model for the power spectra \cite{Oliva:2015:Bayesian_Nonparametric_Kernel-Learning}, but it is not clear whether it is possible to build the causality constraint into such a construction.
+An alternative approach instead induces a nonparametric power spectrum by placing a Gaussian Process prior over the system's impulse response function \cite{Tobar:2015:Learning_Stationary}.
+Critically, the so-called Gaussian Process Convolution Model (GPCM), does not incorporate a causality constraint---the impulse response in question should be thought of as an acausal spatial response.
+In this paper we revisit the GPCM and explicitly build in a causality constraint whilst retaining the flexibility of modelling the kernel nonparametrically.
+In addition to these modelling contributions, we develop enhanced variational inference and learning schemes for the causal and acausal variants of the GPCM---collapsed variational bounds and structured approximating distributions---that significantly improve statistical accuracy.
+These modelling and inferential contributions are demonstrated on synthetic and real-world signals.
 

### 2017-02-24 doc/paper.tex
Figures of toy experiment; better late than never

@@
 \usepackage{cancel}
+
+\usepackage{color}
+\definecolor{darkgreen}{rgb}{0,.5,0}
+\definecolor{mycyan}{rgb}{0,.9,.9}
+
+\usepackage{tabularx}
+
+\usepackage{tikz}
 
@@
 
-% \subsection{Ornstein-Uhlenbeck Process}
-% The Ornstein-Uhlenbeck process is a noisy relaxation process that can be formulated as a Gaussian process with kernel of the form $k(r)=\exp(- \alpha |r|)$ \cite{Rasmussen:2006:Gaussian_Processes}.
-This corresponds to the particular case of the CGPCM where $h(t)=\sqrt{2 \alpha} \exp(- \alpha |t|)$.
+\subsection{Speech}
 
-% \subsection{RLC Series Circuit}
-% Consider the voltage response $f(t)$ of a series circuit consisting of a resistance $R$, capacitance $C$, and inductance $L$ excited by current $x(t)$ with unity PSD due to thermal noise of the resistance.
-Application of Kirchhoff's voltage law \cite{Irwin:2010:Basic_Engineering_Circuit_Analysis} yields that
-% \begin{align*}
-%      L f'(t) + R f(t)+\frac{1}{C}\int^t f(t)\id{t} = x(t).
-% \end{align*}
-% Assume that the circuit is critically damped and that $R=2$ and $L=1$.
-Then $f(t)$ is of the form of \cref{eq:model} where $h(t) = t \exp(-t)$.
-Therefore, according to \cref{mod:cgpcm2}, it holds that $f(t) \sim \mathcal{GP}(0,k_{f\cond h}(t-t'))$ where
-% \begin{align*}
-%     k_{f\cond h}(r) = \frac{1}{4}(1+|r|)\exp(-|r|).
-% \end{align*}
+\begin{figure*}[t]
+    \centering
+    GPCM
+    \begin{tabularx}{\linewidth}{
+            >{\centering}>{\hsize=.43\hsize}X  % ??
+            >{\centering}>{\hsize=.2\hsize}X
+            >{\centering}>{\hsize=.2\hsize}X
+            >{\centering}>{\hsize=.2\hsize}X
+        }
+        $f\cond h$ & $h$ & $k_{f\cond h}$ & PSD of $k_{f\cond h}$
+    \end{tabularx}
+    \includegraphics[width=\linewidth, height=3cm]{resources/learning_known_kernels_causal_sample_gpcm_cropped.pdf}
+    CGPCM
+    \includegraphics[width=\linewidth, height=3cm]{resources/learning_known_kernels_causal_sample_cgpcm_cropped.pdf}
+    \caption{Toy.
+Here {\color{red}red} denotes truth; {\color{darkgreen}green} denotes observations, the emperical autocorrelation, and the periodogram for respectively the function $f\cond h$, the kernel $k_{f\cond h}$, and PSD of $k_{f\cond h}$; {\color{blue}blue} denotes quantities learned through the MF approximation; and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation.}
+    \label{fig:toy_causal_sample}
+\end{figure*}
 
-% {\color{red} Experiment.}
-
-% \subsection{Unvoiced Fricatives}
-% Unvoiced fricatives are consonants produced by forcing air through a narrow constriction in the vocal tract, thereby creating a noisy, turbulant sound \cite{Taylor:2009:Text-To-Speech_Synthesis}.
+\begin{figure*}[t]
+    \centering
+    GPCM
+    \begin{tabularx}{\linewidth}{
+            >{\centering}>{\hsize=.43\hsize}X  % ??
+            >{\centering}>{\hsize=.2\hsize}X
+            >{\centering}>{\hsize=.2\hsize}X
+            >{\centering}>{\hsize=.2\hsize}X
+        }
+        $f\cond h$ & $h$ & $k_{f\cond h}$ & PSD of $k_{f\cond h}$
+    \end{tabularx}
+    \includegraphics[width=\linewidth, height=3cm]{resources/learning_known_kernels_acausal_sample_gpcm_cropped.pdf}
+    CGPCM
+    \includegraphics[width=\linewidth, height=3cm]{resources/learning_known_kernels_acausal_sample_cgpcm_cropped.pdf}
+    \caption{Toy.
+Here {\color{red}red} denotes truth; {\color{darkgreen}green} denotes observations, the emperical autocorrelation, and the periodogram for respectively the function $f\cond h$, kernel $k_{f\cond h}$, and power spectral density $\log \mathcal{F}\{k_{f\cond h}\}$; {\color{blue}blue} denotes quantities learned through the MF approximation; and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation.}
+    \label{fig:toy_scausal_sample}
+\end{figure*}
 
 \subsection{Head-Related Transfer Function}
-A head-related transfer function (HRTF) is a transfer function that describes how an ear receives sound from a certain point in space.
-
-\subsection{Speech}
-
-{\color{red} Experiment.}
-
-{\color{red} Experiment.}
-
-\subsection{Population Growth}
+The head-related transfer function (HRTF) describes how the reception of sound by the ear depends on the position from which the sound originates.
 

### 2017-02-24 doc/paper.tex
Prettyfied images

@@
 \Crefname{equation}{Equation}{Equations}
+\crefname{figure}{figure}{figures}
+\Crefname{figure}{Figure}{Figures}
 
@@
 \subsection{Learning Known Kernels}
-
-{\color{red} Experiment.}
-
-\subsection{Speech}
-
 \begin{figure*}[t]
@@
     \end{tabularx}
-    \includegraphics[width=\linewidth, height=3cm]{resources/learning_known_kernels_causal_sample_gpcm_cropped.pdf}
+    \includegraphics[width=\linewidth, height=3cm]{resources/cropped/learning_known_kernels_causal_sample_gpcm.pdf}
     CGPCM
-    \includegraphics[width=\linewidth, height=3cm]{resources/learning_known_kernels_causal_sample_cgpcm_cropped.pdf}
+    \includegraphics[width=\linewidth, height=3cm]{resources/cropped/learning_known_kernels_causal_sample_cgpcm.pdf}
     \caption{Toy.
-Here {\color{red}red} denotes truth; {\color{darkgreen}green} denotes observations, the emperical autocorrelation, and the periodogram for respectively the function $f\cond h$, the kernel $k_{f\cond h}$, and PSD of $k_{f\cond h}$; {\color{blue}blue} denotes quantities learned through the MF approximation; and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation.}
+Here {\color{red}red} denotes truth; {\color{darkgreen}green} denotes observations, the emperical autocorrelation, and the periodogram for respectively the function $f\cond h$, the kernel $k_{f\cond h}$, and PSD of $k_{f\cond h}$; {\color{blue}blue} denotes quantities learned through the MF approximation; and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation.
+Signals normalised to unity power.}
     \label{fig:toy_causal_sample}
@@
     \end{tabularx}
-    \includegraphics[width=\linewidth, height=3cm]{resources/learning_known_kernels_acausal_sample_gpcm_cropped.pdf}
+    \includegraphics[width=\linewidth, height=3cm]{resources/cropped/learning_known_kernels_acausal_sample_gpcm.pdf}
     CGPCM
-    \includegraphics[width=\linewidth, height=3cm]{resources/learning_known_kernels_acausal_sample_cgpcm_cropped.pdf}
+    \includegraphics[width=\linewidth, height=3cm]{resources/cropped/learning_known_kernels_acausal_sample_cgpcm.pdf}
     \caption{Toy.
-Here {\color{red}red} denotes truth; {\color{darkgreen}green} denotes observations, the emperical autocorrelation, and the periodogram for respectively the function $f\cond h$, kernel $k_{f\cond h}$, and power spectral density $\log \mathcal{F}\{k_{f\cond h}\}$; {\color{blue}blue} denotes quantities learned through the MF approximation; and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation.}
-    \label{fig:toy_scausal_sample}
+Here {\color{red}red} denotes truth; {\color{darkgreen}green} denotes observations, the emperical autocorrelation, and the periodogram for respectively the function $f\cond h$, kernel $k_{f\cond h}$, and PSD of $k_{f\cond h}$; {\color{blue}blue} denotes quantities learned through the MF approximation; and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation.
+Signals normalised to unity power.}
+    \label{fig:toy_acausal_sample}
 \end{figure*}
+
+Check out \cref{fig:toy_causal_sample,fig:toy_acausal_sample}.
+
+\subsection{Speech}
 

### 2017-02-24 doc/resources/crop.sh

@@ -1,0 +1,4 @@
+#!/usr/bin/env zsh
+for f in *.pdf; do
+    pdfcrop --margins 0 $f cropped/$f
+done

### 2017-02-24 doc/paper.tex
Typo on PSD

@@
         }
-        $f\cond h$ & $h$ & $k_{f\cond h}$ & PSD of $k_{f\cond h}$
+        $f\cond h$ & $h$ & $k_{f\cond h}$ & PSD of $f\cond h$
     \end{tabularx}
@@
     \caption{Toy.
-Here {\color{red}red} denotes truth; {\color{darkgreen}green} denotes observations, the emperical autocorrelation, and the periodogram for respectively the function $f\cond h$, the kernel $k_{f\cond h}$, and PSD of $k_{f\cond h}$; {\color{blue}blue} denotes quantities learned through the MF approximation; and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation.
+Here {\color{red}red} denotes truth; {\color{darkgreen}green} denotes observations, the emperical autocorrelation, and the periodogram for respectively the function $f\cond h$, the kernel $k_{f\cond h}$, and PSD of $f\cond h$; {\color{blue}blue} denotes quantities learned through the MF approximation; and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation.
 Signals normalised to unity power.}
@@
     \caption{Toy.
-Here {\color{red}red} denotes truth; {\color{darkgreen}green} denotes observations, the emperical autocorrelation, and the periodogram for respectively the function $f\cond h$, kernel $k_{f\cond h}$, and PSD of $k_{f\cond h}$; {\color{blue}blue} denotes quantities learned through the MF approximation; and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation.
+Here {\color{red}red} denotes truth; {\color{darkgreen}green} denotes observations, the emperical autocorrelation, and the periodogram for respectively the function $f\cond h$, kernel $k_{f\cond h}$, and PSD of $f\cond h$; {\color{blue}blue} denotes quantities learned through the MF approximation; and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation.
 Signals normalised to unity power.}

### 2017-02-24 doc/paper.tex
Fixed autocorrelation in plots

@@
 \section{Modelling Causally-Generated Stationary Time Series}
-Consider modelling the stationary time series $f$.
-Since many dynamical systems in nature can accurately be described by an initial value problem or a linear system, we let $f$ be the solution of a time-invariant linear initial value problem with causal Green's function $h$ and forcing function $x$, or equivalently the system response of a time-invariant linear system with causal impulse response $h$ and excitation $x$.
+Consider modelling a stationary time series $f$.
+Motivated by the fact that many dynamical systems in nature can accurately be described by an initial value problem or a linear system, we let $f$ be the solution of a time-invariant linear initial value problem with causal Green's function $h$ and forcing function $x$, or equivalently the system response of a time-invariant linear system with causal impulse response $h$ and excitation $x$.
 It then holds that
@@
 \section{Experiments}
-We used TensorFlow \cite{Abadi:2016:TensorFlow_A_System_for_Large-Scale} to implement the CGPCM.\footnote{The implementation can be found at \url{https://github.com/wesselb/cgpcm}.} The main issue in implementing the CGPCM is that computation of the matrices in \cref{app:moments_f} requires the bivariate normal cumulative density function (BNCDF) \cite{Genz:2004:Numerical_Computation_of_Rectangular_Bivariate}.
+We used TensorFlow \cite{Abadi:2016:TensorFlow_A_System_for_Large-Scale} to implement the CGPCM.\footnote{The implementation can be found at
+% \url{https://github.com/wesselb/cgpcm}.
+--- ---
+} The main issue in implementing the CGPCM is that computation of the matrices in \cref{app:moments_f} requires evaluation of the bivariate normal cumulative density function (BNCDF) \cite{Genz:2004:Numerical_Computation_of_Rectangular_Bivariate}.
 Since the BNCDF is not available in TensorFlow, we used the implementation of the BNCDF from the R package \texttt{pbivnorm}\footnote{See \url{https://github.com/brentonk/pbivnorm}.} to implement the BNCDF and its gradient in TensorFlow.
-We furthermore implementated an algorithm that symbolically solves the integrals in \cref{app:moments_f} and constructs the corresponding matrices afterwards; namely, manually solving these expressions  otherwise would be tedious and error prone.
+We furthermore implementated an algorithm that symbolically solves the integrals in \cref{app:moments_f} and constructs the corresponding matrices afterwards---manually solving these expressions would be tedious and error prone.
 
@@
     CGPCM
+    \begin{tabularx}{\linewidth}{
+            >{\centering}>{\hsize=.43\hsize}X  % ??
+            >{\centering}>{\hsize=.2\hsize}X
+            >{\centering}>{\hsize=.2\hsize}X
+            >{\centering}>{\hsize=.2\hsize}X
+        }
+        $f\cond h$ & $h$ & $k_{f\cond h}$ & PSD of $f\cond h$
+    \end{tabularx}
     \includegraphics[width=\linewidth, height=3cm]{resources/cropped/learning_known_kernels_causal_sample_cgpcm.pdf}
@@
 Here {\color{red}red} denotes truth; {\color{darkgreen}green} denotes observations, the emperical autocorrelation, and the periodogram for respectively the function $f\cond h$, the kernel $k_{f\cond h}$, and PSD of $f\cond h$; {\color{blue}blue} denotes quantities learned through the MF approximation; and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation.
-Signals normalised to unity power.}
+All signals are normalised to unity power.}
     \label{fig:toy_causal_sample}
@@
     CGPCM
+    \begin{tabularx}{\linewidth}{
+            >{\centering}>{\hsize=.43\hsize}X  % ??
+            >{\centering}>{\hsize=.2\hsize}X
+            >{\centering}>{\hsize=.2\hsize}X
+            >{\centering}>{\hsize=.2\hsize}X
+        }
+        $f\cond h$ & $h$ & $k_{f\cond h}$ & PSD of $f\cond h$
+    \end{tabularx}
     \includegraphics[width=\linewidth, height=3cm]{resources/cropped/learning_known_kernels_acausal_sample_cgpcm.pdf}
     \caption{Toy.
-Here {\color{red}red} denotes truth; {\color{darkgreen}green} denotes observations, the emperical autocorrelation, and the periodogram for respectively the function $f\cond h$, kernel $k_{f\cond h}$, and PSD of $f\cond h$; {\color{blue}blue} denotes quantities learned through the MF approximation; and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation.
-Signals normalised to unity power.}
+Here {\color{red}red} denotes truth; {\color{darkgreen}green} denotes observations, the emperical autocorrelation, and the periodogram for respectively the function $f\cond h$, the kernel $k_{f\cond h}$, and PSD of $f\cond h$; {\color{blue}blue} denotes quantities learned through the MF approximation; and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation.
+All signals are normalised to unity power.}
     \label{fig:toy_acausal_sample}

### 2017-02-24 doc/paper.tex
HRIR results

@@
 Note that the sample appears smooth for $h(0)=0$ and becomes rougher as $|h(0)|$ increases.
+{\color{red} something about pointyness?}
 
@@
 
-\subsection{Head-Related Transfer Function}
-The head-related transfer function (HRTF) describes how the reception of sound by the ear depends on the position from which the sound originates.
+\subsection{Head-Related Impulse Response}
+The head-related impulse response (HRIR) describes how the reception of sound by the ear depends on the position from which the sound originates.
+\begin{figure*}[t]
+    \centering
+    \begin{tabularx}{\linewidth}{>{\centering}X>{\centering}X}
+        GPCM & CGPCM
+    \end{tabularx}
+    \begin{tabularx}{\linewidth}{>{\centering}X>{\centering}X>{\centering}X>{\centering}X}
+        $k_{f\cond h}$ & $h$ & $k_{f\cond h}$ & $h$
+    \end{tabularx}
+    \includegraphics[width=.49\linewidth, height=3cm]{resources/hrtf_gpcm.pdf}
+    \includegraphics[width=.49\linewidth, height=3cm]{resources/hrtf_cgpcm.pdf}
+    \caption{HRIR.
+Here {\color{red}red} denotes truth, {\color{blue}blue} denotes quantities learned through the MF approximation, and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation.
+All signals are normalised to unity power.}
+    \label{fig:hrir}
+\end{figure*}
 

### 2017-02-24 doc/paper.tex
Making pass through...

@@
 \section{Modelling Causally-Generated Stationary Time Series}
+%
 Consider modelling a stationary time series $f$.
@@
 \end{align}
-We denote integration from negative infinity or to positive infinity by omitting the respective limit from the integral.
+For parsimony, we denote integration from negative infinity or to positive infinity by omitting the respective limit from the integral throughout.
 Note that $f(t)$ depends on $x(\tau)$ only for $\tau \le t$; this reveals \cref{eq:model}'s causal nature.
@@
 \begin{align}
-    &k_{f\cond h}(t,t')\nonumber \\
-    &\quad= \int^t h(t- \tau)\int^{t'}h(t'- \tau')
+    k_{f\cond h}(t,t')&= \!\int^t \! \int^{t'}\!h(t- \tau) \vspace{-1mm} h(t'- \tau')
         \mathbb{E}[ x(\tau) x(\tau')]\id{\tau'} \id{\tau} \nonumber \\
-    &\quad= \int^{t \land t'} h(t - \tau) h(t' - \tau) \id{\tau} \nonumber \\
-    &\quad= \int_0 h(|t - t'| + \tau)h(\tau) \id{\tau} \label{eq:kernel} \\
-    &\quad= k_{f\cond h}(t-t') \nonumber
+    &= \int^{t \land t'} \! h(t - \tau) h(t' - \tau) \id{\tau} \nonumber \\
+    &= \int_0 h(|t - t'| + \tau)h(\tau) \id{\tau} = k_{f\cond h}(t-t')  \label{eq:kernel} 
 \end{align}
-where $t \land t'$ denotes the minimum of $t$ and $t'$. Alternatively, \cref{eq:model} can in that case be interpreted as the continuous-time generalisation of a causal moving-average filter.
+where $t \land t'$ denotes the minimum of $t$ and $t'$.
+The restriction of $x$ to white noise is not overly restrictive since any causal stationary process can be represented with sufficiently rich $h$.
+An alternative lens through which to view the model is as the continuous-time generalisation of a causal moving-average filter.
 
-Since $h$ consitutes an unknown function, we choose to model it using a Gaussian process; that is, we let $h \sim \mathcal{GP}(0,k_h(t,t'))$.
-In choosing $k_h(t,t')$, we require that $f$ has finite power, since every real-world signal has so as well.
-Following \citet{Tobar:2015:Learning_Stationary}, this can be achieved by letting $h$ decay to zero at infinity sufficiently quickly: let $h\cond g = w g$ where $w(t)= \exp(- \alpha t^2)$ and $g \sim \mathcal{GP}(0,k_g(t-t'))$ has finite power; in that case,
+One of the focusses in this work is to develop flexible prior distributions over the filter $h$ that can support essentially arbitrarily complex structure.
+We therefore choose to model it using a Gaussian process $h \sim \mathcal{GP}(0,k_h(t,t'))$.
+The covariance function describing the prior over the filter$k_h(t,t')$ should be carefully chosen.
+One important constraint on the covariance function comes from the fact that every real-world signal $f$ has finite power.
+This can be satisfied by letting the filter variance decay to zero at infinity sufficiently quickly \citet{Tobar:2015:Learning_Stationary}: let $h\cond g = w g$ where $w(t)= \exp(- \alpha t^2)$ and $g \sim \mathcal{GP}(0,k_g(t-t'))$ has finite power; in that case, % RET there's a question as to whether exponential decay might be more appropriate in the current case
 \begin{align*}
@@
 Equivalently, we let $k_h(t,t')=\exp(- \alpha (t^2 + t^{\prime 2}))k_g(t-t')$.
-To retain flexibility in the prior on $h$, we let $k_g$ be an exponentiated square.
+To retain flexibility in the prior on $h$, we let $k_g$ be an exponentiated quadratic.
 We thus have that $k_h(t,t')=\exp(- \alpha (t^2 + t^{\prime 2}) - \gamma(t-t')^2)$.

### 2017-02-24 doc/paper.tex
Some edits

--- 2017-02-24 doc/paper.tex
@@
 \subsection{Roughness of Sample Paths}
+\label{subsec:roughness}
 We show that the CGCPM models both differentiable and nondifferentiable phenomena, the latter in various levels of roughness.
@@
 \section{Inference}
+\label{sec:inference}
 Let $y(t)\cond \sim \mathcal{N}(f(t),\sigma^2)$ for all $t$ be a noisy version of $f$. Given some observations $e=(y(t_1),\ldots,y(t_n))$ of $y$, we wish to compute $p(f,h,x\cond e)$.
@@
 \section{Experiments}
-We used TensorFlow \cite{Abadi:2016:TensorFlow_A_System_for_Large-Scale} to implement the CGPCM.\footnote{The implementation can be found at
-% \url{https://github.com/wesselb/cgpcm}.
---- ---
-} The main issue in implementing the CGPCM is that computation of the matrices in \cref{app:moments_f} requires evaluation of the bivariate normal cumulative density function (BNCDF) \cite{Genz:2004:Numerical_Computation_of_Rectangular_Bivariate}.
-Since the BNCDF is not available in TensorFlow, we used the implementation of the BNCDF from the R package \texttt{pbivnorm}\footnote{See \url{https://github.com/brentonk/pbivnorm}.} to implement the BNCDF and its gradient in TensorFlow.
-We furthermore implementated an algorithm that symbolically solves the integrals in \cref{app:moments_f} and constructs the corresponding matrices afterwards---manually solving these expressions would be tedious and error prone.
+In this section we test the CGPCM by applying it to synthetic and real-world signals.
+We aim to show that in certain situations the causality constraint provides a useful inductive bias that leads to better predictions.
 
-\subsection{Learning Known Kernels}
-\begin{figure*}[t]
-    \centering
-    GPCM
-    \begin{tabularx}{\linewidth}{
-            >{\centering}>{\hsize=.43\hsize}X  % ??
-            >{\centering}>{\hsize=.2\hsize}X
-            >{\centering}>{\hsize=.2\hsize}X
-            >{\centering}>{\hsize=.2\hsize}X
-        }
-        $f\cond h$ & $h$ & $k_{f\cond h}$ & PSD of $f\cond h$
-    \end{tabularx}
-    \includegraphics[width=\linewidth, height=3cm]{resources/cropped/learning_known_kernels_causal_sample_gpcm.pdf}
-    CGPCM
-    \begin{tabularx}{\linewidth}{
-            >{\centering}>{\hsize=.43\hsize}X  % ??
-            >{\centering}>{\hsize=.2\hsize}X
-            >{\centering}>{\hsize=.2\hsize}X
-            >{\centering}>{\hsize=.2\hsize}X
-        }
-        $f\cond h$ & $h$ & $k_{f\cond h}$ & PSD of $f\cond h$
-    \end{tabularx}
-    \includegraphics[width=\linewidth, height=3cm]{resources/cropped/learning_known_kernels_causal_sample_cgpcm.pdf}
-    \caption{Toy.
-Here {\color{red}red} denotes truth; {\color{darkgreen}green} denotes observations, the emperical autocorrelation, and the periodogram for respectively the function $f\cond h$, the kernel $k_{f\cond h}$, and PSD of $f\cond h$; {\color{blue}blue} denotes quantities learned through the MF approximation; and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation.
-All signals are normalised to unity power.}
-    \label{fig:toy_causal_sample}
-\end{figure*}
+% We used TensorFlow \cite{Abadi:2016:TensorFlow_A_System_for_Large-Scale} to implement the CGPCM.\footnote{The implementation can be found at
+% % \url{https://github.com/wesselb/cgpcm}.
+% --- ---
+% } The main issue in implementing the CGPCM is that computation of the matrices in \cref{app:moments_f} requires evaluation of the bivariate normal cumulative density function (BNCDF) \cite{Genz:2004:Numerical_Computation_of_Rectangular_Bivariate}.
+Unfortunately, the BNCDF is not available in TensorFlow; we therefore we used the implementation of the BNCDF from the R package \texttt{pbivnorm}\footnote{See \url{https://github.com/brentonk/pbivnorm}.} to implement the BNCDF and its gradient in TensorFlow.
+Moreover, we implementated an algorithm that symbolically solves the integrals in \cref{app:moments_f}; hence we provide no closed-form expression for those integrals.
+
+\subsection{Learning Rough Signals}
+Consider a Gaussian process with a fixed kernel. Then, according to \cref{mod:cgpcm2}'s interpretation, the CGPCM should be able to learn this kernel. We compare the CGPCM and the GPCM.
 
@@
 
-Check out \cref{fig:toy_causal_sample,fig:toy_acausal_sample}.
+\Cref{fig:toy_acausal_sample} shows the results for fitting the GPCM and CGPCM to a noisy random sample from the GPCM.
+The sample consists of 300 data points.
+We used 51 evenly-spaced inducing points for $h$ and 80 evenly-spaced inducing point for $x$.
+All parameters were initialised randomly and the procedure from \cref{sec:inference} was followed.
 
-\subsection{Speech}
+We make several observations.
+First, the GPCM and CGPCM appear to both succesfully fit the function $f \cond h$ and the kernel $k_{f\cond h}$.
+This shows that the CGPCM can indeed learn kernels.
+Second, the sample from the GPCM appears smooth, which means that the learned filter $h$ for the CGPCM should be nearly zero near zero.
+A quick inspection of \cref{fig:toy_acausal_sample} verifies that this is indeed the case.
+Finally, the learned filter $h(t)$ for the GPCM appears uncertain for $t\le 0$.
+This a direct consequence of the causality contraint: \cref{eq:model} shows that $f\cond h$ does not depend on $h(t)$ for $t \le 0$.
 
-\subsection{Head-Related Impulse Response}
+\begin{figure*}[t]
+    \centering
+    GPCM
+    \begin{tabularx}{\linewidth}{
+            >{\centering}>{\hsize=.43\hsize}X  % ??
+            >{\centering}>{\hsize=.2\hsize}X
+            >{\centering}>{\hsize=.2\hsize}X
+            >{\centering}>{\hsize=.2\hsize}X
+        }
+        $f\cond h$ & $h$ & $k_{f\cond h}$ & PSD of $f\cond h$
+    \end{tabularx}
+    \includegraphics[width=\linewidth, height=3cm]{resources/cropped/learning_known_kernels_causal_sample_gpcm.pdf}
+    CGPCM
+    \begin{tabularx}{\linewidth}{
+            >{\centering}>{\hsize=.43\hsize}X  % ??
+            >{\centering}>{\hsize=.2\hsize}X
+            >{\centering}>{\hsize=.2\hsize}X
+            >{\centering}>{\hsize=.2\hsize}X
+        }
+        $f\cond h$ & $h$ & $k_{f\cond h}$ & PSD of $f\cond h$
+    \end{tabularx}
+    \includegraphics[width=\linewidth, height=3cm]{resources/cropped/learning_known_kernels_causal_sample_cgpcm.pdf}
+    \caption{Toy.
+Here {\color{red}red} denotes truth; {\color{darkgreen}green} denotes observations, the emperical autocorrelation, and the periodogram for respectively the function $f\cond h$, the kernel $k_{f\cond h}$, and PSD of $f\cond h$; {\color{blue}blue} denotes quantities learned through the MF approximation; and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation.
+All signals are normalised to unity power.}
+    \label{fig:toy_causal_sample}
+\end{figure*}
+
+\Cref{fig:toy_causal_sample} shows the results for fitting the GPCM and CGPCM to instead a noiseless random sample from the CGPCM.
+Since the drawn filter $h$ satisfies $|h(0)|>0$, the sample is rough and thus appears noisy.
+The challenge for the models is to not mistake this sample roughness for observation noise.
+
+Observe that the CGPCM fits the filter $h$ and kernel $k_{f \cond h}$ well, whereas GPCM's fit is more mediocre.
+Despite roughly getting the correlation structure right, the GPCM mistakes the sample roughness for observation noisy and provides a poor fit of the function $f\cond h$.
+The CGPCM, on the other hand, correctly recognises the sample roughness and provides an excellent fit of $f \cond h$.
+This discrepancy is explained by inspecting the fit of the PSD: the GPCM is unable to fit the higher frequencies in the sample, whereas the CGPCM does so succesfully.
+
+The key difference between \cref{fig:toy_acausal_sample,fig:toy_causal_sample} is that in the former noise is added independently, whereas in the latter noise is part of the process dynamics.
+Therefore our experiments indicate that the GPCM and CPGCM can both succesfully handle independently added noise, but only the CGPCM succesfully fits noisy dynamics.
+
+% \subsection{Power Spectral Density Estimation}
+% \begin{figure*}[t]
+%     \centering
+%     \begin{tabularx}{\linewidth}{>{\centering}X>{\centering}X}
+%         GPCM & CGPCM
+%     \end{tabularx}
+%     \begin{tabularx}{\linewidth}{>{\centering}X>{\centering}X>{\centering}X>{\centering}X}
+%         $k_{f\cond h}$ & PSD of $f\cond h$ & $k_{f\cond h}$ & PSD of $f \cond h$
+%     \end{tabularx}
+%     % \includegraphics[width=.49\linewidth, height=3cm]{resources/psd_gpcm.pdf}
+%     % \includegraphics[width=.49\linewidth, height=3cm]{resources/psd_cgpcm.pdf}
+%     \caption{PSD.
+Here {\color{red}red} denotes truth, {\color{blue}blue} denotes quantities learned through the MF approximation, and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation.
+All signals are normalised to unity power.}
+%     \label{fig:psd}
+% \end{figure*}
+
+% Check out \cref{fig:psd}.
+
+\subsection{Head-Related Impulse Response Estimation}
 The head-related impulse response (HRIR) describes how the reception of sound by the ear depends on the position from which the sound originates.
+In particular, the received sound is convolved with the HRIR before it is sensed by the ear.
+Since this is exactly
+
 \begin{figure*}[t]
@@
 
+Check out \cref{fig:hrir}.
+
 \section{Discussion}
@@
 Inference is performed by finding an initial approximation through a variational MF approach, and then refining this approximation through a variational SMF approach combined with ESS.
-The GPCM has been validated {\color{red} discussion of results of experiments here.}
+The GPCM has been validated on synthesis and real-world signals.
+{\color{red} some here}
 
-In future research on the CGPCM, various directions can be explored: First, the model could be extented to multidimensional input and output spaces by following a construction similar to that by \citet{Bruinsma:2016:GGPCM}.
+Various directions could be explored in future research: First, the model could be extented to multidimensional input and output spaces by following a construction similar to that by \citet{Bruinsma:2016:GGPCM}.
 Second, the application of more efficient sparse approximation for $h$ and $x$, such as the tree-structured approximation \cite{Bui:2014:Tree-Structured_Gaussian}, would scale the CGPCM to larger data sets.

### 2017-02-24 doc/paper.tex
Fixed motivation for first experiment

@@
 \subsection{Learning Rough Signals}
-Consider a Gaussian process with a fixed kernel. Then, according to \cref{mod:cgpcm2}'s interpretation, the CGPCM should be able to learn this kernel. We compare the CGPCM and the GPCM.
+In \cref{subsec:roughness} we showed that the CGPCM models both smooth and rough isgnals, whereas the GPCM only models smooth signals.
+We perform two experiments that show the significance of this modelling capability.
 
@@
 In particular, the received sound is convolved with the HRIR before it is sensed by the ear.
-Since this is exactly
+We consider the problem of determining the HRIR given some signal sensed by the ear.
 

### 2017-02-24 doc/paper.tex
Updated section 2

--- 2017-02-24 doc/paper.tex
@@
 We thus have that $k_h(t,t')=\exp(- \alpha (t^2 + t^{\prime 2}) - \gamma(t-t')^2)$.
-Note that $\alpha$ determines $k_f$'s length scale and $\gamma$ its flexibility.
+Note that $\alpha$ determines the typical temporal extent of the filter and $\gamma$ the typical time-scale over which it varies.
+These assumptions also serve to make inference well-posed; without them, any shifted version of the filter results in an identical statistical model for the data meaning the model is unidentifiable.
 
-Through specifying a prior on $x$ and a prior on $h$ we have specified a prior on $f$.
+The prior distributions over $x$ and $h$ induce a prior distribution on $f$.
 Further including a scale $\sigma_f$ to control $f$'s prior power, we call this prior on $f$ the Causal Gaussian Process Convolution Model (CGPCM).

### 2017-02-24 doc/paper.tex
More edits

@@
         h &\sim \mathcal{GP}(0, k_h(t,t')), \\
-        f \cond h &\sim \mathcal{GP}(0,  \sigma_f^2\int_0 h(|t-t'|+\tau)h(\tau)\id{\tau}).
+        f \cond h &\sim \mathcal{GP}\left(0,  \sigma_f^2\int_0 h(|t-t'|+\tau)h(\tau)\id{\tau} \right).
     \end{align*}
@@
 \subsection{Sampling from the CGPCM}
-Sampling from the CGPCM is difficult because the integrals in \cref{mod:cgpcm,mod:cgpcm2} depend on the entirety of $h$ and $x$.
-To resolve this issue, we follow \citet{Tobar:2015:Learning_Stationary} and approximate \cref{eq:kernel} using quadrature \cite{Minka:2000:Quadrature_GP} by conditioning on finitely many values $u=(h(t_{u,1}),\ldots,h(t_{u,n_u}))\sim \mathcal{N}(0,K_u)$ of $h$:
+Sampling from the CGPCM is challenging because the integrals in \cref{mod:cgpcm,mod:cgpcm2} depend on the entirety of $h$ and $x$.
+To resolve this issue, we follow \citet{Tobar:2015:Learning_Stationary} and approximate \cref{eq:kernel} using Bayesian quadrature \cite{Minka:2000:Quadrature_GP} by conditioning on finitely many values $u=(h(t_{u,1}),\ldots,h(t_{u,n_u}))\sim \mathcal{N}(0,K_u)$ of $h$ which induces a  distribution over covariance functions.
+The mean of this distribution is an accurate approximation for $k_{f\cond h}$ if sufficient conditioning values are used,
 \begin{align*}
@@
     &= \int_0 k_h(|r| + \tau, \tau) \id{\tau} \\
-    &\phan{=}+ \tr M^u \int_{0} k_h(t_u, |r|+ \tau)k_h(\tau, t_u^T) \id{\tau}
+    &\phan{=}+ \mathrm{trace} \left( M^u \int_{0} k_h(t_u, |r|+ \tau)k_h(\tau, t_u^T) \id{\tau} \right)
 \end{align*}
 where $M^{u}=K_u^{-1}uu^T K_u^{-1}-K_u$.
-Unlike the GPCM, the kernel approximation $\mathbb{E}[k_{f\cond h}(r)\cond u]$ of the CGPCM will not simply be a mixture of Gaussians; it will be a more complicated expression instead.
+This expression was used to compute the model samples shown in figure 1.
+%Unlike the GPCM, the kernel approximation $\mathbb{E}[k_{f\cond h}(r)\cond u]$ of the CGPCM will not simply be a mixture of Gaussians; it will be a more complicated expression instead.
 
@@
 \label{subsec:roughness}
-We show that the CGCPM models both differentiable and nondifferentiable phenomena, the latter in various levels of roughness.
-This modelling capability holds false for the GPCM, whose sample paths are always differentiable.
-
-Let $h$ be fixed and decay to zero at infinity.
+We now show that the CGCPM can capture both differentiable and non-differentiable phenomena, the latter in various levels of roughness.
+In this regard, the model is more flexible than the original GPCM whose sample paths are always differentiable.
+Let $h$ be a fixed filter that decays to zero at infinity.
 We claim that $f$'s sample paths are almost surely everywhere differentiable if $h(0)=0$, and almost surely nowhere differentiable if $h(0)\neq 0$.
-To show this, let
+To show this, let $g(r) = \int_0 h(r + \tau) h(\tau) \id{\tau}$ so that
 \begin{align*}
-    g(r) = \int_0 h(r + \tau) h(\tau) \id{\tau}
-\end{align*}
-so that
-\begin{align*}
-    k_{f\cond h}(r)
-    &= g(|r|) \\
-    &= g(0) + g'(0)|r| + \frac{1}{2}g''(0)r^2 + \O(|r|^3).
+    k_{f\cond h}(r) = \! g(|r|) \! = \! g(0) \! + \! g'(0)|r| + \frac{1}{2}g''(0)r^2 + \O(|r|^3).
 \end{align*}
@@
 \begin{align*}
-    g'(0)
-    &= \int_0 h'(\tau) h(\tau) \id{\tau} \\
-    &= \ll[ h^2(\tau)\rr]_0 - \int_0 h(\tau) h'(\tau) \id{\tau},
+    g'(0)= \int_0 h'(\tau) h(\tau) \id{\tau} = \ll[ h^2(\tau)\rr]_0 - \int_0 h(\tau) h'(\tau) \id{\tau},
 \end{align*}
@@
 
-In summary, the CGPCM models differentiable phenoma if $h(0)=0$ and nondifferentiable phenoma if $h(0)\neq 0$. In the latter case $|h(0)|$ quantifies the level of roughness.
+In summary, the CGPCM models differentiable phenomena if $h(0)=0$ and non-differentiable phenoma if $h(0)\neq 0$. In the latter case $|h(0)|$ quantifies the level of roughness.
 

### 2017-02-24 doc/paper.tex
final edits for now

@@
     \begin{align*}
-        x &\sim \mathcal{GP}(0,\delta(t-t')), \\
-        h &\sim \mathcal{GP}(0, k_h(t,t')), \\
-        f\cond h, x &= \sigma_f \int^t h(t- \tau)x(\tau)\id{\tau}.
+        &x \sim \mathcal{GP}(0,\delta(t-t')), \;
+        h \sim \mathcal{GP}(0, k_h(t,t')), \\
+        &f\cond h, x = \sigma_f \int^t h(t- \tau)x(\tau)\id{\tau}.
     \end{align*}
@@
 
-In summary, the CGPCM models differentiable phenomena if $h(0)=0$ and non-differentiable phenoma if $h(0)\neq 0$. In the latter case $|h(0)|$ quantifies the level of roughness.
+In summary, the CGPCM models differentiable phenomena if $h(0)=0$ and non-differentiable phenomena if $h(0)\neq 0$.
+In the latter case $|h(0)|$ quantifies the level of roughness.
+Performing inference for the filter will therefore automatically infer the differentiability of the underlying process from possibly noisy data.
 

### 2017-02-24 doc/paper.tex
Updates

--- 2017-02-24 doc/paper.tex
@@
 We thus have that $k_h(t,t')=\exp(- \alpha (t^2 + t^{\prime 2}) - \gamma(t-t')^2)$.
-Note that $\alpha$ determines the typical temporal extent of the filter and $\gamma$ the typical time-scale over which it varies.
-These assumptions also serve to make inference well-posed; without them, any shifted version of the filter results in an identical statistical model for the data meaning the model is unidentifiable.
+Note that $\alpha$ determines $k_f$'s length scale and $\gamma$ its flexibility.
 
-The prior distributions over $x$ and $h$ induce a prior distribution on $f$.
+Through specifying a prior on $x$ and a prior on $h$ we have specified a prior on $f$.
 Further including a scale $\sigma_f$ to control $f$'s prior power, we call this prior on $f$ the Causal Gaussian Process Convolution Model (CGPCM).
@@
 \subsection{Learning Rough Signals}
-Consider a Gaussian process with a fixed kernel. Then, according to \cref{mod:cgpcm2}'s interpretation, the CGPCM should be able to learn this kernel. We compare the CGPCM and the GPCM.
+In \cref{subsec:roughness} we showed that the CGPCM models both smooth and rough signals, whereas the GPCM only models smooth signals.
+We perform two experiments that show the significance of this modelling capability.
 
@@
 The sample consists of 300 data points.
-We used 51 evenly-spaced inducing points for $h$ and 80 evenly-spaced inducing point for $x$.
+We used 51 evenly-spaced inducing points for $h$ and 80 evenly-spaced inducing points for $x$.
 All parameters were initialised randomly and the procedure from \cref{sec:inference} was followed.
@@
 
-\Cref{fig:toy_causal_sample} shows the results for fitting the GPCM and CGPCM to instead a noiseless random sample from the CGPCM.
+\Cref{fig:toy_causal_sample} shows the results for fitting the GPCM and CGPCM to instead a \textit{noiseless} random sample from the CGPCM.
 Since the drawn filter $h$ satisfies $|h(0)|>0$, the sample is rough and thus appears noisy.
@@
 Observe that the CGPCM fits the filter $h$ and kernel $k_{f \cond h}$ well, whereas GPCM's fit is more mediocre.
-Despite roughly getting the correlation structure right, the GPCM mistakes the sample roughness for observation noisy and provides a poor fit of the function $f\cond h$.
+Despite roughly getting the correlation structure right, the GPCM mistakes the sample roughness for observation noise and provides a poor fit of the function $f\cond h$.
 The CGPCM, on the other hand, correctly recognises the sample roughness and provides an excellent fit of $f \cond h$.
@@
 We have presented the CGPCM as a model for causally-generated stationary time series.
-Inference is performed by finding an initial approximation through a variational MF approach, and then refining this approximation through a variational SMF approach combined with ESS.
+Inference in the CGPCM is performed by finding an initial approximation through a variational MF approach, and then refining this approximation through a variational SMF approach combined with ESS.
 The GPCM has been validated on synthesis and real-world signals.

### 2017-02-24 doc/paper.tex
Progress

@@
 Sampling from the CGPCM is challenging because the integrals in \cref{mod:cgpcm,mod:cgpcm2} depend on the entirety of $h$ and $x$.
-To resolve this issue, we follow \citet{Tobar:2015:Learning_Stationary} and approximate \cref{eq:kernel} using Bayesian quadrature \cite{Minka:2000:Quadrature_GP} by conditioning on finitely many values $u=(h(t_{u,1}),\ldots,h(t_{u,n_u}))\sim \mathcal{N}(0,K_u)$ of $h$ which induces a  distribution over covariance functions.
+To resolve this issue, we follow \citet{Tobar:2015:Learning_Stationary} and approximate \cref{eq:kernel} using Bayesian quadrature \cite{Minka:2000:Quadrature_GP} by conditioning on finitely many values $u=(h(t_{u,1}),\ldots,h(t_{u,n_u}))\sim \mathcal{N}(0,K_u)$ of $h$, which induces a  distribution over covariance functions.
 The mean of this distribution is an accurate approximation for $k_{f\cond h}$ if sufficient conditioning values are used,
@@
 where $M^{u}=K_u^{-1}uu^T K_u^{-1}-K_u$.
-This expression was used to compute the model samples shown in figure 1.
+This expression was used to compute the model samples shown in \cref{fig:interpolation}.
 %Unlike the GPCM, the kernel approximation $\mathbb{E}[k_{f\cond h}(r)\cond u]$ of the CGPCM will not simply be a mixture of Gaussians; it will be a more complicated expression instead.
@@
 \label{subsec:roughness}
-We now show that the CGCPM can capture both differentiable and non-differentiable phenomena, the latter in various levels of roughness.
+We now show that the CGCPM can capture both differentiable and nondifferentiable phenomena, the latter in various levels of roughness.
 In this regard, the model is more flexible than the original GPCM whose sample paths are always differentiable.
@@
 Note that the sample appears smooth for $h(0)=0$ and becomes rougher as $|h(0)|$ increases.
-{\color{red} something about pointyness?}
 
@@
 
-Variational mean-field (MF) approaches to inference, like the one just presented, are often computationally efficient, but they are known to suffer from certain biases \cite{MacKay:2002:Information_Theory_Learning,Turner:2011:Two_Problems_With_Variational_Expectation,Murphy:2012:Probabilistic_Perspective}.
+Variational mean-field approaches to inference, like the one just presented, are often computationally efficient, but they are known to suffer from certain biases \cite{MacKay:2002:Information_Theory_Learning,Turner:2011:Two_Problems_With_Variational_Expectation,Murphy:2012:Probabilistic_Perspective}.
 We further refine the MF approximation to alleviate these biases.
@@
 In this section we test the CGPCM by applying it to synthetic and real-world signals.
-We aim to show that in certain situations the causality constraint provides a useful inductive bias that leads to better predictions.
+We show that in certain situations the causality constraint provides a useful inductive bias that leads to better predictions.
 
@@
 \subsection{Learning Rough Signals}
-In \cref{subsec:roughness} we showed that the CGPCM models both smooth and rough signals, whereas the GPCM only models smooth signals.
+In \cref{subsec:roughness} we showed that the CGPCM models both smooth and rough signals, whereas the GPCM models only smooth signals.
 We perform two experiments that show the significance of this modelling capability.
@@
     \includegraphics[width=\linewidth, height=3cm]{resources/cropped/learning_known_kernels_acausal_sample_cgpcm.pdf}
-    \caption{Toy.
-Here {\color{red}red} denotes truth; {\color{darkgreen}green} denotes observations, the emperical autocorrelation, and the periodogram for respectively the function $f\cond h$, the kernel $k_{f\cond h}$, and PSD of $f\cond h$; {\color{blue}blue} denotes quantities learned through the MF approximation; and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation.
-All signals are normalised to unity power.}
+    \caption{Predictions of the GPCM and CGPCM after learning a noisy random sample from the GPCM.
+This sample consists of 300 data points.
+We used 51 evenly-spaced inducing points for $h$ and 80 evenly-spaced inducing points for $x$; inducing points are indicated by black dots in the figure.
+All parameters were initialised randomly and the procedure from \cref{sec:inference} was followed.
+All signals are normalised to unity power.
+Finally, {\color{red}red} denotes truth; {\color{darkgreen}green} denotes observations, the emperical autocorrelation, and the periodogram for respectively the function $f\cond h$, the kernel $k_{f\cond h}$, and the PSD of $f\cond h$; {\color{blue}blue} denotes quantities learned through the MF approximation; and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation.
+For any learned quantity it holds that lines correspond to means and areas correpond to 95\% central credible regions.}
     \label{fig:toy_acausal_sample}
@@
 \Cref{fig:toy_acausal_sample} shows the results for fitting the GPCM and CGPCM to a noisy random sample from the GPCM.
-The sample consists of 300 data points.
-We used 51 evenly-spaced inducing points for $h$ and 80 evenly-spaced inducing points for $x$.
-All parameters were initialised randomly and the procedure from \cref{sec:inference} was followed.
-
 We make several observations.
@@
     \includegraphics[width=\linewidth, height=3cm]{resources/cropped/learning_known_kernels_causal_sample_cgpcm.pdf}
-    \caption{Toy.
-Here {\color{red}red} denotes truth; {\color{darkgreen}green} denotes observations, the emperical autocorrelation, and the periodogram for respectively the function $f\cond h$, the kernel $k_{f\cond h}$, and PSD of $f\cond h$; {\color{blue}blue} denotes quantities learned through the MF approximation; and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation.
-All signals are normalised to unity power.}
+    \caption{Predictions of the GPCM and CGPCM after learning a noiseless random sample from the CGPCM.
+The sample consists of 300 data points.
+We used 51 evenly-spaced inducing points for $h$ and 80 evenly-spaced inducing points for $x$; inducing points are indicated by black dots in the figure.
+All parameters were initialised randomly and the procedure from \cref{sec:inference} was followed.
+All signals are normalised to unity power.
+Finally, {\color{red}red} denotes truth; {\color{darkgreen}green} denotes observations, the emperical autocorrelation, and the periodogram for respectively the function $f\cond h$, the kernel $k_{f\cond h}$, and the PSD of $f\cond h$; {\color{blue}blue} denotes quantities learned through the MF approximation; and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation.
+For any learned quantity it holds that lines correspond to means and areas correpond to 95\% central credible regions.}
     \label{fig:toy_causal_sample}
@@
 The challenge for the models is to not mistake this sample roughness for observation noise.
-
-Observe that the CGPCM fits the filter $h$ and kernel $k_{f \cond h}$ well, whereas GPCM's fit is more mediocre.
+Observe that the CGPCM fits the filter $h$ and kernel $k_{f \cond h}$ well, whereas GPCM's fit is mediocre.
 Despite roughly getting the correlation structure right, the GPCM mistakes the sample roughness for observation noise and provides a poor fit of the function $f\cond h$.
@@
 The key difference between \cref{fig:toy_acausal_sample,fig:toy_causal_sample} is that in the former noise is added independently, whereas in the latter noise is part of the process dynamics.
-Therefore our experiments indicate that the GPCM and CPGCM can both succesfully handle independently added noise, but only the CGPCM succesfully fits noisy dynamics.
+Therefore \cref{fig:toy_acausal_sample,fig:toy_causal_sample} indicate that the GPCM and CPGCM can both handle independently added noise, but only the CGPCM fits noisy dynamics.
 
-% \subsection{Power Spectral Density Estimation}
+Finally, note that predictions through the SMF approximation consistently give better-calibrated credible regions than predictions through the MF approximation.
+We will further investigate issue in the following experiment.
+
+\subsection{Speech Modelling}
+Coming up.
+
 % \begin{figure*}[t]
@@
 The head-related impulse response (HRIR) describes how the reception of sound by the ear depends on the position from which the sound originates.
-In particular, the received sound is convolved with the HRIR before it is sensed by the ear.
-We consider the problem of determining the HRIR given some signal sensed by the ear.
+Specifically, the received sound is convolved with the HRIR before it is sensed by the ear, and the HRIR depends on the direction from which the sound comes.
+We consider the problem of inferring the HRIR after some sound has been sensed.
+Suppose that the sound is background noise and can be modelled by white noise.
+Then assuming a Gaussian process prior over the HRIR recovers exactly the CGPCM; therefore, in that case, we can infer the HRIR by inferring the filter $h$ in the CGPCM.
 
@@
     \includegraphics[width=.49\linewidth, height=3cm]{resources/hrtf_cgpcm.pdf}
-    \caption{HRIR.
+    \caption{Predictions of the GPCM and CGPCM after learning a HRIR convolved with white noise.
+The sample consists of 500 data points.
+We used 101 evenly-spaced inducing points for $h$ and 150 evenly-spaced inducing points for $x$; inducing points are indicated by black dots in the figure.
+All parameters were initialised randomly and the procedure from \cref{sec:inference} was followed.
+All signals are normalised to unity power.
 Here {\color{red}red} denotes truth, {\color{blue}blue} denotes quantities learned through the MF approximation, and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation.
-All signals are normalised to unity power.}
+For any learned quantity it holds that lines correspond to means and areas correpond to 95\% central credible regions.}
     \label{fig:hrir}
@@
 
-Check out \cref{fig:hrir}.
+We convolve a HRIR from a KEMAR\footnote{See \url{http://sound.media.mit.edu/resources/KEMAR.html}.} dummy head microphone with white noise to simulate the signal that would have been sensed by the ear.
+\Cref{fig:hrir} shows the results of fitting the GPCM and CGPCM to the resulting signal.
+Despite the fact that both models correctly infer the kernel $k_{f \cond h}$, only the CGPCM is able to provide a good prediction for the HRIR.
+This shows that the inductive bias built in by the causality constraint in fact does help when dealing with causal phenomena.
 
 \section{Discussion}
-We have presented the CGPCM as a model for causally-generated stationary time series.
-Inference in the CGPCM is performed by finding an initial approximation through a variational MF approach, and then refining this approximation through a variational SMF approach combined with ESS.
-The GPCM has been validated on synthesis and real-world signals.
-{\color{red} some here}
+We have presented the CGPCM as a model of causal, spectrally complex dynamical phenomena.
+Inference in the CGPCM is performed in a two-step procedure: first a collapsed varational MF approach is used to obtain an initial approximation, and then this approximation is refined through a variational SMF approach combined with ESS.
+The CGPCM has been tested and validated on synthesised and real-world signals and showed promising results.
+In particular, the CGPCM showed the capability of modelling rough signals without attributing this roughness to observation noise.
 

### 2017-02-24 doc/paper.tex
Added prior over kernels visualisation

@@
 Finally, a sample $f\cond h$ is drawn from $\mathcal{GP}(0,k_{f \cond h}(t-t'))$.
+\Cref{fig:prior} visualises the corresponding prior over kernels.
 
 The CPGCM is similar to the latent force model presented by \citet{Alvarez:2009:Latent_Force_Models}; however, whereas we let $h$ be free-form and let $x$ be white noise, \citet{Alvarez:2009:Latent_Force_Models} specify $h$ deterministically and let $x$ be free-form.
+
+ \begin{figure}[t]
+    \centering
+    \begin{tabularx}{\linewidth}{>{\centering}X>{\centering}X}
+        GPCM & CGPCM
+    \end{tabularx}
+    \includegraphics[width=\linewidth, height=3cm]{resources/prior.pdf}
+    \caption{Prior distribution over kernels in the GPCM and CGPCM}
+    \label{fig:prior}
+\end{figure}
 
@@
 Inference in the CGPCM is performed in a two-step procedure: first a collapsed varational MF approach is used to obtain an initial approximation, and then this approximation is refined through a variational SMF approach combined with ESS.
-The CGPCM has been tested and validated on synthesised and real-world signals and showed promising results.
-In particular, the CGPCM showed the capability of modelling rough signals without attributing this roughness to observation noise.
+The CGPCM has been tested and validated on synthesised and real-world signals and shows encouraging results.
+In particular, the CGPCM shows the capability of modelling rough signals without attributing this roughness to observation noise.
 

### 2017-02-24 doc/paper.tex
Updated figures and prior over PSD + connection to experiments

@@
 % Break in equations...
-% \allowdisplaybreaks
+\allowdisplaybreaks
 
@@
 Finally, a sample $f\cond h$ is drawn from $\mathcal{GP}(0,k_{f \cond h}(t-t'))$.
-\Cref{fig:prior} visualises the corresponding prior over kernels.
+\Cref{fig:prior} visualises the corresponding prior over kernels, and \cref{fig:prior_psd} visualises the corresponding prior over PSDs.
+Note that in \cref{fig:prior_psd} the prior for the GPCM has a low-pass structure, whereas that for the CGPCM decays more slowly.
 
@@
     \end{tabularx}
-    \includegraphics[width=\linewidth, height=3cm]{resources/prior.pdf}
-    \caption{Prior distribution over kernels in the GPCM and CGPCM}
+    \includegraphics[width=\linewidth, height=3cm]{resources/cropped/prior.pdf}
+    \caption{Prior distribution over kernels in the GPCM and CGPCM. Lines correspond to means and gradients correspond to uncertainty.}
     \label{fig:prior}
+\end{figure}
+
+\begin{figure}[t]
+    \centering
+    \begin{tabularx}{\linewidth}{>{\centering}X>{\centering}X}
+        GPCM & CGPCM
+    \end{tabularx}
+    \includegraphics[width=\linewidth, height=3cm]{resources/cropped/prior_psd.pdf}
+    \caption{Prior distribution over PSDs in the GPCM and CGPCM. Lines correspond to means and gradients correspond to uncertainty.}
+    \label{fig:prior_psd}
 \end{figure}
@@
 \begin{figure*}[t]
-    % \vskip 0.2in
+    \vskip 0.1in
     \centering
-    \includegraphics[width=\linewidth]{resources/interpolation.pdf}
+    \includegraphics[width=\linewidth]{resources/cropped/interpolation.pdf}
     \caption{Generative process of the CGPCM.
@@
 The CGPCM, on the other hand, correctly recognises the sample roughness and provides an excellent fit of $f \cond h$.
-This discrepancy is explained by inspecting the fit of the PSD: the GPCM is unable to fit the higher frequencies in the sample, whereas the CGPCM does so succesfully.
+This discrepancy is explained by inspecting the fit of the PSD: the GPCM is unable to fit the higher frequencies in the sample, whereas the CGPCM does so succesfully; this is consistent with the priors depicted in \cref{fig:prior_psd}.
 

### 2017-02-24 doc/paper.tex
Various updates

@@
 \Cref{fig:prior} visualises the corresponding prior over kernels, and \cref{fig:prior_psd} visualises the corresponding prior over PSDs.
-Note that in \cref{fig:prior_psd} the prior for the GPCM has a low-pass structure, whereas that for the CGPCM decays more slowly.
+Note that the prior over PSDs for the GPCM has a confident low-pass structure, whereas that for the CGPCM has support for more slowly decaying spectra.
 
@@
     \begin{tabularx}{\linewidth}{>{\centering}X>{\centering}X}
-        GPCM & CGPCM
+        ~~GPCM & ~~~~~~CGPCM
     \end{tabularx}
     \includegraphics[width=\linewidth, height=3cm]{resources/cropped/prior_psd.pdf}
-    \caption{Prior distribution over PSDs in the GPCM and CGPCM. Lines correspond to means and gradients correspond to uncertainty.}
+    \caption{Prior distribution over PSDs in the GPCM and CGPCM. The unit on the $y$ axis is decibel. Lines correspond to means and gradients correspond to uncertainty.}
     \label{fig:prior_psd}
@@
 
-\subsection{Speech Modelling}
+\subsection{Speech}
 Coming up.

### 2017-02-24 doc/paper.tex
First draft

@@
 The CGPCM models the kernel of a Gaussian process nonparametrically and has an inductive bias towards causally-generated time series.
-We develop an algorithm to perform inference and apply the CGPCM to tasks involving real-world data.
+We develop an algorithm to perform inference and demonstrate the CGPCM on synthetic and real-world signals.
 \end{abstract}
@@
 When developing statistical models for such phenomena this prior knowledge---causality and spectral richness---should be leveraged in order to exclude any unrealisable system from the model prior whilst at the same time allowing the model to have the capacity to capture varied spectral content that may be only slowly revealed as more data are seen.
-The goal of this paper is to develop Gaussian Process (GP) models together with associated inference and learning schemes that serve this purpose.
+The goal of this paper is to develop Gaussian process (GP) models together with associated inference and learning schemes that serve this purpose.
 
-A widely-used model for stationary time series are Gaussian Processes, which place a prior distribution over the latent function underlying the time series $f:\R \to \R$ by assuming that any finite collection of function values $f(t_1),\ldots,f(t_n)$ is multivariate-Gaussian distributed.
-The key modelling decision in using a Gaussian Process is the choice of covariance $k_f(t,t')$ between any two function values $f(t)$ and $f(t')$---$k_f$ is also called the \textit{kernel} of $f$.
-This choice of kernel\footnote{The mean of the function is assumed to be zero without loss of generality.
-\cite{Rasmussen:2006:Gaussian_Processes}} encodes prior information about the function $f$.
+Gaussian processes are a widely-used model for stationary time series.
+They place a prior distribution over the latent function underlying the time series $f:\R \to \R$ by assuming that any finite collection of function values $f(t_1),\ldots,f(t_n)$ is multivariate-Gaussian distributed.
+The key modelling decision in using a Gaussian process is the choice of covariance $k_f(t,t')$ between any two function values $f(t)$ and $f(t')$---$k_f$ is also called the \textit{kernel} of $f$.
+This choice of kernel encodes prior information about the function $f$.
 Consequently, a large research effort has been devoted towards developing flexible kernels \cite{Duvenaud:2014:Automatic_Construction,Wilson:2013:Spectral_Mixture,Tobar:2015:Learning_Stationary,Tobar:2015:Inter-Domain_Inducing}.
 
-Of particular interest for the current application are GP models that use nonparametric models of the power spectra of the GPs---doubly non-parametric models---as they have the capacity to flexibly model signals with arbitrary spectral complexity.
+Of particular interest for the current application are GP models that use nonparametric models of the power spectra of the GPs---doubly nonparametric models---as they have the capacity to flexibly model signals with arbitrary spectral complexity.
 One approach employs a Dirichlet Process Mixture Model for the power spectra \cite{Oliva:2015:Bayesian_Nonparametric_Kernel-Learning}, but it is not clear whether it is possible to build the causality constraint into such a construction.
-An alternative approach instead induces a nonparametric power spectrum by placing a Gaussian Process prior over the system's impulse response function \cite{Tobar:2015:Learning_Stationary}.
+An alternative approach instead induces a nonparametric power spectrum by placing a Gaussian process prior over the system's impulse response function \cite{Tobar:2015:Learning_Stationary}.
 Critically, the so-called Gaussian Process Convolution Model (GPCM), does not incorporate a causality constraint---the impulse response in question should be thought of as an acausal spatial response.
@@
 We therefore choose to model it using a Gaussian process $h \sim \mathcal{GP}(0,k_h(t,t'))$.
-The covariance function describing the prior over the filter$k_h(t,t')$ should be carefully chosen.
+The covariance function describing the prior over the filter $k_h(t,t')$ should be carefully chosen.
 One important constraint on the covariance function comes from the fact that every real-world signal $f$ has finite power.
-This can be satisfied by letting the filter variance decay to zero at infinity sufficiently quickly \citet{Tobar:2015:Learning_Stationary}: let $h\cond g = w g$ where $w(t)= \exp(- \alpha t^2)$ and $g \sim \mathcal{GP}(0,k_g(t-t'))$ has finite power; in that case, % RET there's a question as to whether exponential decay might be more appropriate in the current case
+This can be satisfied by letting the filter variance decay to zero at infinity sufficiently quickly \cite{Tobar:2015:Learning_Stationary}: let $h\cond g = w g$ where $w(t)= \exp(- \alpha t^2)$ and $g \sim \mathcal{GP}(0,k_g(t-t'))$ has finite power; in that case, % RET there's a question as to whether exponential decay might be more appropriate in the current case
 \begin{align*}
@@
 \Cref{fig:prior} visualises the corresponding prior over kernels, and \cref{fig:prior_psd} visualises the corresponding prior over PSDs.
-Note that the prior over PSDs for the GPCM has a confident low-pass structure, whereas that for the CGPCM has support for more slowly decaying spectra.
+Note that the prior over PSDs for the GPCM confidently has a low-pass structure, whereas that for the CGPCM has support for more slowly decaying spectra.
 
@@
 \end{figure}
-
 \begin{figure}[t]
@@
 where $M^{u}=K_u^{-1}uu^T K_u^{-1}-K_u$.
-This expression was used to compute the model samples shown in \cref{fig:interpolation}.
+This expression was used to compute the model samples shown in \cref{fig:prior,fig:prior_psd,fig:interpolation}.
 %Unlike the GPCM, the kernel approximation $\mathbb{E}[k_{f\cond h}(r)\cond u]$ of the CGPCM will not simply be a mixture of Gaussians; it will be a more complicated expression instead.
@@
 \end{align*}
-Again, setting the variations $\delta \mathcal{L} / \delta q(u)$ and $\delta \mathcal{L} / \delta q(z\cond u)$ to zero and solving for for respectively $q(u)$ and $q(z\cond u)$ yields that
+Again, setting the variations $\delta \mathcal{L} / \delta q(u)$ and $\delta \mathcal{L} / \delta q(z\cond u)$ to zero and solving for respectively $q(u)$ and $q(z\cond u)$ yields that
 \begin{align}
@@
 \section{Experiments}
-In this section we test the CGPCM by applying it to synthetic and real-world signals.
-We show that in certain situations the causality constraint provides a useful inductive bias that leads to better predictions.
+We test the CGPCM by applying it to synthetic and real-world signals. We show that in certain situations the causality constraint provides a useful inductive bias that leads to better predictions.
 
@@
     \caption{Predictions of the GPCM and CGPCM after learning a noisy random sample from the GPCM.
-This sample consists of 300 data points.
+The sample consists of 300 data points.
 We used 51 evenly-spaced inducing points for $h$ and 80 evenly-spaced inducing points for $x$; inducing points are indicated by black dots in the figure.
@@
 All signals are normalised to unity power.
-Finally, {\color{red}red} denotes truth; {\color{darkgreen}green} denotes observations, the emperical autocorrelation, and the periodogram for respectively the function $f\cond h$, the kernel $k_{f\cond h}$, and the PSD of $f\cond h$; {\color{blue}blue} denotes quantities learned through the MF approximation; and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation.
+{\color{red}Red} denotes truth; {\color{darkgreen}green} denotes observations, the emperical autocorrelation, and the periodogram for respectively the function $f\cond h$, the kernel $k_{f\cond h}$, and the PSD of $f\cond h$; {\color{blue}blue} denotes quantities learned through the MF approximation; and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation.
 For any learned quantity it holds that lines correspond to means and areas correpond to 95\% central credible regions.}
@@
 
-Finally, note that predictions through the SMF approximation consistently give better-calibrated credible regions than predictions through the MF approximation.
-We will further investigate issue in the following experiment.
-
-\subsection{Speech}
-Coming up.
+Finally, note that in nearly every case predictions through the SMF approximation have better calibrated credible regions than predictions through the MF approximation.
+% \subsection{Speech}
+% Coming up.
 
@@
 \subsection{Head-Related Impulse Response Estimation}
-The head-related impulse response (HRIR) describes how the reception of sound by the ear depends on the position from which the sound originates.
+The head-related impulse response (HRIR) describes how the ear's reception of sound depends on the direction from which the sound comes.
 Specifically, the received sound is convolved with the HRIR before it is sensed by the ear, and the HRIR depends on the direction from which the sound comes.
@@
 Suppose that the sound is background noise and can be modelled by white noise.
-Then assuming a Gaussian process prior over the HRIR recovers exactly the CGPCM; therefore, in that case, we can infer the HRIR by inferring the filter $h$ in the CGPCM.
+Then assuming a Gaussian process prior over the HRIR recovers the CGPCM exactly; therefore, in that case, we can infer the HRIR via inferring the filter $h$ in the CGPCM.
 
@@
 All signals are normalised to unity power.
-Here {\color{red}red} denotes truth, {\color{blue}blue} denotes quantities learned through the MF approximation, and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation.
+{\color{red}Red} denotes truth, {\color{blue}blue} denotes quantities learned through the MF approximation, and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation.
 For any learned quantity it holds that lines correspond to means and areas correpond to 95\% central credible regions.}
@@
 Inference in the CGPCM is performed in a two-step procedure: first a collapsed varational MF approach is used to obtain an initial approximation, and then this approximation is refined through a variational SMF approach combined with ESS.
-The CGPCM has been tested and validated on synthesised and real-world signals and shows encouraging results.
+The CGPCM has been tested on synthesised and real-world signals and shows encouraging results.
 In particular, the CGPCM shows the capability of modelling rough signals without attributing this roughness to observation noise.

### 2017-02-24 doc/paper.tex
Rephrasing

@@
 \section{Experiments}
-We test the CGPCM by applying it to synthetic and real-world signals. We show that in certain situations the causality constraint provides a useful inductive bias that leads to better predictions.
+We test the CGPCM by applying it to synthetic and real-world signals. We show that in certain situations the causality constraint provides an useful inductive bias that leads to better predictions.
 
@@
 
-Finally, note that in nearly every case predictions through the SMF approximation have better calibrated credible regions than predictions through the MF approximation.
+Finally, note that the SMF approximation consistently has better calibrated credible regions compared to the MF approximation.
+For example, in the filter $h$, kernel $k_{f\cond h}$, PSD of $f \cond h$ of the CGPCM in \cref{fig:toy_causal_sample}, the MF approximation fails to capture the truth in its credible region, whereas the SMF approximation does capture the truth.
 % \subsection{Speech}

### 2017-02-24 doc/paper.tex
Rephrasing in abstract
Rephrasing of captions of fig 1 and 2

@@
 
-\usepackage{icml2017}  % option: accepted
+\usepackage[accepted]{icml2017}  % option: accepted
 \icmltitlerunning{Learning Causally-Generated Stationary Time Series}
@@
 The CGPCM models the kernel of a Gaussian process nonparametrically and has an inductive bias towards causally-generated time series.
-We develop an algorithm to perform inference and demonstrate the CGPCM on synthetic and real-world signals.
+We develop enhanced variational inference and learning schemes and demonstrate the CGPCM on synthetic and real-world signals.
 \end{abstract}
@@
 \end{align}
-For parsimony, we denote integration from negative infinity or to positive infinity by omitting the respective limit from the integral throughout.
+For parsimony, we denote integration from negative infinity and to positive infinity by omitting the respective limit from the integral throughout.
 Note that $f(t)$ depends on $x(\tau)$ only for $\tau \le t$; this reveals \cref{eq:model}'s causal nature.
@@
     \includegraphics[width=\linewidth, height=3cm]{resources/cropped/prior.pdf}
-    \caption{Prior distribution over kernels in the GPCM and CGPCM. Lines correspond to means and gradients correspond to uncertainty.}
+    \caption{Prior distribution over kernels in the GPCM and CGPCM. Lines correspond to means and gradients indicate marginal variance.}
     \label{fig:prior}
@@
     \includegraphics[width=\linewidth, height=3cm]{resources/cropped/prior_psd.pdf}
-    \caption{Prior distribution over PSDs in the GPCM and CGPCM. The unit on the $y$ axis is decibel. Lines correspond to means and gradients correspond to uncertainty.}
+    \caption{Prior distribution over PSDs in the GPCM and CGPCM. The unit on the $y$ axis is decibel. Lines correspond to means and gradients indicate marginal variance.}
     \label{fig:prior_psd}

### 2017-02-24 doc/paper.tex
More minor edits

@@
 Note that $\alpha$ determines the typical temporal extent of the filter and $\gamma$ the typical time-scale over which it varies.
-These assumptions also serve to make inference well-posed; without them, any shifted version of the filter results in an identical statistical model for the data meaning the model is unidentifiable.
+These assumptions also serve to make inference well-posed; without them, any shifted version of the filter results in an identical statistical model for the data, meaning that the model is unidentifiable.
 
@@
 
-\Cref{mod:cgpcm2} reveals the CGCPM as a Gaussian process in which the kernel, or equivalently the power spectral density (PSD), is modelled nonparametrically.
+\Cref{mod:cgpcm2} reveals the CGCPM as a Gaussian process in which the kernel, or equivalently the power spectral density, is modelled nonparametrically.
 \Cref{fig:interpolation} illustrates the generative process of the CGPCM: First, a filter $h$ is generated.
@@
     \caption{Predictions of the GPCM and CGPCM after learning a noisy random sample from the GPCM.
-The sample consists of 300 data points.
+The learned sample consists of 300 data points.
 We used 51 evenly-spaced inducing points for $h$ and 80 evenly-spaced inducing points for $x$; inducing points are indicated by black dots in the figure.
@@
     \caption{Predictions of the GPCM and CGPCM after learning a noiseless random sample from the CGPCM.
-The sample consists of 300 data points.
+The learned sample consists of 300 data points.
 We used 51 evenly-spaced inducing points for $h$ and 80 evenly-spaced inducing points for $x$; inducing points are indicated by black dots in the figure.
@@
     \caption{Predictions of the GPCM and CGPCM after learning a HRIR convolved with white noise.
-The sample consists of 500 data points.
+The learned sample consists of 500 data points.
 We used 101 evenly-spaced inducing points for $h$ and 150 evenly-spaced inducing points for $x$; inducing points are indicated by black dots in the figure.
@@
     &\quad= -\frac{n}{2}\log 2 \pi \sigma^2 - \frac{1}{2 \sigma^2} \la e^{2}(t) - 2 \sigma_f u^T A^{hx}(t) e(t) z\\
-    &\quad\phan{=}  + \sigma_f^2 ( b(t) + u^T B^{h}(t) u+ z^T B^{x}(t) z + (u^T A^{hx}(t) z)^2 ) \ra_t
+    &\quad\phan{=}  + \sigma_f^2 ( b(t) + u^T B^{h}(t) u+ z^T B^{x}(t) z \\
+    &\quad\phan{=}  + (u^T A^{hx}(t) z)^2 ) \ra_t
 \end{align*}

### 2017-02-24 doc/paper.tex
Slight edits to inference section.

@@
 \end{align*}
-is some interdomain transformation with filter $r$ \cite{Lazaro-Gredilla:2009:Inter-Domain_Gaussian_Processes_for_Sparse,Alvarez:2010:Efficient_Multioutput_Gaussian_Processes_Through,Tobar:2015:Learning_Stationary}.
-In choosing $r$, we make sure that $s=T[x]$ has power at the majority of frequencies present in the signal that we aim to model.
-We follow \citet{Tobar:2015:Learning_Stationary} and let $r(t)=\exp(-\omega t^2)$, which fits baseband signals.
+is some interdomain transformation of the white-noise process with filter $r$ \cite{Lazaro-Gredilla:2009:Inter-Domain_Gaussian_Processes_for_Sparse,Alvarez:2010:Efficient_Multioutput_Gaussian_Processes_Through,Tobar:2015:Learning_Stationary}.
+The inter-domain transformation is necessary because white noise is uncorrelated and so placing inducing points directly in this domain would not be able to capture meaningful structure in the posterior.
+In choosing $r$, we make sure that $s=T[x]$ has power at the majority of frequencies present in the signal that we aim to model, thereby enabling the inducing variables to capture posterior dependencies in the components of the white noise that pass through the filter and therefore affect the data.
+We follow \citet{Tobar:2015:Learning_Stationary} and let $r(t)=\exp(-\omega t^2)$, which fits the low frequency components of the posterior.
 
-Specifically, let the \textit{mean-field} (MF) approximation $q(f,u,z)=p(f\cond u, z)q(u)q(z)$ of $p(f,u,z\cond e)$ be such that it is closest to $p(f,u,z\cond e)$ in Kullback-Leibler divergence.
-Rearranging
+Let the \textit{mean-field} (MF) approximation $q(f,u,z)=p(f\cond u, z)q(u)q(z)$ of $p(f,u,z\cond e)$ be such that it is closest to $p(f,u,z\cond e)$ in Kullback-Leibler divergence. Rearranging
 \begin{align*}
@@
 
-To optimise $\mathcal{L}$ with respect to $q(u)$ and $q(z)$ we set their respective variations $\delta \mathcal{L} / \delta q(u)$ and $\delta \mathcal{L} / \delta q(z)$ to zero; then solving for $q(u)$ and $q(z)$ yields that
+To optimise $\mathcal{L}$ with respect to $q(u)$ and $q(z)$ we set their respective variations $\delta \mathcal{L} / \delta q(u)$ and $\delta \mathcal{L} / \delta q(z)$ to zero; then solving for $q(u)$ and $q(z)$ yields
 \begin{align}
@@
 
-We can, however, significantly speed up the optimisation by solving for either $q(u)$ or $q(z)$ analytically. Substituting the optimal form of $q(z)$ back in $\mathcal{L}$ yields that
+One of the new contributions of this work is the finding that we can significantly speed up the optimisation by solving for either $q(u)$ or $q(z)$ analytically.
+Substituting the optimal form of $q(z)$ back in $\mathcal{L}$ yields
 \begin{align}
@@
 Variational mean-field approaches to inference, like the one just presented, are often computationally efficient, but they are known to suffer from certain biases \cite{MacKay:2002:Information_Theory_Learning,Turner:2011:Two_Problems_With_Variational_Expectation,Murphy:2012:Probabilistic_Perspective}.
-We further refine the MF approximation to alleviate these biases.
+We further refine the MF approximation to alleviate these biases, which is the second major improvement to inference and learning provided by this work.
 
@@
 
-Following an argument similar, we again derive the corresponding ELBO:
+Following a similar argument to the above, we again derive the corresponding ELBO:
 \begin{align*}
@@
 \end{align*}
-Again, setting the variations $\delta \mathcal{L} / \delta q(u)$ and $\delta \mathcal{L} / \delta q(z\cond u)$ to zero and solving for respectively $q(u)$ and $q(z\cond u)$ yields that
+Again, setting the variations $\delta \mathcal{L} / \delta q(u)$ and $\delta \mathcal{L} / \delta q(z\cond u)$ to zero and solving for respectively $q(u)$ and $q(z\cond u)$ yields 
 \begin{align}

### 2017-02-24 doc/paper.tex
small edits

@@
 Observe that the CGPCM fits the filter $h$ and kernel $k_{f \cond h}$ well, whereas GPCM's fit is mediocre.
-Despite roughly getting the correlation structure right, the GPCM mistakes the sample roughness for observation noise and provides a poor fit of the function $f\cond h$.
+Despite approximately capturing the correlation structure, the GPCM mistakes the sample roughness for observation noise and provides a poor fit of the function $f\cond h$.
 The CGPCM, on the other hand, correctly recognises the sample roughness and provides an excellent fit of $f \cond h$.

### 2017-02-24 doc/paper.tex
altered the HRTF section

@@
 \subsection{Head-Related Impulse Response Estimation}
-The head-related impulse response (HRIR) describes how the ear's reception of sound depends on the direction from which the sound comes.
-Specifically, the received sound is convolved with the HRIR before it is sensed by the ear, and the HRIR depends on the direction from which the sound comes.
-We consider the problem of inferring the HRIR after some sound has been sensed.
+The head-related impulse response (HRIR) describes how sounds entering the ear canal are filtered by the outer ear.
+The form of the filter depends on the location of the sound source.
+The filter introduces complex spectral cues into the sounds that are used by the brain to infer source location.
+We consider the problem of inferring the HRIR from an incoming sound pressure waveform.
 Suppose that the sound is background noise and can be modelled by white noise.
-Then assuming a Gaussian process prior over the HRIR recovers the CGPCM exactly; therefore, in that case, we can infer the HRIR via inferring the filter $h$ in the CGPCM.
+Then assuming a Gaussian process prior over the HRIR we recover the CGPCM exactly; in other words, we can infer the HRIR via inferring the filter $h$ in the CGPCM.
 
@@
 Despite the fact that both models correctly infer the kernel $k_{f \cond h}$, only the CGPCM is able to provide a good prediction for the HRIR.
-This shows that the inductive bias built in by the causality constraint in fact does help when dealing with causal phenomena.
+This shows that the inductive bias built in by the causality constraint does in fact help when dealing with causal phenomena.
 

### 2017-02-24 doc/paper.tex
Extra remark

--- 2017-02-24 doc/paper.tex
@@
 
-Let the \textit{mean-field} (MF) approximation $q(f,u,z)=p(f\cond u, z)q(u)q(z)$ of $p(f,u,z\cond e)$ be such that it is closest to $p(f,u,z\cond e)$ in Kullback-Leibler divergence. Rearranging
+Let the \textit{mean-field} (MF) approximation $q(f,u,z)=p(f\cond u, z)q(u)q(z)$ of $p(f,u,z\cond e)$ be such that it is closest to $p(f,u,z\cond e)$ in Kullback-Leibler divergence.
+Contrary to the formulation of \citet{Tobar:2015:Learning_Stationary}, we have analytically integrated out $h$ and $x$ in $p(f\cond u, z)=\int p(f, h, x\cond u, z)\id{h}\id{x}$ prior to performing inference.
+It is indeed the case that $p(f\cond u, z)$ is intractable, but its first two moment can actually be computed, and only those we need in formulating our inference scheme.
+Rearranging
 \begin{align*}
@@
 A quick inspection of \cref{fig:toy_acausal_sample} verifies that this is indeed the case.
-Finally, the learned filter $h(t)$ for the GPCM appears uncertain for $t\le 0$.
+Finally, the learned filter $h(t)$ for the CGPCM appears uncertain for $t\le 0$.
 This a direct consequence of the causality contraint: \cref{eq:model} shows that $f\cond h$ does not depend on $h(t)$ for $t \le 0$.
@@
 Finally, note that the SMF approximation consistently has better calibrated credible regions compared to the MF approximation.
-For example, in the filter $h$, kernel $k_{f\cond h}$, PSD of $f \cond h$ of the CGPCM in \cref{fig:toy_causal_sample}, the MF approximation fails to capture the truth in its credible region, whereas the SMF approximation does capture the truth.
+For example, the MF approximation fails to capture the truth in its credible region for the filter $h$, kernel $k_{f\cond h}$, and PSD of $f \cond h$ of the CGPCM in \cref{fig:toy_causal_sample}; in contrast, the SMF approximation does capture the truth.
 % \subsection{Speech}

### 2017-02-24 doc/paper.tex
Typo

@@
 Contrary to the formulation of \citet{Tobar:2015:Learning_Stationary}, we have analytically integrated out $h$ and $x$ in $p(f\cond u, z)=\int p(f, h, x\cond u, z)\id{h}\id{x}$ prior to performing inference.
-It is indeed the case that $p(f\cond u, z)$ is intractable, but its first two moment can actually be computed, and only those we need in formulating our inference scheme.
+It is indeed the case that $p(f\cond u, z)$ is intractable, but its first two moments can actually be computed, and only those we need in formulating our inference scheme.
 Rearranging

### 2017-02-24 doc/paper.tex
Changes

@@
 \begin{abstract}
-We investigate the notion of causality in modelling stationary time series.
+% We investigate the notion of causality in modelling stationary time series.
 Extending on the work by \citet{Tobar:2015:Learning_Stationary}, we present the Causal Gaussian Process Convolution Model (CGPCM).
@@
 We develop enhanced variational inference and learning schemes and demonstrate the CGPCM on synthetic and real-world signals.
+Extending on the work by \citet{Tobar:2015:Learning_Stationary}, we present the Causal Gaussian Process Convolution Model (CGPCM), a doubly nonparametric model for causal, spectrally complex dynamical phenomena.
+The CGPCM is a generative model in which white noise is passed through a causal, nonparametric-window moving-average filter, which we show to be equivalent to a Gaussian process with a nonparametric kernel that is biased towards causally-generated signals.
+We develop enhanced variational inference and learning schemes for the causal and acausal variants of the GPCM that significantly improve statistical accuracy.
+These modelling and inferential contributions are finally demonstrated on synthetic and real-world signals.
 \end{abstract}

### 2017-02-24 doc/paper.tex
Final version

--- 2017-02-24 doc/paper.tex
@@
 \begin{abstract}
-We investigate the notion of causality in modelling stationary time series.
-Extending on the work by \citet{Tobar:2015:Learning_Stationary}, we present the Causal Gaussian Process Convolution Model (CGPCM).
-The CGPCM models the kernel of a Gaussian process nonparametrically and has an inductive bias towards causally-generated time series.
-We develop enhanced variational inference and learning schemes and demonstrate the CGPCM on synthetic and real-world signals.
+This paper presents the Causal Gaussian Process Convolution Model (CGPCM), a doubly nonparametric model for causal, spectrally complex dynamical phenomena.
+The CGPCM is a generative model in which white noise is passed through a causal, nonparametric-window moving-average filter, which we show to be equivalent to a Gaussian process with a nonparametric kernel that is biased towards causally-generated signals.
+We develop enhanced variational inference and learning schemes for the causal and previous acausal variants of the GPCM \citet{Tobar:2015:Learning_Stationary} that significantly improve statistical accuracy.
+These modelling and inferential contributions are finally demonstrated on synthetic and real-world signals.
 \end{abstract}

### 2017-02-28 src/config.py
First commit of source, still broken

@@ -1,0 +1,5 @@
+import tensorflow as tf
+
+
+reg = 1e-5
+dtype = tf.float64

### 2017-02-28 src/core/dist.py

@@ -1,0 +1,94 @@
+import numpy as np
+import tensorflow as tf
+
+from core.utils import *
+
+
+class Normal:
+    """
+    Normal distribution.
+
+    :param variance: variance
+    :param mean: mean
+    """
+    def __init__(self, variance, mean=None):
+        self.dims = shape(variance)[-1]
+        if mean is None:
+            self.mean = zeros([self.dims, 1])
+        else:
+            self.mean = mean
+        self.variance = variance
+
+    @property
+    def m2(self):
+        """
+        Get the second moment.
+
+        :return: second moment
+        """
+        return self.variance + outer(self.mean)
+
+    def log_pdf(self, x):
+        """
+        Get the log of the pdf evaluated for a feature matrix `x`.
+
+        :param x: points to evaluate at
+        :return: log pdf
+        """
+        if shape(x)[1] != self.dims:
+            raise RuntimeError('Dimensionality of data points does not match '
+                               'distribution')
+        n = shape(x)[0]  # Number of data points
+        chol = tf.cholesky(self.variance)
+        y = trisolve(chol, tf.transpose(x) - self.mean)
+        return -(log_det(chol) + n * to_float(self.dims) * log_2_pi()
+                 + tf.reduce_sum(y * y)) / 2.
+
+    def kl(self, other):
+        """
+        Compute the KL divergence with respect to another normal.
+
+        :param other: other normal
+        :return: KL divergence
+        """
+        if self.dims != other.dims:
+            raise RuntimeError('KL divergence can only be computed between '
+                               'distributions of the same dimensionality')
+        chol_self = tf.cholesky(self.variance)
+        chol_other = tf.cholesky(other.variance)
+        mu_diff_part = tf.reduce_sum(trisolve(chol_other,
+                                              other.mean - self.mean) ** 2)
+        trace_part = tf.reduce_sum(trisolve(chol_other, chol_self) ** 2)
+        return .5 * (trace_part + mu_diff_part - to_float(self.dims)
+                     + log_det(chol_other) - log_det(chol_self))
+
+    def sample(self, num=1):
+        """
+        Generate a sample.
+
+        :param num: number of samples
+        :return: sample
+        """
+        sample = randn([self.dims, num])
+        return mul(tf.cholesky(self.variance), sample) + self.mean
+
+
+class Uniform:
+    """
+    Uniform distribution.
+
+    :param lb: lower bound
+    :param ub: upper bound
+    """
[…]

### 2017-02-28 src/core/expq.py

@@ -1,0 +1,563 @@
+from core.utils import *
+import operator
+
+if __name__ == '__main__':
+    # For debugging convenience
+    sess = tf.Session()
+
+
+def const(x):
+    """
+    Constant polynomial.
+
+    :param x: constant
+    :return: polynomial
+    """
+    return Poly(Term(x))
+
+
+def var(x, power=1):
+    """
+    Polynomial consisting of just a single variable.
+
+    :param x: variable name
+    :param power: power
+    :return: polynomial
+    """
+    return Poly(Term(1., Factor(x, power)))
+
+
+def kh(alpha, gamma, x, y):
+    """
+    Kernel of :math:`h` process.
+
+    :param alpha: :math:`\\alpha`
+    :param gamma: :math:`\\gamma`
+    :param x: first variable
+    :param y: second variable
+    :return: exponentiated quadratic
+    """
+    return ExpQ(-const(alpha) * (x ** 2 + y ** 2)
+                - const(gamma) * (x - y) ** 2)
+
+
+def kxs(omega, t1, t2):
+    """
+    Kernel between :math:`s` and :math:`x` processes.
+
+    :param omega: :math:`\\omega`
+    :param x: first variable
+    :param y: second variable
+    :return: exponentiated quadratic
+    """
+    return ExpQ(-const(omega) * (t1 - t2) ** 2)
+
+
+class Factor:
+    """
+    Variable raised to some power.
+
+    :param var: variable name
+    :param power: power
+    """
+
+    def __init__(self, var, power):
+        self._var = var
+        self._power = power
+
+    @property
+    def var(self):
+        return self._var
+
+    @property
+    def power(self):
+        return self._power
+
+    def eval(self, **var_map):
+        """
+        Evaluate factor.
+
+        :param var_map: variable mapping
+        :return: evaluated variable
+        """
+        return var_map[self._var] ** self._power
+
+    def __eq__(self, other):
+        return self._var == other.var and self._power == other.power
+
+    def __str__(self):
+        return '{}^{}'.format(self._var, self._power)
+
+    def __mul__(self, other):
+        if is_numeric(other) and other == 1:
+            return self
+        if not other.var == self._var:
+            raise ValueError('other Factor must be function of same variable')
+        return Factor(self.var, self.power + other.power)
+
+    def __rmul__(self, other):
+        return self.__mul__(other)
+
+    def __hash__(self):
+        return hash(self._var) + hash(self._power)
+
+
+class Term:
+    """
[…]

### 2017-02-28 src/core/gpcm.py

@@ -1,0 +1,481 @@
+import scipy.stats
+
+from core import expq as expq
+from core.dist import Normal
+from core.kernel import DEQ
+from core.samplers import ESS
+import core.utils as utils
+from core.utils import *
+
+
+class GPCM:
+    """
+    Gaussian Process Convolution Model.
+    """
+
+    def __init__(self, pars, th, tx, t, y, sess):
+        self.th, self.tx = th, tx
+        self.nh, self.nx = shape(th)[0], shape(tx)[0]
+        self.pars = pars
+        self.t = t
+        self.y = y
+        self.sess = sess
+        self._init_kernels()
+
+        self.var_map = {'th1': expand_dims(th, 2, 3),
+                        'th2': expand_dims(th, 3, 2),
+                        'tx1': expand_dims(tx, 4, 1),
+                        'tx2': expand_dims(tx, 5, 0)}
+        self.kh = lambda t1, t2: expq.kh(pars['alpha'], pars['gamma'], t1, t2)
+        self.kxs = lambda t1, t2: expq.kxs(pars['omega'], t1, t2)
+
+        # Some variables
+        self.tau1 = expq.var('tau1')
+        self.tau2 = expq.var('tau2')
+        self.tau = expq.var('tau')
+        self.t1 = expq.var('t1')
+        self.t2 = expq.var('t2')
+        self.th1 = expq.var('th1')
+        self.th2 = expq.var('th2')
+        self.tx1 = expq.var('tx1')
+        self.tx2 = expq.var('tx2')
+        self.min_t1_t2 = expq.var('min_t1_t2')
+        self.min_t1_0 = expq.var('min_t1_0')
+        self.min_t1_tx1 = expq.var('min_t1_tx1')
+        self.min_t1_tx2 = expq.var('min_t1_tx2')
+
+        # Some expressions
+        self.expq_a = self.kh(self.t1 - self.tau, self.t2 - self.tau)
+        self.expq_Ahh = self.kh(self.t1 - self.tau, self.th1) \
+                        * self.kh(self.th2, self.t2 - self.tau)
+        self.expq_Axx = self.kh(self.t1 - self.tau1, self.t2 - self.tau2) \
+                        * self.kxs(self.tau1, self.tx1) \
+                        * self.kxs(self.tx2, self.tau2)
+        self.expq_Ahx = self.kh(self.t1 - self.tau, self.th1) \
+                        * self.kxs(self.tau, self.tx1)
+
[…]

### 2017-02-28 src/core/kernel.py

@@ -1,0 +1,75 @@
+import abc
+from utils import *
+
+
+class Kernel:
+    __metaclass__ = abc.ABCMeta
+
+    def __init__(self, hyps):
+        self._hyps = hyps
+
+    def __getitem__(self, hyp):
+        return self._hyps[hyp]
+
+    @abc.abstractmethod
+    def __call__(self, points):
+        """
+        Evaluate the kernel for a matrix of points.
+
+        :param points: points to evaluate kernel at
+        :returns: kernel matrix
+        """
+        pass
+
+
+class DEQ(Kernel):
+    """
+    Decaying exponentiated-quadratic kernel.
+
+    :param hyps: dictionary containing the following hyperparameters:
+
+                    - `s2`: :math:`\\sigma^2`,
+                    - `alpha`: :math:`\\alpha`, and
+                    - `gamma`: :math:`\\gamma`
+    """
+    def __call__(self, x, y=None):
+        if y is None:
+            y = x
+        dists2, norms2_x, norms2_y = pw_dists2(x, y)
+        return self['s2'] * tf.exp(-self['alpha'] * (norms2_x + norms2_y)
+                                   - self['gamma'] * dists2)
+
+
+class UhlenbeckOrnstein(Kernel):
+    """
+    Uhlenbeck-Ornstein kernel.
+
+    :param hyps: dictionary containing the following hyperparameters:
+
+                    - `s2`: :math:`\\sigma^2`,
+                    - `gamma`: :math:`\\gamma`
+    """
+    def __call__(self, x, y=None):
+        if y is None:
+            y = x
+        dists2, norms2_x, norms2_y = pw_dists2(x, y)
+        return self['s2'] * tf.exp(-self['gamma'] * dists2 ** .5)
+
+
+class RLCSeriesCircuit(Kernel):
+    """
+    Kernel corresponding to an RLC series circuit excited by white noise due to
+    thermal noise of the resistance.
+
+    :param hyps: dictionary containing the following hyperparameters:
+
+                    - `s2`: :math:`\\sigma^2`,
+                    - `gamma`: :math:`\\gamma`
+    """
+    def __call__(self, x, y=None):
+        if y is None:
+            y = x
+        dists2, norms2_x, norms2_y = pw_dists2(x, y)
+        dists = self['gamma'] * dists2 ** .5
+        return self['s2'] * (1 + dists) * tf.exp(-dists)
+

### 2017-02-28 src/core/plotter.py

@@ -1,0 +1,324 @@
+from mpl_toolkits.mplot3d import Axes3D
+import matplotlib as mpl
+import abc
+import warnings
+import itertools
+with warnings.catch_warnings():
+    warnings.simplefilter('ignore')
+    import matplotlib.pyplot as plt
+
+
+def to_rbga(color):
+    """ Convert `color` to a tuple `(red, blue, green, alpha)`. """
+    return mpl.colors.colorConverter.to_rgba(color, alpha=1.0)
+
+
+# Mapping for values from strings
+val_map = {'True': True, 'False': False}
+
+
+def dict_map(source, mapping):
+    """
+    Replace values in dictionary `mapping` with their values in `source`
+    if possible and return result as a new dictionary. If a value is of the
+    form `val:x`, then this value will be set to `x` and parsed
+    appropriately.
+
+    :param source: source dictionary
+    :param mapping: mapping
+    :return: mapping dictionary
+    """
+    result = {}
+    for key, val in mapping.items():
+        if len(val) > 5 and val[:4] == 'val:':
+            val_parsed = val[4:]
+            if val_parsed in val_map:
+                result[key] = val_map[val_parsed]
+            else:
+                result[key] = val_parsed
+        elif val in source and source[val] != 'undefined':
+            result[key] = source[val]
+    return result
+
+
+class Plotter:
+    """
+    Sane and consistent configuration of plots.
+
+    :param \*\*kw_args: configuration adjustments
+    """
+    _config = {'axes_color': 'black',
+               'axes_width': 1,
+               'axes_labelpad': 8,
+               'grid_color': '0.8',
+               'grid_width': .5,
+               'grid_style': '--',
+               'font_family': 'Adobe Caslon Pro',
+               'font_size': 11,
+               'fig_size': (8, 6),
+               'legend_color': '0.95',
+               'line_color': 'undefined',
+               'marker_style': 'None',
+               'line_width': 1,
+               'line_style': '-',
+               'colorbar_shrink': 0.8,
+               'colorbar_aspect': 20,
+               'marker_color': 'undefined',
+               'marker_size': 5,
+               'surface_rstride': 1,
+               'surface_cstride': 1,
+               'surface_antialiased': False,
+               'surface_line_width': 0,
+               'cmap': mpl.cm.coolwarm,
+               'fill_alpha': 0.25,
+               'fill_color': 'undefined',
+               'label': 'undefined'
+               }
+
[…]

### 2017-02-28 src/core/samplers.py

@@ -1,0 +1,130 @@
+import numpy.random as rng
+import numpy as np
+from time import time
+
+
+class ESS(object):
+    """
+    Elliptical slice sampling algorithm.
+
+    :param log_lik: function that returns the log value of a quantity that is
+                    proportional to the likelihood of the target distribution
+    :param sample_prior: function that generates a sample from the prior of
+                         the target distribution
+    :param x_init: initial state
+    """
+
+    def __init__(self, log_lik, sample_prior, x_init=None):
+        self._log_lik = log_lik
+        self._sample_prior = sample_prior
+        if x_init is None:
+            self._x = sample_prior()
+        else:
+            self._x = x_init
+        self._log_lik_x = log_lik(self._x)
+
+    def move(self, x, log_lik=None):
+        """
+        Move sampler to a new state.
+
+        :param x: new state
+        :param log_lik: log likelihood at new state
+        """
+        self._x = x
+        if log_lik is None:
+            self._log_lik_x = self._log_lik(x)
+
+    def update(self, log_lik):
+        """
+        Update the log likelihood function.
+
+        :param log_lik: log likelihood
+        """
+        self._log_lik = log_lik
+
+    def _establish_ellipse(self):
+        """
+        Establish an ellipse, which is required to draw the new state.
+        """
+        self._y = self._sample_prior()
+
+    def _draw_proposal(self, theta_l, theta_u):
+        """
+        Draw a proposal for the next state given a bracket for :math:`\\theta`.
+
+        :param theta_l: lower bound of bracket
+        :param theta_u: upper bound of bracket
+        :return: proposal
+        """
+        self._theta = rng.uniform(theta_l, theta_u)
+        self._x_proposed = np.cos(self._theta) * self._x \
+                           + np.sin(self._theta) * self._y
+        self._log_lik_x_proposed = self._log_lik(self._x_proposed)
+
+    def _draw_bracket(self):
+        """
+        Draw a bracket for :math:`\\theta`.
+
+        :return: bracket
+        """
+        theta = rng.uniform(0, 2 * np.pi)
+        return theta - 2 * np.pi, theta
+
+    def _draw(self, (theta_l, theta_u), u, attempts=1):
+        """
+        Draw new state given a bracket for :math:`\\theta`.
+
+        :param theta_l: lower bound of bracket
+        :param theta_u: upper bound of bracket
+        :param u: slice height
+        :param attempts: number of attempts
+        :return: new state
+        """
+        self._draw_proposal(theta_l, theta_u)
[…]

### 2017-02-28 src/core/utils.py

@@ -1,0 +1,640 @@
+import tensorflow as tf
+import numpy as np
+import colorsys
+import scipy.stats
+import tensorflow.contrib.distributions as tf_dists
+from tensorflow.python.client import timeline
+from tensorflow.python.framework import ops
+
+from rpy2.robjects import r
+from rpy2.robjects.packages import importr
+import rpy2.robjects.numpy2ri
+
+import config
+
+rpy2.robjects.numpy2ri.activate()
+importr("pbivnorm")
+
+lower_perc = 100 * scipy.stats.norm.cdf(-2)
+upper_perc = 100 * scipy.stats.norm.cdf(2)
+
+inf = np.inf
+
+
+def pw_dists2(x, y):
+    """
+    Compute pairwise distances between the rows of `x` and `y`.
+
+    :param x: first set of features
+    :param y: second set of features
+    :return: squared distances, norms of `x`, and norms of `y`
+    """
+    norms2_x = tf.reduce_sum(x ** 2, reduction_indices=[1])[:, None]
+    norms2_y = tf.reduce_sum(y ** 2, reduction_indices=[1])[None, :]
+    return norms2_x - 2 * mul(x, y, adj_b=True) + norms2_y, norms2_x, norms2_y
+
+
+def transp(x):
+    """
+    Batch transpose a tensor.
+
+    :param x: tensor
+    :return: tensor where the last two dimensions are transposed
+    """
+    perm = range(len(shape(x)))
+    perm[-2], perm[-1] = perm[-1], perm[-2]
+    return tf.transpose(x, perm=perm)
+
+
+def trace(x):
+    """
+    Batch trace a tensor.
+
+    :param x: tensor
+    :return: traces of the last two dimensions of the tensor
+    """
+    return tf.reduce_sum(tf.matrix_diag_part(x), [-1])
+
+
+def eye(n):
+    """
+    :math:`N`-dimensional identity matrix.
+
+    :param n: number of dimensions
+    :return: identity matrix
+    """
+    return tf.diag(tf.ones([n], dtype=config.dtype))
+
+
+def invert_perm(perm):
+    """
+    Invert a permutation.
+
+    :param perm: permutation
+    :return: inverse of permutation
+    """
+    n = len(perm)
+    inverse_perm = range(n)
+    for i in range(n):
+        inverse_perm[perm[i]] = i
+    return inverse_perm
+
+
+def randn(size):
+    """
+    Alias for `tf.random_normal`.
+    """
[…]

### 2017-02-28 src/doc.sh

@@ -1,0 +1,15 @@
+#!/usr/bin/env bash
+
+function title {
+    echo ">> $1"
+}
+
+title "Generating documentation source"
+python generate_doc_source.py
+title "Building documentation"
+# latexpdf
+make clean html
+
+if [ "$1" = "open" ]; then
+    open build/html/index.html
+fi

### 2017-02-28 src/generate_doc_source.py

@@ -1,0 +1,166 @@
+import os
+import subprocess as sp
+import sqlite3
+import datetime
+
+
+documentation = [
+    {'package': 'core',
+     'title': 'Core Modules',
+     'description': 'Important modules that make up the system',
+     'subpackages': []},
+     {'package': 'learn',
+      'title': 'Learning Tasks',
+      'description': 'Scripts to perform learning tasks',
+      'subpackages': []}
+]
+
+
+def walk_dir(dir_path):
+    """
+    Get all the files in a directory.
+
+    :param dir_path: path of directory
+    :return: files in directory
+    """
+    files = []
+    for file in os.listdir(dir_path):
+        if file.startswith('.'):
+            continue
+        file_path = os.path.join(dir_path, file)
+        if os.path.isfile(file_path):
+            files += [file_path]
+    return files
+
+
+def filter_files(files, exts):
+    """
+    Filter files according to a list of extensions.
+
+    :param files: list of files
+    :param exts: accepted extensions
+    :return: list of accepted files
+    """
+    return filter(lambda f: os.path.splitext(f)[1] in exts, files)
+
+
+line1 = '=' * 100
+line2 = '-' * 100
+line3 = '^' * 100
+
+template_index = '''
+GGPCM Documentation
+{line1:s}
+
+.. toctree::
+   :maxdepth: 2
+
+   {modules:s}
+
+
+Indices and Tables
+{line1:s}
+
+* :ref:`genindex`
+* :ref:`modindex`
+* :ref:`search`
+'''
+
+template_module = '''
+{module:s}
+{line2:s}
+.. automodule:: {module:s}
+'''
+
+template_package = '''
+{title:s}
+{line1:s}
+{description:s}
+{modules:s}
+'''
+
+template_experiments = '''
+Experiments on Branch ``{branch:s}``
+{line1:s}
+{experiments:s}
+'''
+
+template_experiment = '''
+{label:s} ({timestamp:s})
+{line2:s}
+Reason
+{line3:s}
+{reason:s}
+
+Outcome
+{line3:s}
+{outcome:s}
+'''
+
+
+
+class Documentation:
+    """
+    Generator of documentation source.
+    """
+
+    def __init__(self, source_dir='source'):
+        self._source_dir = source_dir
+
+    def generate(self):
+        """
+        Generate documentation source.
+        """
+        self._write_index()
+        self._write_packages()
+
+    def _write_packages(self):
+        for package in documentation:
+            with open(os.path.join(self._source_dir,
[…]

### 2017-02-28 src/learn/hrtf.py

@@ -1,0 +1,309 @@
+from tensorflow.contrib.opt.python.training.external_optimizer import \
+    ScipyOptimizerInterface as SciPyOpt
+import pickle
+import scipy.io.wavfile as wav
+import scipy.io as sio
+
+from core.gpcm import VGPCM
+from core.plotter import Plotter2D
+from core.utils import *
+
+
+def normalise_energy(x, dt, causal, i_mid=None):
+    single = False
+    if not (type(x) == list or type(x) == tuple):
+        x = [x]
+        single = True
+    if causal:
+        if i_mid is None:
+            i_mid = (len(x[0]) - 1) / 2
+        energy = np.sum(x[0][i_mid:] ** 2 * dt)
+    else:
+        energy = np.sum(x[0] ** 2 * dt)
+    return x[0] / energy ** .5 if single \
+        else map(lambda y: y / energy ** .5, x)
+
+
+def normalise(x):
+    if type(x) == list or type(x) == tuple:
+        return map(lambda y: y / max(x[0]), x)
+    else:
+        return x / max(x)
+
+
+sess = SessionDecoratorDebug(tf.Session())
+
+seed = 21474836
+tf.set_random_seed(seed)
+np.random.seed(seed)
+
+cgpcm = True
+do_train = True
+do_plot = True
+
+if do_train:
+    print 'Loading data...'
+    n = 500  # Data under consideration
+    choose = 500  # Number of training points
+    # imp_region = (100, 100)
+    nh = 101
+    nx = 150
+    nk = 501
+    k_len = .3
+    noise_init = 0.1 ** 2
+    gamma_init = length_scale(.05 * k_len)
+    omega_init = length_scale(2. / nx)
+    k_stretch = 2
+
+    train_iters_pre = 500
+    train_iters = 0
+    ess_burn = 200
+    ess_samps = 200
+
+    noise = 0
+
+    causal = cgpcm
+
+    pars = {'alpha': length_scale(k_len),
+            'gamma': var_pos(to_float(gamma_init), 'gamma'),
+            'omega': var_pos(to_float(omega_init), 'omega'),
+            'K': 1, 'M': 1, 'N': 1,
+            's2': var_pos(to_float(noise_init), 's2'),
+            's2_f': var_pos(to_float(1.), 's2_f')}
+    initialise_uninitialised_variables(sess)
+
+    t = np.linspace(-1, 1, n)[:, None]
[…]

### 2017-02-28 src/learn/timit.py

@@ -1,0 +1,280 @@
+from tensorflow.contrib.opt.python.training.external_optimizer import \
+    ScipyOptimizerInterface as SciPyOpt
+import pickle
+import scipy.io as sio
+
+from core.gpcm import VGPCM
+from core.plotter import Plotter2D
+from core.utils import *
+
+
+def normalise_energy(x, dt, causal):
+    single = False
+    if not (type(x) == list or type(x) == tuple):
+        x = [x]
+        single = True
+    if causal:
+        energy = np.sum(x[0][(len(x[0]) - 1) / 2:] ** 2 * dt)
+    else:
+        energy = np.sum(x[0] ** 2 * dt)
+    return x[0] / energy ** .5 if single \
+        else map(lambda y: y / energy ** .5, x)
+
+
+def normalise(x):
+    if type(x) == list or type(x) == tuple:
+        return map(lambda y: y / max(x[0]), x)
+    else:
+        return x / max(x)
+
+
+sess = SessionDecoratorDebug(tf.Session())
+
+seed = 21474836
+tf.set_random_seed(seed)
+np.random.seed(seed)
+
+mod_name = 'cgpcm---fit15'
+
+cgpcm = True
+do_train = True
+do_plot = True
+
+if do_train:
+    print 'Loading data...'
+    m = 800  # Chunk size
+    n = 500  # Subsampling of chunk
+    choose = 250  # Further subsampling
+    nh = 120
+    nx = 120
+    nk = 501
+    k_len = .6
+    noise_init = 0.1 ** 2
+    gamma_init = length_scale(.05 * k_len)
+    omega_init = length_scale(2. / nx)
+    k_stretch = 2
+
+    train_iters_pre = 500
+    train_iters = 0
+    ess_burn = 200
+    ess_samps = 200
+
+    noise = 0
+
+    causal = cgpcm
+
+    pars = {'alpha': length_scale(k_len),
+            'gamma': var_pos(to_float(gamma_init), 'gamma'),
+            'omega': var_pos(to_float(omega_init), 'omega'),
+            'K': 1, 'M': 1, 'N': 1,
+            's2': var_pos(to_float(noise_init), 's2'),
+            's2_f': var_pos(to_float(1.), 's2_f')}
+    initialise_uninitialised_variables(sess)
+
+    t = np.linspace(-1, 1, n)[:, None]
+    th = np.linspace(-k_stretch * k_len, k_stretch * k_len, nh)[:, None]
+    tx = np.linspace(-1, 1, nx)[:, None]
[…]

### 2017-02-28 src/learn/timit2.py

@@ -1,0 +1,306 @@
+from tensorflow.contrib.opt.python.training.external_optimizer import \
+    ScipyOptimizerInterface as SciPyOpt
+import pickle
+import scipy.io as sio
+
+from core.gpcm import VGPCM
+from core.plotter import Plotter2D
+from core.utils import *
+
+
+def normalise_energy(x, dt, causal):
+    single = False
+    if not (type(x) == list or type(x) == tuple):
+        x = [x]
+        single = True
+    if causal:
+        energy = np.sum(x[0][(len(x[0]) - 1) / 2:] ** 2 * dt)
+    else:
+        energy = np.sum(x[0] ** 2 * dt)
+    return x[0] / energy ** .5 if single \
+        else map(lambda y: y / energy ** .5, x)
+
+
+def normalise(x):
+    if type(x) == list or type(x) == tuple:
+        return map(lambda y: y / max(x[0]), x)
+    else:
+        return x / max(x)
+
+
+sess = SessionDecoratorDebug(tf.Session())
+
+seed = 21474836
+tf.set_random_seed(seed)
+np.random.seed(seed)
+
+cgpcm = True
+do_train = True
+do_plot = True
+
+if do_train:
+    print 'Loading data...'
+    data_size = 800  # Fragment length
+    n = 300  # Data under consideration
+    choose = 200  # Number of training points
+    imp_region = (150, 200)
+    nh = 80
+    nx = 120
+    nk = 501
+    k_len = .4
+    noise_init = 0.001 ** 2
+    gamma_init = length_scale(.4 * k_len)
+    omega_init = length_scale(2. / nx)
+    k_stretch = 2
+
+    train_iters_pre = 100
+    train_iters = 0
+    ess_burn = 100
+    ess_samps = 100
+
+    noise = 0
+
+    causal = cgpcm
+
+    pars = {'alpha': length_scale(k_len),
+            'gamma': var_pos(to_float(gamma_init), 'gamma'),
+            'omega': var_pos(to_float(omega_init), 'omega'),
+            'K': 1, 'M': 1, 'N': 1,
+            's2': var_pos(to_float(noise_init), 's2'),
+            's2_f': var_pos(to_float(1.), 's2_f')}
+    initialise_uninitialised_variables(sess)
+
+    t = np.linspace(-1, 1, n)[:, None]
+    th = np.linspace(-k_stretch * k_len, k_stretch * k_len, nh)[:, None]
[…]

### 2017-02-28 src/learn/toy.py

@@ -1,0 +1,336 @@
+from tensorflow.contrib.opt.python.training.external_optimizer import \
+    ScipyOptimizerInterface as SciPyOpt
+import pickle
+
+from core.gpcm import AKM, VGPCM
+from core.plotter import Plotter2D
+from core.utils import *
+
+
+def normalise_energy(x, dt, causal):
+    single = False
+    if not (type(x) == list or type(x) == tuple):
+        x = [x]
+        single = True
+    if causal:
+        energy = np.sum(x[0][(len(x[0]) - 1) / 2:] ** 2 * dt)
+    else:
+        energy = np.sum(x[0] ** 2 * dt)
+    return x[0] / energy ** .5 if single \
+        else map(lambda y: y / energy ** .5, x)
+
+
+def normalise(x):
+    if type(x) == list or type(x) == tuple:
+        return map(lambda y: y / max(x[0]), x)
+    else:
+        return x / max(x)
+
+
+sess = SessionDecoratorDebug(tf.Session())
+
+# Causal sample:
+# Seed: 214748361 + 118
+# Noise: 0
+# Noise init: 0.01 ** 2
+# Redraw noise: 3
+# Sample alpha: kernel_length
+# Sample gamma: .8 * kernel_length
+
+# Acausal sample:
+# Seed: 21474836 + 202
+# Noise: 0.15 ** 2
+# Noise init: 0.15 ** 2
+# Redraw noise: 0
+# Sample alpha: .8 * kernel_length
+# Sample gamma: .8 * kernel_length
+
+seed = 21474836 + 202
+tf.set_random_seed(seed)
+np.random.seed(seed)
+
+do_train = True
+do_train_fn = 'approx_q'
+do_plot = True
+do_plot_fn = 'known_kernels_acs_acm'
+
+if do_train:
+    print 'Loading data...'
+    n = 300
+    choose = 300
+    nh = 51
+    nx = 80
+    nk = 501
+    k_len = .3
+    noise_init = 0.15 ** 2
+    gamma_init = length_scale(.6 * k_len)
+    omega_init = length_scale(2. / nx)
+    k_stretch = 3
+
+    kernel_len = .5 * k_len
+    noise = 0.15 ** 2
+
+    train_iters_pre = 1000
+    train_iters = 0
+    ess_burn = 500
+    ess_samps = 200
+
+    test_kernel = False
+    test_sample = False
+    redraw_noise = 0
+    init_at_optimal_mean = False
+
[…]

### 2017-02-28 src/source/conf.py

@@ -1,0 +1,367 @@
+# -*- coding: utf-8 -*-
+#
+# Price Impact documentation build configuration file, created by
+# sphinx-quickstart on Tue Oct  4 10:40:50 2016.
+#
+# This file is execfile()d with the current directory set to its
+# containing dir.
+#
+# Note that not all possible configuration values are present in this
+# autogenerated file.
+#
+# All configuration values have a default; values that are commented out
+# serve to show the default.
+
+# If extensions (or modules to document with autodoc) are in another directory,
+# add these directories to sys.path here. If the directory is relative to the
+# documentation root, use os.path.abspath to make it absolute, like shown here.
+
+import os
+import sys
+sys.path.insert(0, os.path.abspath('..'))
+
+# -- General configuration ------------------------------------------------
+
+# If your documentation needs a minimal Sphinx version, state it here.
+#
+# needs_sphinx = '1.0'
+
+# Add any Sphinx extension module names here, as strings. They can be
+# extensions coming with Sphinx (named 'sphinx.ext.*') or your custom
+# ones.
+extensions = [
+    'sphinx.ext.autodoc',
+    'sphinx.ext.doctest',
+    'sphinx.ext.todo',
+    'sphinx.ext.coverage',
+    'sphinx.ext.mathjax',
+    'sphinx.ext.viewcode',
+]
+
+# Add any paths that contain templates here, relative to this directory.
+templates_path = ['_templates']
+
+# The suffix(es) of source filenames.
+# You can specify multiple suffix as a list of string:
+#
+# source_suffix = ['.rst', '.md']
+source_suffix = '.rst'
+
+# The encoding of source files.
+#
+# source_encoding = 'utf-8-sig'
+
+# The master toctree document.
+master_doc = 'index'
+
+# General information about the project.
+project = u'GGPCM'
+copyright = u'Wessel Bruinsma'
+author = u'Wessel Bruinsma'
+
+# The version info for the project you're documenting, acts as replacement for
+# |version| and |release|, also used in various other places throughout the
+# built documents.
+#
+# The short X.Y version.
+version = u'0.1'
+# The full version, including alpha/beta/rc tags.
+release = u'0.1'
+
+# The language for content autogenerated by Sphinx. Refer to documentation
+# for a list of supported languages.
+#
+# This is also used if you do content translation via gettext catalogs.
+# Usually you set "language" from the command line for these cases.
+language = None
+
+# There are two options for replacing |today|: either, you set today to some
[…]

### 2017-02-28 src/test_bvn_cdf_benchmark.py

@@ -1,0 +1,30 @@
+from core.utils import _mvn_cdf, _bvn_cdf
+import numpy as np
+from time import time
+
+
+def bench(fn, name, iters=100):
+    print 'Benchmark: {}'.format(name)
+    start = time()
+    for i in range(iters):
+        fn()
+    dur = time() - start
+    print '  Duration:           {:.3f} s'.format(dur)
+    print '  Duration/iteration: {:.3f} s'.format(dur / iters)
+
+
+num = 50000
+x = np.random.rand(num, 2)
+mu = np.array([[0.2, 0.4]])
+var = np.array([[3., 0.1], [0.1, 4.]])
+
+print 'Warming up...'
+_mvn_cdf(np.zeros(2), np.zeros(2), np.eye(2))
+
+bench(lambda: _mvn_cdf(x, mu, var), '_mvn_cdf')
+bench(lambda: _bvn_cdf(x, mu, var), '_bvn_cdf')
+
+
+
+
+

### 2017-02-28 src/test_bvn_cdf_gradient.py

@@ -1,0 +1,54 @@
+from core.utils import *
+
+
+with tf.Session() as sess:
+    # Some complicated expression to test more advanced gradients
+    v = ones(3)
+    fac = randn([3]).eval()
+    div = to_float(5)
+    rho = sum(v * fac) / div
+
+    x = tf.constant(np.array([[0.1, 0.2], [0.2, 0.3]]))
+    y = bvn_cdf2(x, rho)
+
+    delta = 1e-8
+
+    print 'y1 wrt x11'
+    e = tf.constant(np.array([[1., 0], [0, 0]])) * delta / 2.
+    y1 = bvn_cdf2(x - e, rho)
+    y2 = bvn_cdf2(x + e, rho)
+    print (y2[0].eval() - y1[0].eval()) / delta
+    print tf.gradients([y[0]], x)[0].eval()[0, 0]
+
+    print 'y2 wrt x21'
+    e = tf.constant(np.array([[0, 0], [1., 0]])) * delta / 2.
+    y1 = bvn_cdf2(x - e, rho)
+    y2 = bvn_cdf2(x + e, rho)
+    print (y2[1].eval() - y1[1].eval()) / delta
+    print tf.gradients([y[1]], x)[0].eval()[1, 0]
+
+    print 'y1 wrt rho'
+    e = to_float(1) * delta / 2
+    y1 = bvn_cdf2(x, rho - e)
+    y2 = bvn_cdf2(x, rho + e)
+    print (y2[0].eval() - y1[0].eval()) / delta
+    print tf.gradients([y[0]], rho)[0].eval()
+
+    print 'y2 wrt v1'
+    e = tf.constant(np.array([1., 0, 0])) * delta / 2
+    rho1 = sum((v - e) * fac) / div
+    rho2 = sum((v + e) * fac) / div
+    y1 = bvn_cdf2(x, rho1)
+    y2 = bvn_cdf2(x, rho2)
+    print (y2[1].eval() - y1[1].eval()) / delta
+    print tf.gradients([y[1]], v)[0].eval()[0]
+
+    print 'y2 wrt v3'
+    e = tf.constant(np.array([0, 0, 1.])) * delta / 2
+    rho1 = sum((v - e) * fac) / div
+    rho2 = sum((v + e) * fac) / div
+    y1 = bvn_cdf2(x, rho1)
+    y2 = bvn_cdf2(x, rho2)
+    print (y2[1].eval() - y1[1].eval()) / delta
+    print tf.gradients([y[1]], v)[0].eval()[2]
+

### 2017-03-01 src/core/dist.py
Refactoring of core code

@@ -1,16 +1,14 @@
-import numpy as np
-import tensorflow as tf
-
-from core.utils import *
+from core.tfutil import *
 
 
-class Normal:
+class Normal(object):
     """
     Normal distribution.
 
     :param variance: variance
     :param mean: mean
     """
+
     def __init__(self, variance, mean=None):
         self.dims = shape(variance)[-1]
         if mean is None:
@@ -73,13 +71,14 @@ class Normal:
         return mul(tf.cholesky(self.variance), sample) + self.mean
 
 
-class Uniform:
+class Uniform(object):
     """
     Uniform distribution.
 
     :param lb: lower bound
     :param ub: upper bound
     """
+
     def __init__(self, lb, ub):
         self.lb = lb
         self.ub = ub
@@ -91,4 +90,3 @@ class Uniform:
         :return: sample
         """
         return self.lb + (self.ub - self.lb) * tf.random_uniform([])
-

### 2017-03-01 src/core/eq.py

@@ -1,4 +1,5 @@
-from core.utils import *
+from core.tfutil import *
+from core.util import *
 import operator
 
 if __name__ == '__main__':
@@ -27,33 +28,60 @@ def var(x, power=1):
     return Poly(Term(1., Factor(x, power)))
 
 
+def kh_constructor(alpha, gamma):
+    """
+    Create a function that calls `kh` for some given hyperparameters.
+    :param alpha: kernel decay parameter :math:`\\alpha`
+    :param gamma: kernel length scale parameter :math:`\\gamma`
+    :return: constructor
+    """
+
+    def constructor(x, y):
+        return kh(alpha, gamma, x, y)
+
+    return constructor
+
+
 def kh(alpha, gamma, x, y):
     """
     Kernel of :math:`h` process.
 
-    :param alpha: :math:`\\alpha`
-    :param gamma: :math:`\\gamma`
+    :param alpha: kernel decay parameter :math:`\\alpha`
+    :param gamma: kernel length scale parameter :math:`\\gamma`
     :param x: first variable
     :param y: second variable
     :return: exponentiated quadratic
     """
-    return ExpQ(-const(alpha) * (x ** 2 + y ** 2)
-                - const(gamma) * (x - y) ** 2)
+    return EQ(-const(alpha) * (x ** 2 + y ** 2)
+              - const(gamma) * (x - y) ** 2)
 
 
-def kxs(omega, t1, t2):
+def kxs_constructor(omega):
+    """
+    Create a function that calls `kxs` for some given hyperparameter.
+    :param omega: kernel length scale parameter :math:`\\omega`
+    :return: constructor
+    """
+
+    def constructor(x, y):
+        return kxs(omega, x, y)
+
+    return constructor
+
+
+def kxs(omega, x, y):
     """
     Kernel between :math:`s` and :math:`x` processes.
 
-    :param omega: :math:`\\omega`
+    :param omega: kernel length scale parameter :math:`\\omega`
     :param x: first variable
     :param y: second variable
     :return: exponentiated quadratic
     """
-    return ExpQ(-const(omega) * (t1 - t2) ** 2)
+    return EQ(-const(omega) * (x - y) ** 2)
 
 
-class Factor:
+class Factor(object):
     """
     Variable raised to some power.
 
@@ -77,7 +105,7 @@ class Factor:
         """
         Evaluate factor.
 
-        :param var_map: variable mapping
+        :param \*\*var_map: variable mapping
[…]

### 2017-03-01 src/core/gpcm.py

@@ -1,60 +1,57 @@
-import scipy.stats
+import util
+import eq
+from par import Parametrisable
+from dist import Normal
+from kernel import DEQ
+from sample import ESS
+from tfutil import *
+from util import *
 
-from core import expq as expq
-from core.dist import Normal
-from core.kernel import DEQ
-from core.samplers import ESS
-import core.utils as utils
-from core.utils import *
 
-
-class GPCM:
+class GPCM(Parametrisable):
     """
-    Gaussian Process Convolution Model.
+    Causal Gaussian Process Convolution Model.
     """
 
-    def __init__(self, pars, th, tx, t, y, sess):
-        self.th, self.tx = th, tx
-        self.nh, self.nx = shape(th)[0], shape(tx)[0]
-        self.pars = pars
-        self.t = t
-        self.y = y
-        self.sess = sess
+    _required_pars = ['t', 'y', 'sess',
+                      'th', 'tx',
+                      's2', 's2_f', 'alpha', 'gamma', 'omega',
+                      'causal', 'causal_id']
+
+    def __init__(self, **kw_args):
+        Parametrisable.__init__(self, **kw_args)
+        self.nh, self.nx = shape(self.th)[0], shape(self.tx)[0]
+        self._precomputed = False
         self._init_kernels()
+        self._init_var_map()
+        self._init_vars()
+        self._init_expressions()
 
-        self.var_map = {'th1': expand_dims(th, 2, 3),
-                        'th2': expand_dims(th, 3, 2),
-                        'tx1': expand_dims(tx, 4, 1),
-                        'tx2': expand_dims(tx, 5, 0)}
-        self.kh = lambda t1, t2: expq.kh(pars['alpha'], pars['gamma'], t1, t2)
-        self.kxs = lambda t1, t2: expq.kxs(pars['omega'], t1, t2)
+    def _init_expressions(self):
+        kh = eq.kh_constructor(self.alpha, self.gamma)
+        kxs = eq.kxs_constructor(self.omega)
+        self.expq_a = kh(self.t1 - self.tau, self.t2 - self.tau)
+        self.expq_Ahh = kh(self.t1 - self.tau, self.th1) \
+                        * kh(self.th2, self.t2 - self.tau)
+        self.expq_Axx = kh(self.t1 - self.tau1, self.t2 - self.tau2) \
+                        * kxs(self.tau1, self.tx1) \
+                        * kxs(self.tx2, self.tau2)
+        self.expq_Ahx = kh(self.t1 - self.tau, self.th1) \
[…]

### 2017-03-01 src/core/kernel.py

@@ -1,15 +1,10 @@
 import abc
-from utils import *
+from tfutil import *
+from par import Parametrisable
 
 
-class Kernel:
+class Kernel(Parametrisable):
     __metaclass__ = abc.ABCMeta
-
-    def __init__(self, hyps):
-        self._hyps = hyps
-
-    def __getitem__(self, hyp):
-        return self._hyps[hyp]
 
     @abc.abstractmethod
     def __call__(self, points):
@@ -26,50 +21,33 @@ class DEQ(Kernel):
     """
     Decaying exponentiated-quadratic kernel.
 
-    :param hyps: dictionary containing the following hyperparameters:
-
-                    - `s2`: :math:`\\sigma^2`,
-                    - `alpha`: :math:`\\alpha`, and
-                    - `gamma`: :math:`\\gamma`
+    :param s2: kernel variance :math:`\\sigma^2`
+    :param alpha: kernel decay parameter :math:`\\alpha`
+    :param gamma: kernel length scale parameter :math:`\\gamma`
     """
+
+    _required_pars = ['s2', 'alpha', 'gamma']
+
     def __call__(self, x, y=None):
         if y is None:
             y = x
-        dists2, norms2_x, norms2_y = pw_dists2(x, y)
-        return self['s2'] * tf.exp(-self['alpha'] * (norms2_x + norms2_y)
-                                   - self['gamma'] * dists2)
+        dists2, norms2_x, norms2_y = pw_dists2(x, y, output_norms=True)
+        return self.s2 * tf.exp(-self.alpha * (norms2_x + norms2_y)
+                                - self.gamma * dists2)
 
 
-class UhlenbeckOrnstein(Kernel):
+class Exponential(Kernel):
     """
-    Uhlenbeck-Ornstein kernel.
+    Exponential kernel.
 
-    :param hyps: dictionary containing the following hyperparameters:
-
-                    - `s2`: :math:`\\sigma^2`,
-                    - `gamma`: :math:`\\gamma`
+    :param s2: kernel variance :math:`\\sigma^2`
+    :param gamma: kernel length scale parameter :math:`\\gamma`
     """
+
+    _required_pars = ['s2', 'gamma']
+
     def __call__(self, x, y=None):
         if y is None:
             y = x
-        dists2, norms2_x, norms2_y = pw_dists2(x, y)
-        return self['s2'] * tf.exp(-self['gamma'] * dists2 ** .5)
-
-
-class RLCSeriesCircuit(Kernel):
[…]

### 2017-03-01 src/core/par.py

@@ -1,0 +1,19 @@
+class Parametrisable(object):
+    """
+    Class whose keywords given to the constructor become attributes.
+
+    Required parameters can be specified through overriding `_required_pars`.
+    """
+    _required_pars = []
+
+    def __init__(self, **pars):
+        # Verify that all required parameters are specified
+        if not all(k in pars for k in self._required_pars):
+            unspecified_pars = [k for k in self._required_pars
+                                if k not in pars]
+            formatted = ', '.join('"{}"'.format(k) for k in unspecified_pars)
+            raise RuntimeError('must specify {}'.format(formatted))
+
+        # Set parameters
+        for k, v in pars.items():
+            setattr(self, k, v)

### 2017-03-01 src/core/plot.py

@@ -2,15 +2,20 @@ from mpl_toolkits.mplot3d import Axes3D
 import matplotlib as mpl
 import abc
 import warnings
-import itertools
 with warnings.catch_warnings():
     warnings.simplefilter('ignore')
     import matplotlib.pyplot as plt
 
 
-def to_rbga(color):
-    """ Convert `color` to a tuple `(red, blue, green, alpha)`. """
-    return mpl.colors.colorConverter.to_rgba(color, alpha=1.0)
+def to_rbga(color, alpha=1.0):
+    """
+    Convert `color` to a tuple `(red, blue, green, alpha)`.
+
+    :param color: color to convert
+    :param alpha: alpha
+    :return: converter color
+    """
+    return mpl.colors.colorConverter.to_rgba(color, alpha=alpha)
 
 
 # Mapping for values from strings
@@ -41,7 +46,7 @@ def dict_map(source, mapping):
     return result
 
 
-class Plotter:
+class Plotter(object):
     """
     Sane and consistent configuration of plots.
 
@@ -55,14 +60,14 @@ class Plotter:
                'grid_style': '--',
                'font_family': 'Adobe Caslon Pro',
                'font_size': 11,
-               'fig_size': (8, 6),
                'legend_color': '0.95',
+               'legend_location': 'upper right',
                'line_color': 'undefined',
-               'marker_style': 'None',
                'line_width': 1,
                'line_style': '-',
                'colorbar_shrink': 0.8,
                'colorbar_aspect': 20,
+               'marker_style': 'None',
                'marker_color': 'undefined',
                'marker_size': 5,
                'surface_rstride': 1,
@@ -72,45 +77,21 @@ class Plotter:
                'cmap': mpl.cm.coolwarm,
                'fill_alpha': 0.25,
                'fill_color': 'undefined',
-               'label': 'undefined'
-               }
+               'label': 'undefined',
+               'figure_size': (8, 6),
+               'figure_toolbar': 'None',  # None | toolbar2
+               'figure_autolayout': True}
 
     def __init__(self, **kw_args):
-        self._generate_abbrevs()
-        self._config.update(self._deabbrev(kw_args))
         self._config_global()
         self.plt = plt
         self._first = True
         self.plots = []
 
-    def _abbrev(self, xs, pref=''):
-        prefs = set([pref + x[len(pref)] for x in xs])
-        prefixed = [(x, list(set(filter(lambda y: y.startswith(x), xs))))
-                    for x in prefs]
-        abbrevs = {}
[…]

### 2017-03-01 src/core/sample.py

@@ -14,6 +14,8 @@ class ESS(object):
     :param x_init: initial state
     """
 
+    _min_theta = 1e-10
+
     def __init__(self, log_lik, sample_prior, x_init=None):
         self._log_lik = log_lik
         self._sample_prior = sample_prior
@@ -81,7 +83,7 @@ class ESS(object):
         :return: new state
         """
         self._draw_proposal(theta_l, theta_u)
-        theta_violation = abs(self._theta) < 1e-10
+        theta_violation = abs(self._theta) < self._min_theta
         if self._log_lik_x_proposed > u or theta_violation:
             # Proposal accepted
             if theta_violation:
@@ -120,6 +122,7 @@ class ESS(object):
                 eta = (time() - start) / (i + 1) * (num - i - 1)
                 time_attempt = 1000 * (end_it - start_it) / attempts
                 print 'ESS:'
+                print '  burning:      {}'.format('yes' if discard else 'no')
                 print '  iteration:    {:d}/{:d}'.format(i + 1, num)
                 print '  time left:    {:.0f} sec'.format(eta)
                 print '  attempts:     {}'.format(attempts)

### 2017-03-01 src/core/tfutil.py

@@ -1,37 +1,47 @@
 import tensorflow as tf
 import numpy as np
-import colorsys
-import scipy.stats
 import tensorflow.contrib.distributions as tf_dists
 from tensorflow.python.client import timeline
 from tensorflow.python.framework import ops
+
+import config
 
 from rpy2.robjects import r
 from rpy2.robjects.packages import importr
 import rpy2.robjects.numpy2ri
 
-import config
-
 rpy2.robjects.numpy2ri.activate()
 importr("pbivnorm")
 
-lower_perc = 100 * scipy.stats.norm.cdf(-2)
-upper_perc = 100 * scipy.stats.norm.cdf(2)
 
-inf = np.inf
-
-
-def pw_dists2(x, y):
+def pw_dists2(x, y, output_norms=False):
     """
     Compute pairwise distances between the rows of `x` and `y`.
 
     :param x: first set of features
     :param y: second set of features
-    :return: squared distances, norms of `x`, and norms of `y`
+    :param output_norms: boolean that indicates whether to also output the
+                         norms of `x` and `y`
+    :return: squared distances, further outputs norms of `x` and norms of `y`
+             if `output_norms` is set to `True`
     """
     norms2_x = tf.reduce_sum(x ** 2, reduction_indices=[1])[:, None]
     norms2_y = tf.reduce_sum(y ** 2, reduction_indices=[1])[None, :]
-    return norms2_x - 2 * mul(x, y, adj_b=True) + norms2_y, norms2_x, norms2_y
+    if output_norms:
+        return norms2_x - 2 * mul(x, y, adj_b=True) + norms2_y, norms2_x, \
+               norms2_y
+    else:
+        return norms2_x - 2 * mul(x, y, adj_b=True) + norms2_y
+
+
+def ctransp(x):
+    """
+    Batch conjugate transpose a tensor.
+
+    :param x: tensor
+    :return: tensor where the last two dimensions are conjugate transposed
+    """
+    return tf.conj(transp(x))
 
 
 def transp(x):
@@ -159,10 +169,10 @@ def trisolve3(a, b, c, lower_a=True, lower_c=True, adj_a=False, adj_c=False):
     :return: :math:`A^{-1}BC^{-1}`
     """
     return trisolve(a,
-                    transp(trisolve(c,
-                                    transp(b),
-                                    lower_a=lower_c,
-                                    adj_a=not adj_c)),
+                    ctransp(trisolve(c,
[…]

### 2017-03-01 src/core/util.py

@@ -1,0 +1,128 @@
+import colorsys
+import scipy.stats
+import numpy as np
+
+# Some handy constants
+lower_perc = 100 * scipy.stats.norm.cdf(-2)
+upper_perc = 100 * scipy.stats.norm.cdf(2)
+inf = np.inf
+
+
+def color_range(num, saturation=1.0, value=1.0):
+    """
+    Generate a range of colors that span hue.
+
+    :param num: num of colors in the range
+    :param saturation: saturation
+    :param value: value
+    :return: list of RGB colors
+    """
+    hsvs = [(x * 1.0 / num, saturation, value) for x in range(num)]
+    return map(lambda x: colorsys.hsv_to_rgb(*x), hsvs)
+
+
+def length_scale(ls):
+    """
+    Construct constant from length scale.
+
+    :param ls: length scale
+    :return: constant
+    """
+    return .5 / ls ** 2
+
+
+def is_inf(x):
+    """
+    Check if a variable is infinite.
+
+    :param x: variable to check
+    :return: boolean whether `x` is infinite
+    """
+    return is_numeric(x) and np.isinf(x)
+
+
+def is_numeric(x):
+    """
+    Check if a variable is numeric.
+
+    :param x: variable
+    :return: boolean indicating whether `x` is numeric
+    """
+    try:
+        float(x)
+        return True
+    except ValueError:
+        return False
+    except AttributeError:
+        return False
+
+
+def fft(*args, **kw_args):
+    """
+    Alias for `np.fft.fft` which afterwards applies `np.fft.fftshift`.
+    """
+    return np.fft.fftshift(np.fft.fft(*args, **kw_args))
+
+
+def fft_freq(*args, **kw_args):
+    """
+    Alias for `np.fft.fft_freq` which afterwards applies `np.fft.fftshift`.
+    """
+    return np.fft.fftshift(np.fft.fftfreq(*args, **kw_args))
+
+
+def psd(*args, **kw_args):
+    """
+    Similar to `np.fft.fft`, but computes the PSD in dB instead.
+    """
+    return 10 * np.log10(np.abs(fft(*args, **kw_args)))
+
+
+def zero_pad(x, num, axis=0):
+    """
+    Zero pad an array.
+
+    :param x: array
+    :param axis: axis to zero pad along
+    :param num: number of zeros
+    :return: zero-padded array
+    """
+    zeros = np.zeros(num)
+    if axis > 0:
+        for i in range(axis - 1):
[…]

### 2017-03-02 src/core/cgpcm.py
Further refactoring of core code

@@ -8,12 +8,80 @@ from tfutil import *
 from util import *
 
 
-class GPCM(Parametrisable):
+def cgpcm_pars(sess,
+               t,
+               y,
+               nx,
+               nh,
+               k_len,
+               k_wiggles,
+               causal,
+               causal_id=False):
+    """
+    Create parameters for the CGPCM.
+
+    :param sess: TensorFlow session
+    :param t: times
+    :param y: observations
+    :param nx: number of inducing points for noise
+    :param nh: number of inducing points for filter
+    :param k_len: length of kernel
+    :param k_wiggles: number of wiggles in kernel
+    :param causal: boolean that indicates whether the model is causal
+    :param causal_id: boolean that indicates whether the interdomain
+                      transformation is causal
+    :return: parameters
+    """
+    # Equal prior power
+    if causal:
+        k_len *= 2
+
+    # Hyperparameters
+    alpha = length_scale(k_len)
+    gamma = length_scale(float(k_len) / k_wiggles)
+
+    # Hyperparameter for x and inducing points for x
+    if nx > 0:
+        omega = length_scale((max(t) - min(t)) / nx)
+        tx = np.linspace(min(t), max(t), nx)
+
+    # Make sure of an odd number of inducing points for h, so as to have one
+    # at zero
+    if nh % 2 == 0:
+        nh += 1
+
+    # Inducing points for h
+    if causal:
+        th = np.linspace(0, 2 * k_len, nh)
+        dth = th[1] - th[0]
+        th -= dth * 3
+    else:
+        th = np.linspace(-2 * k_len, 2 * k_len, nh)
+
+    pars = {'sess': sess,
+            't': tf.constant(t),
+            'y': tf.constant(y),
+            'th': tf.constant(th),
+            's2': var_pos(to_float(1e-4), 's2'),
+            's2_f': var_pos(to_float(1), 's2_f'),
+            'alpha': var_pos(to_float(alpha), 'alpha'),
+            'gamma': var_pos(to_float(gamma), 'gamma'),
+            'causal': causal,
+            'causal_id': causal_id}
+    if nx > 0:
+        pars['tx'] = tf.constant(tx)
+        pars['omega'] = var_pos(to_float(omega), 'omega')
+
+    initialise_uninitialised_variables(sess)
[…]

### 2017-03-02 src/core/data.py

@@ -1,0 +1,246 @@
+import scipy.io as sio
+import scipy.io.wavfile as sio_wav
+
+from tfutil import *
+from util import *
+import cgpcm
+import kernel
+
+
+class Data(object):
+    """
+    Data.
+
+    :param t: times
+    :param y: values
+    """
+
+    def __init__(self, t, y):
+        self.t = t
+        self.y = y
+        self.n = shape(t)[0]
+
+    def subsample(self, n):
+        """
+        Subsample the data.
+
+        :param n: number of points in subsample
+        :return: subsampled data and remainder of data
+        """
+        return self._split(np.random.choice(self.n, n, replace=False))
+
+    def fragment(self, length, start=0):
+        """
+        Fragment the data.
+
+        :param length: length of fragment
+        :param start: start of fragment
+        :return: fragmented data
+        """
+        return self._split(range(start, start + length))
+
+    def _split(self, i_in):
+        i_in = sorted(i_in)
+        i_out = sorted(list(set(range(self.n)) - set(i_in)))
+        return Data(self.t[i_in], self.y[i_in]), \
+               Data(self.t[i_out], self.y[i_out])
+
+    def smse(self, y):
+        """
+        Compute the standardised mean squared error for some predictions.
+
+        :param y: predictions
+        :return: standardised means squared error
+        """
+        return smse(y, self.y)
+
+    def mll(self, mu, std):
+        """
+        Compute the mean log loss for some predictions.
+
+        :param mu: mean of predictions
+        :param std: standard deviation of predictions
+        :return: mean log loss
+        """
+        return mll(mu, std, self.y)
+
+    def make_noisy(self, var):
+        """
+        Make the data noisy.
+
+        :param var: variance of noise
+        :return: noisy data
+        """
+        noise = var ** .5 * np.random.randn(*shape(self.y))
+        return Data(self.t, self.y + noise)
+
+    @property
+    def mean(self):
+        """
+        Mean of data.
+        """
+        return np.mean(self.y)
+
+    @property
+    def std(self):
+        """
+        Standard deviation of data.
+        """
+        return np.std(self.y, ddof=1)
+
+    def __neg__(self):
[…]

### 2017-03-02 src/core/eq.py

@@ -2,10 +2,6 @@ from core.tfutil import *
 from core.util import *
 import operator
 
-if __name__ == '__main__':
-    # For debugging convenience
-    sess = tf.Session()
-
 
 def const(x):
     """
@@ -31,6 +27,7 @@ def var(x, power=1):
 def kh_constructor(alpha, gamma):
     """
     Create a function that calls `kh` for some given hyperparameters.
+
     :param alpha: kernel decay parameter :math:`\\alpha`
     :param gamma: kernel length scale parameter :math:`\\gamma`
     :return: constructor
@@ -59,6 +56,7 @@ def kh(alpha, gamma, x, y):
 def kxs_constructor(omega):
     """
     Create a function that calls `kxs` for some given hyperparameter.
+
     :param omega: kernel length scale parameter :math:`\\omega`
     :return: constructor
     """
@@ -562,30 +560,3 @@ class EQ(object):
             return str(self._const)
         else:
             return '{} exp({})'.format(self._const, str(self._poly))
-
-
-if __name__ == '__main__':
-    # Small test case. Probably should use units tests.
-
-    t1, t2, t3 = var('t1'), var('t2'), var('t3')
-
-    eq = EQ(- const(1) * t1 ** 2
-            - const(2) * t2 ** 2
-            - const(.5) * t1 * t2
-            - const(2) * t1 * t3
-            + const(3) * t2
-            + const(4))
-    var_map = {'t3': tf.constant(np.eye(2))}
-    print sess.run(eq.integrate_half('t1', 't2', **var_map))
-    print 'Mathematica\'s NIntegrate: 55.8181 and 11.7677'
-
-    print sess.run(eq.integrate_box(('t1', const(-1), const(2)),
-                                    ('t2', t3, const(3)),
-                                    **var_map))
-    print 'Mathematica\'s NIntegrate: 318.354 and 217.392'
-
-    eq = EQ(const(-1) * t1 ** 2
-            + const(-.5) * t1
-            + const(4.))
-    print sess.run(eq.integrate_half('t1'))
-    print 'Mathematica\'s NIntegrate: 65.7397'

### 2017-03-02 src/core/eq_test.py

@@ -1,0 +1,46 @@
+import os
+
+os.environ['TF_CPP_MIN_LOG_LEVEL'] = '3'
+import tensorflow as tf
+
+from eq import *
+import unittest
+
+
+class TestEQ(unittest.TestCase):
+    def setUp(self):
+        self.sess = tf.Session()
+        self.t1 = var('t1')
+        self.t2 = var('t2')
+        self.t3 = var('t3')
+        self.exp1 = EQ(- const(1) * self.t1 ** 2
+                       - const(2) * self.t2 ** 2
+                       - const(.5) * self.t1 * self.t2
+                       - const(2) * self.t1 * self.t3
+                       + const(3) * self.t2
+                       + const(4))
+        a
+        self.exp2 = EQ(const(-1) * self.t1 ** 2
+                       + const(-.5) * self.t1
+                       + const(4))
+        self.var_map = {'t3': tf.constant(np.eye(2))}
+
+    def test1(self):
+        ref = np.array([[55.8181, 11.7677],
+                        [11.7677, 55.8181]])
+        res = self.sess.run(self.exp1.integrate_box(('t1', -inf, 0),
+                                                    ('t2', -inf, 0),
+                                                    **self.var_map))
+        np.testing.assert_almost_equal(res, ref)
+
+    def test2(self):
+        ref = np.array([[318.354, 217.392],
+                        [217.392, 318.354]])
+        box = [('t1', const(-1), const(2)), ('t2', self.t3, const(3))]
+        res = self.sess.run(self.exp1.integrate_box(box, **self.var_map))
+        np.testing.assert_almost_equal(res, ref)
+
+    def test3(self):
+        ref = 65.7397
+        res = self.sess.run(self.exp2.integrate_half('t1'))
+        np.testing.assert_almost_equal(res, ref)

### 2017-03-02 src/core/par.py

@@ -17,3 +17,5 @@ class Parametrisable(object):
         # Set parameters
         for k, v in pars.items():
             setattr(self, k, v)
+
+        self._pars = pars

### 2017-03-02 src/core/util.py

@@ -51,6 +51,8 @@ def is_numeric(x):
     try:
         float(x)
         return True
+    except TypeError:
+        return False
     except ValueError:
         return False
     except AttributeError:
@@ -126,3 +128,14 @@ def vlinspace(*args, **kw_args):
     Alias for `np.linspace`, but returns a vector instead.
     """
     return np.array([np.linspace(*args, **kw_args)]).T
+
+
+def nearest_index(xs, y):
+    """
+    Get the index of the element in `xs` that is nearest to `y`.
+
+    :param xs: list
+    :param y: element
+    :return: index of element in `xs` that is nearest to `y`
+    """
+    return np.argmin(np.abs(xs - y))

### 2017-03-02 src/learn/hrtf.py
Added tasks in raw form

@@ -4,9 +4,9 @@ import pickle
 import scipy.io.wavfile as wav
 import scipy.io as sio
 
-from core.gpcm import VGPCM
-from core.plotter import Plotter2D
-from core.utils import *
+from core.cgpcm import VCGPCM
+from core.plot import Plotter2D
+from core.tfutil import *
 
 
 def normalise_energy(x, dt, causal, i_mid=None):
@@ -78,7 +78,7 @@ if do_train:
     tk = np.linspace(-k_stretch * 2 * k_len, k_stretch * 2 * k_len, nk)[:,
          None]
 
-    h = wav.read('R-10e355a.wav')[1].astype(float)
+    h = wav.read('KEMAR_R10e355a.wav')[1].astype(float)
     x = np.random.randn(1000)
     y = np.convolve(x, h)
     y = (y - np.mean(y)) / np.std(y, ddof=1)  # Normalise
@@ -102,14 +102,14 @@ if do_train:
     # t_sample = np.concatenate((t[:imp_region[0]], t[imp_region[1] - 1:]))
 
     print 'Constructing model...'
-    mod = VGPCM(pars,
-                tf.constant(th[:, 0]),
-                tf.constant(tx[:, 0]),
-                t_sample,
-                y_sample,
-                sess,
-                causal=causal,
-                causal_id=False)
+    mod = VCGPCM(pars,
+                 tf.constant(th[:, 0]),
+                 tf.constant(tx[:, 0]),
+                 t_sample,
+                 y_sample,
+                 sess,
+                 causal=causal,
+                 causal_id=False)
     initialise_uninitialised_variables(sess)
     #
     # if train_iters_pre == 0 and train_iters == 0:

### 2017-03-02 src/learn/timit.py

@@ -3,9 +3,9 @@ from tensorflow.contrib.opt.python.training.external_optimizer import \
 import pickle
 import scipy.io as sio
 
-from core.gpcm import VGPCM
-from core.plotter import Plotter2D
-from core.utils import *
+from core.cgpcm import VCGPCM
+from core.plot import Plotter2D
+from core.tfutil import *
 
 
 def normalise_energy(x, dt, causal):
@@ -77,7 +77,7 @@ if do_train:
     tk = np.linspace(-k_stretch * 2 * k_len, k_stretch * 2 * k_len, nk)[:,
          None]
 
-    audio = sio.loadmat('audioWS.mat')['audio']
+    audio = sio.loadmat('TIMIT_unknown.mat')['audio']
     margin = 0
     y = audio[
         11499 + margin:11499 + margin + m].squeeze()  # Section from paper
@@ -95,14 +95,14 @@ if do_train:
     t_sample = t[choice]
 
     print 'Constructing model...'
-    mod = VGPCM(pars,
-                tf.constant(th[:, 0]),
-                tf.constant(tx[:, 0]),
-                t_sample,
-                y_sample,
-                sess,
-                causal=causal,
-                causal_id=False)
+    mod = VCGPCM(pars,
+                 tf.constant(th[:, 0]),
+                 tf.constant(tx[:, 0]),
+                 t_sample,
+                 y_sample,
+                 sess,
+                 causal=causal,
+                 causal_id=False)
     initialise_uninitialised_variables(sess)
 
     if train_iters_pre == 0 and train_iters == 0:

### 2017-03-02 src/learn/timit2.py

@@ -3,9 +3,9 @@ from tensorflow.contrib.opt.python.training.external_optimizer import \
 import pickle
 import scipy.io as sio
 
-from core.gpcm import VGPCM
-from core.plotter import Plotter2D
-from core.utils import *
+from core.cgpcm import VCGPCM
+from core.plot import Plotter2D
+from core.tfutil import *
 
 
 def normalise_energy(x, dt, causal):
@@ -76,7 +76,7 @@ if do_train:
     tk = np.linspace(-k_stretch * 2 * k_len, k_stretch * 2 * k_len, nk)[:,
          None]
 
-    audio = sio.loadmat('audioWS.mat')['audio']
+    audio = sio.loadmat('TIMIT_unknown.mat')['audio']
     margin = 200
     y = audio[11499 + margin:11499 + margin + data_size].squeeze()
     y = (y - np.mean(y)) / np.std(y, ddof=1)  # Normalise
@@ -99,14 +99,14 @@ if do_train:
     free_inds[free_inds >= imp_region[0]] += imp_region[1] - imp_region[0]
 
     print 'Constructing model...'
-    mod = VGPCM(pars,
-                tf.constant(th[:, 0]),
-                tf.constant(tx[:, 0]),
-                t_sample,
-                y_sample,
-                sess,
-                causal=causal,
-                causal_id=False)
+    mod = VCGPCM(pars,
+                 tf.constant(th[:, 0]),
+                 tf.constant(tx[:, 0]),
+                 t_sample,
+                 y_sample,
+                 sess,
+                 causal=causal,
+                 causal_id=False)
     initialise_uninitialised_variables(sess)
 
     if train_iters_pre == 0 and train_iters == 0:

### 2017-03-02 src/learn/toy.py

@@ -2,9 +2,9 @@ from tensorflow.contrib.opt.python.training.external_optimizer import \
     ScipyOptimizerInterface as SciPyOpt
 import pickle
 
-from core.gpcm import AKM, VGPCM
-from core.plotter import Plotter2D
-from core.utils import *
+from core.cgpcm import AKM, VCGPCM
+from core.plot import Plotter2D
+from core.tfutil import *
 
 
 def normalise_energy(x, dt, causal):
@@ -117,14 +117,14 @@ if do_train:
     akm.sample['h'] = h_samp
 
     if test_kernel:
-        k = akm.construct_kernel(causal=causal_sample)
+        k = akm.k(causal=causal_sample)
         p = Plotter2D()
         p.plot(tk, k)
         p.show()
 
-    k_true, y, _ = akm.construct_sample(causal=causal_sample)
+    k_true, y, _ = akm.f(causal=causal_sample)
     k_true, y = normalise(k_true[:, None]), y[:, None]
-    h_true = normalise_energy(akm.construct_filter(tf.constant(tk)),
+    h_true = normalise_energy(akm.h(tf.constant(tk)),
                               dt=tk[1, 0] - tk[0, 0], causal=causal_sample)
     y_noisy = y + np.random.randn(n, 1) * np.sqrt(noise)
 
@@ -139,14 +139,14 @@ if do_train:
     t_sample = t[choice, 0]
 
     print 'Constructing model...'
-    mod = VGPCM(pars,
-                tf.constant(th[:, 0]),
-                tf.constant(tx[:, 0]),
-                t_sample,
-                y_sample,
-                sess,
-                causal=causal,
-                causal_id=False)
+    mod = VCGPCM(pars,
+                 tf.constant(th[:, 0]),
+                 tf.constant(tx[:, 0]),
+                 t_sample,
+                 y_sample,
+                 sess,
+                 causal=causal,
+                 causal_id=False)
     initialise_uninitialised_variables(sess)
 
     if init_at_optimal_mean:

### 2017-03-02 src/sample_akm_causal.py
Added sample scripts

@@ -1,0 +1,165 @@
+from core.cgpcm import AKM
+from core.plot import Plotter2D
+from core.util import *
+from core.kernel import DEQ
+from core.dist import Normal
+
+sess = tf.Session()
+
+seed = 2147483647
+tf.set_random_seed(seed)
+np.random.seed(seed)
+
+
+def length_scale(ls):
+    return .5 / ls ** 2
+
+
+def r2z(r):
+    return 2 ** .5 * r / (2 - r ** 2) ** .5
+
+
+k_len = .5
+n = 150
+nh = 31
+nx = 100
+noise_init = 1e-3
+noise = 0
+deq_len = k_len * .3
+pre_train = True
+k_stretch = 2
+pars = {'alpha': length_scale(k_len),
+        'gamma': var_pos(to_float(length_scale(k_len * r2z(.3))), 'gamma'),
+        'omega': length_scale(2. / nx),
+        'K': 1, 'M': 1, 'N': 1,
+        's2': var_pos(to_float(noise_init), 's2'),
+        's2_f': var_pos(to_float(1.), 's2_f')}
+sess.run(tf.global_variables_initializer())
+
+t = np.linspace(-1, 1, n)
+th = np.linspace(-k_stretch * k_len, k_stretch * k_len, nh)
+tx = np.linspace(-1, 1, nx)
+tk = np.linspace(-k_stretch * 2 * k_len, k_stretch * 2 * k_len, 500)
+
+deq = DEQ({'alpha': 0.,
+           'gamma': length_scale(deq_len),
+           's2': 1.})
+y = sess.run(Normal(reg(deq(t[:, None]))).sample())
+k_true = sess.run(deq(tk[:, None], tk[:1][:, None]))
+
+y_noisy = y + np.random.randn(n, 1) * np.sqrt(noise)
+
+print 'Constructing model...'
+mod = AKM(pars,
+          tf.constant(th),
+          tf.constant(tx),
+          t,
+          y_noisy,
+          sess)
+
+print 'Sampling kernels...'
+k_mu, k_lower, k_upper = mod.prior_kernel(tk, iters=5000, causal=False)
+k_mu_c, k_lower_c, k_upper_c = mod.prior_kernel(tk, causal=True, iters=5000)
+
+
[…]

### 2017-03-02 src/sample_akm_causal_interpolation.py

@@ -1,0 +1,156 @@
+from core.cgpcm import AKM
+from core.plot import Plotter2D
+from core.util import *
+from core.kernel import DEQ
+from core.dist import Normal
+
+config.reg = 1e-6
+
+sess = tf.Session()
+
+# 1212
+
+seed = 99998 # 2147483647
+tf.set_random_seed(seed)
+np.random.seed(seed)
+
+
+def length_scale(ls):
+    return .5 / ls ** 2
+
+
+k_len = .4
+gamma_frac = .5
+n = 450
+nh = 31
+nx = 100  # Irrelevant
+noise_init = 1e-3  # Irrelevant
+noise = 0
+k_stretch = 2
+pars = {'alpha': length_scale(k_len),
+        'gamma': var_pos(to_float(length_scale(k_len * gamma_frac)), 'gamma'),
+        'omega': length_scale(2. / nx),
+        'K': 1, 'M': 1, 'N': 1,
+        's2': var_pos(to_float(noise_init), 's2'),
+        's2_f': var_pos(to_float(1.), 's2_f')}
+sess.run(tf.global_variables_initializer())
+
+t = np.linspace(-1, 1, n)
+th = np.linspace(-k_stretch * k_len, k_stretch * k_len, nh)
+tx = np.linspace(-1, 1, nx)
+tk = np.linspace(-k_stretch * 2 * k_len, k_stretch * 2 * k_len, 501)
+tk_pos = np.linspace(0, k_stretch * 2 * k_len, 501)
+
+print 'Constructing model...'
+mod = AKM(pars,
+          tf.constant(th),
+          tf.constant(tx),
+          t,
+          None,
+          sess)
+
+
+# Compute filters
+def h_conditioned(h0):
+    deq = DEQ({'alpha': pars['alpha'],
+               'gamma': pars['gamma'],
+               's2': 1.})
+    thu = np.array([h0])
+    Kh = deq(th[:, None])
+    Ku = deq(thu[:, None])
+    Kuh = deq(thu[:, None], th[:, None])
+    A = trisolve(tf.cholesky(reg(Ku)), Kuh)
+    Kh = reg(Kh - mul(A, A, adj_a=True))
+    h = sess.run(Normal(reg(Kh)).sample())
+    h[(nh - 1) / 2] = h0  # Correct for round-off errors
+    return h
+
+
[…]

### 2017-03-02 src/sample_akm_causal_psd.py

@@ -1,0 +1,102 @@
+from core.cgpcm import AKM
+from core.plot import Plotter2D
+from core.util import *
+from core.kernel import DEQ
+from core.dist import Normal
+
+sess = tf.Session()
+
+# seed = 2147483647
+# tf.set_random_seed(seed)
+# np.random.seed(seed)
+
+
+def length_scale(ls):
+    return .5 / ls ** 2
+
+
+def r2z(r):
+    return 2 ** .5 * r / (2 - r ** 2) ** .5
+
+
+k_len = .2
+frac_gamma = .2
+n = 200
+nh = 200
+nx = 100
+noise_init = 1e-3
+noise = 0
+pre_train = True
+k_stretch = 2
+pars = {'alpha': length_scale(k_len),
+        'gamma': var_pos(to_float(length_scale(frac_gamma * k_len)), 'gamma'),
+        'omega': length_scale(2. / nx),
+        'K': 1, 'M': 1, 'N': 1,
+        's2': var_pos(to_float(noise_init), 's2'),
+        's2_f': var_pos(to_float(1.), 's2_f')}
+sess.run(tf.global_variables_initializer())
+
+t = np.linspace(-1, 1, n)
+th = np.linspace(-k_stretch * k_len, k_stretch * k_len, nh)
+tx = np.linspace(-1, 1, nx)
+tk_stretch = 10
+tk = np.linspace(-tk_stretch * k_len, tk_stretch * k_len, 1000)
+
+print 'Constructing model...'
+mod = AKM(pars,
+          tf.constant(th),
+          tf.constant(tx),
+          t,
+          None,
+          sess)
+
+mod.generate_sample(t, tk)
+# print 'Acausal'
+# k_samp, f_samp, K = mod.construct_sample()
+# print 'Causal'
+# k_samp_c, f_samp_c, K_c = mod.construct_sample(causal=True)
+
+print 'Acausal'
+k_samp = mod.construct_kernel()
+print 'Causal'
+k_samp_c = mod.construct_kernel(causal=True)
+
+
+def psd_freqs(t):
+    return np.fft.fftshift(np.fft.fftfreq(len(t), t[1] - t[0]))
+
+
+def psd(f):
+    return 10 * np.log(np.abs(np.fft.fftshift(np.fft.fft(f))) ** 2)
+
+
[…]

### 2017-03-02 src/test_bvn_cdf_benchmark.py
Code for testing BVNCDF

@@ -1,4 +1,4 @@
-from core.utils import _mvn_cdf, _bvn_cdf
+from core.tfutil import _mvn_cdf, _bvn_cdf
 import numpy as np
 from time import time
 

### 2017-03-02 src/test_bvn_cdf_gradient.py

@@ -1,4 +1,4 @@
-from core.utils import *
+from core.tfutil import *
 
 
 with tf.Session() as sess:

### 2017-03-02 src/generate_doc_source.py
Documentation

@@ -1,18 +1,14 @@
 import os
-import subprocess as sp
-import sqlite3
-import datetime
-
 
 documentation = [
     {'package': 'core',
      'title': 'Core Modules',
      'description': 'Important modules that make up the system',
      'subpackages': []},
-     {'package': 'learn',
-      'title': 'Learning Tasks',
-      'description': 'Scripts to perform learning tasks',
-      'subpackages': []}
+    # {'package': 'learn',
+    #  'title': 'Learning Tasks',
+    #  'description': 'Scripts to perform learning tasks',
+    #  'subpackages': []}
 ]
 
 
@@ -25,7 +21,7 @@ def walk_dir(dir_path):
     """
     files = []
     for file in os.listdir(dir_path):
-        if file.startswith('.'):
+        if file == '.' or file == '..':
             continue
         file_path = os.path.join(dir_path, file)
         if os.path.isfile(file_path):
@@ -49,7 +45,7 @@ line2 = '-' * 100
 line3 = '^' * 100
 
 template_index = '''
-GGPCM Documentation
+CGPCM Documentation
 {line1:s}
 
 .. toctree::
@@ -98,8 +94,7 @@ Outcome
 '''
 
 
-
-class Documentation:
+class Documentation(object):
     """
     Generator of documentation source.
     """
@@ -144,12 +139,14 @@ class Documentation:
     def _render_package(self, package):
         # Gather the packages and the subpackages to be included
         packages = [package['package']] \
-                  + map(lambda x: '{}.{}'.format(package['package'], x),
-                        package['subpackages'])
+                   + map(lambda x: '{}.{}'.format(package['package'], x),
+                         package['subpackages'])
+
         # Extract all modules
         paths = map(self._import_to_path, packages)
         modules = sum(map(lambda x: self._files_to_modules(walk_dir(x)),
                           paths), [])
+
         # Render the modules
         return template_package.format(
             description=package['description'],

### 2017-03-02 src/source/conf.py

@@ -55,7 +55,7 @@ source_suffix = '.rst'
 master_doc = 'index'
 
 # General information about the project.
-project = u'GGPCM'
+project = u'CGPCM'
 copyright = u'Wessel Bruinsma'
 author = u'Wessel Bruinsma'
 
@@ -126,7 +126,7 @@ todo_include_todos = True
 # The theme to use for HTML and HTML Help pages.  See the documentation for
 # a list of builtin themes.
 #
-html_theme = 'agogo'
+html_theme = 'haiku'
 
 def setup(app):
     app.add_stylesheet('custom.css')

### 2017-03-02 src/config.py
Config

@@ -1,5 +1,4 @@
 import tensorflow as tf
 
-
 reg = 1e-5
 dtype = tf.float64

### 2017-03-02 src/core/cgpcm.py
Upgraded to TF1.0

@@ -403,7 +403,7 @@ class VCGPCM(CGPCM):
     def _init_inducing_points(self):
         mean_init = tf.Variable(self.h_prior.sample(), name='muh')
         var_init = tf.Variable(tf.cholesky(self.h_prior.variance), name='Sh')
-        tf.initialize_variables([mean_init, var_init])
+        tf.variables_initializer([mean_init, var_init])
         self.h = Normal(mul(var_init, var_init, adj_a=True), mean_init)
 
     def _qz_natural(self, h_mean, h_m2):

### 2017-03-02 src/core/tfutil.py

@@ -25,8 +25,8 @@ def pw_dists2(x, y, output_norms=False):
     :return: squared distances, further outputs norms of `x` and norms of `y`
              if `output_norms` is set to `True`
     """
-    norms2_x = tf.reduce_sum(x ** 2, reduction_indices=[1])[:, None]
-    norms2_y = tf.reduce_sum(y ** 2, reduction_indices=[1])[None, :]
+    norms2_x = tf.reduce_sum(x ** 2, axis=[1])[:, None]
+    norms2_y = tf.reduce_sum(y ** 2, axis=[1])[None, :]
     if output_norms:
         return norms2_x - 2 * mul(x, y, adj_b=True) + norms2_y, norms2_x, \
                norms2_y
@@ -144,7 +144,7 @@ def mul(a, b, adj_a=False, adj_b=False):
     """
     Alias for `tf.batch_matmul`.
     """
-    return tf.batch_matmul(a, b, adj_x=adj_a, adj_y=adj_b)
+    return tf.matmul(a, b, adj_x=adj_a, adj_y=adj_b)
 
 
 def trisolve(a, b, lower_a=True, adj_a=False):

### 2017-03-14 src/experiment_compare.sh
Added handy experimentation scripts

@@ -1,0 +1,5 @@
+#!/usr/bin/env bash
+set -e
+./train.py -t $@ causal-model -t $@
+./plot.py compare -t $@ causal-model -t $@
+

### 2017-03-14 src/experiment_compare_series.sh

@@ -1,0 +1,12 @@
+#!/usr/bin/env bash
+task=$1
+seed=$2
+to_resample=$3
+opts=${@:4:${#}}
+for (( i=0; i<=$to_resample; i++ ))
+do
+    ./experiment_compare.sh $task seed $seed resample $i $opts
+done
+
+
+

### 2017-03-14 src/train_plot.sh

@@ -1,0 +1,5 @@
+#!/usr/bin/env bash
+set -e
+./train.py -t $@
+./plot.py full -t $@ -s
+

### 2017-03-14 src/core/cgpcm.py
Further changes to core code
Mainly options for tasks, updated plotting, and bug fixes

@@ -31,7 +31,7 @@ class CGPCM(Parametrisable):
 
     @classmethod
     def from_recipe(cls, sess, e, nx, nh, k_len, k_wiggles, causal,
-                    causal_id=False):
+                    causal_id=False, noise_init=1e-4):
         """
         Generate parameters for the CGPCM and construct afterwards.
 
@@ -43,14 +43,14 @@ class CGPCM(Parametrisable):
         :param k_wiggles: number of wiggles in kernel
         :param causal: causal model
         :param causal_id: causal interdomain transformation
+        :param noise_init: initialisation of noise
         :return: CGPCM instance
         """
         # Trainable variables
         vars = {}
 
         # Config
-        noise_init = 1e-4
-        k_stretch = 3
+        k_stretch = 2
         causal_extra_points = 3
 
         # Compute effective kernel length to match equal prior power

### 2017-03-14 src/core/exp.py

@@ -62,6 +62,112 @@ class Task(object):
         del self.mod
 
 
+class TaskOptions(object):
+    """
+    Options for task.
+    """
+
+    def __init__(self):
+        self._options = []
+
+    def add_option(self, name, desc='no description available'):
+        """
+        Add a boolean option.
+
+        :param name: name of options
+        :param desc: description of option
+        """
+        self._options.append({'has_value': False,
+                              'name': name.lower(),
+                              'value': False,
+                              'description': desc,
+                              'required': False})
+
+    def add_value_option(self, name, value_type,
+                         desc='no description available', required=False):
+        """
+        Add an option with a value.
+
+        :param name: name of option
+        :param value_type: type of option, should be callable
+        :param desc: description of option
+        :param required: option is required
+        """
+        self._options.append({'has_value': True,
+                              'name': name.lower(),
+                              'value_type': value_type,
+                              'value': None,
+                              'description': desc,
+                              'required': required})
+
+    def _get_option(self, name):
+        for option in self._options:
+            if name == option['name']:
+                return option
+        raise RuntimeError('option "{}" not found'.format(name))
+
+    def parse(self, args):
+        """
+        Parse arguments.
+
+        :param args: arguments
+        """
+        self._options = sorted(self._options, key=lambda x: x['name'])
+        self._parse_help(args)
+        self._parse_args(args)
+        self._parse_required()
+
+    def _parse_required(self):
+        missing = []
+        for option in self._options:
+            if option['required'] and option['value'] is None:
+                missing.append(option['name'])
+        if len(missing) == 1:
+            raise RuntimeError('missing option "{}"'.format(missing[0]))
+        elif len(missing) > 1:
+            missing_string = ', '.join(['"{}"'.format(x) for x in missing])
+            raise RuntimeError('missing options {}'.format(missing_string))
+
+    def _parse_args(self, args):
+        it = iter(args)
+        for arg in it:
+            option = self._get_option(arg)
+            if option['has_value']:
+                option['value'] = option['value_type'](next(it))
+            else:
+                option['value'] = True
+
+    def _parse_help(self, args):
+        if 'help' in args:
+            out.section('options for task')
+            for option in self._options:
[…]

### 2017-03-14 src/core/out.py

@@ -3,7 +3,7 @@ import sys
 level = 0
 total_width = 25
 width = 25
-indenter = '|  '
+indenter = '    '
 
 
 def _out(msg):

### 2017-03-14 src/core/util.py

@@ -2,6 +2,7 @@ import colorsys
 import scipy.stats
 import numpy as np
 import tensorflow as tf
+import os
 
 from tfutil import shape
 
@@ -173,3 +174,14 @@ def dict2(d=None, **kw_args):
         d = dict(d)
     d.update(kw_args)
     return d
+
+
+def mkdirs(path):
+    """
+    Ensure that all directories in a file path exist.
+
+    :param path: file path
+    """
+    dir_path = os.path.dirname(path)
+    if not os.path.exists(dir_path):
+        os.makedirs(dir_path)

### 2017-03-14 src/plot.py
More refined paths for cache and output

@@ -4,6 +4,7 @@ import pickle
 import argparse
 
 import core.out as out
+import core.util as util
 import core.exp as exp
 
 parser = argparse.ArgumentParser(description='Run a task.')
@@ -33,7 +34,9 @@ out.section_end()
 
 out.section('plotting')
 p, fn = exp.plot(args.plot[0], tasks, options=args.plot[1:])
-p.save('output/out_{}_{}.pdf'.format(args.plot[0], fn))
+fn = 'output/{}/{}.pdf'.format(args.plot[0], fn)
+util.mkdirs(fn)
+p.save(fn)
 if args.show:
     p.show()
 out.section_end()

### 2017-03-14 src/train.py

@@ -40,7 +40,9 @@ for i, task_args in enumerate(args.task):
         # Save resulting task
         out.section('saving')
         task.make_pickleable(sess)
-        with open('tasks/cache/{}.pickle'.format(task.config.fn), 'w') as f:
+        fn = 'tasks/cache/{}.pickle'.format(task.config.fn)
+        util.mkdirs(fn)
+        with open(fn, 'w') as f:
             pickle.dump(task, f)
         out.section_end()
 

### 2017-03-14 src/tasks/toy.py
Working toy experiment in old parametrisation

@@ -1,4 +1,4 @@
-from core.exp import Task, TaskConfig
+from core.exp import Task, TaskConfig, TaskOptions
 from core.cgpcm import VCGPCM
 import core.data as data
 
@@ -6,69 +6,72 @@ import core.data as data
 class Experiment(Task):
     """
     Toy experiment.
-
-    The first option is
-
-    Options:
-
-        - resample: number of times to resample, this option is required and
-                    must be first,
-        - causal: generate a causal sample,
-        - causal-model: use the GPCM instead of the CGPCM, and
-        - post: use posttraining.
     """
 
-    def generate_config(self, options):
-        # Parse options
-        if len(options) == 0:
-            raise RuntimeError('first option is required')
-        resample = int(options[0])
-        causal = 'causal' in options
-        causal_mod = 'causal-model' in options
-        post = 'post' in options
-
-        fn = 'toy_{}_{}_sample_{}{}'.format(resample,
-                                            'causal' if causal else 'acausal',
-                                            'cgpcm' if causal_mod else 'gpcm',
-                                            '_post' if post else '')
+    def generate_config(self, args):
+        options = TaskOptions()
+        options.add_option('causal', 'generate a causal sample')
+        options.add_option('causal-model', 'use the causal model')
+        options.add_option('post', 'use posttraining')
+        options.add_option('no-pre', 'no pretraining')
+        options.add_option('noisy', 'noisy sample')
+        options.add_option('small', 'small sample')
+        options.add_option('extensive', 'extensive training')
+        options.add_value_option('resample', int,
+                                 'number of times to resamples',
+                                 required=True)
+        options.add_value_option('seed', int, 'seed', required=True)
+        options.parse(args)
 
         return TaskConfig(name='Toy Experiment',
-                          seed=85,
-                          fn=fn,
+                          seed=options['seed'],
+                          fn=options.fn(group_by=['causal', 'seed', 'small']),
 
                           # Training options
-                          iters_pre=200,
-                          iters=500,
-                          iters_post=50 if post else 0,
-                          samps=300,
+                          iters_pre=0 if options['no-pre'] else 250,
+                          iters=25000 if options['extensive'] else 1000,
+                          iters_post=50 if options['post'] else 0,
+                          samps=200,
 
                           # Sample options
-                          causal=causal,
-                          causal_model=causal_mod,
-                          resample=resample,
-                          n=500,
-                          nx=120,
-                          nh=31,
-                          k_len=.15,
[…]

### 2017-03-15 src/core/cgpcm.py
Finished up toy experiments

@@ -30,7 +30,7 @@ class CGPCM(Parametrisable):
         self._init_expressions()
 
     @classmethod
-    def from_recipe(cls, sess, e, nx, nh, k_len, k_wiggles, causal,
+    def from_recipe(cls, sess, e, nx, nh, tau_w, tau_f, causal,
                     causal_id=False, noise_init=1e-4):
         """
         Generate parameters for the CGPCM and construct afterwards.
@@ -39,8 +39,8 @@ class CGPCM(Parametrisable):
         :param e: observations
         :param nx: number of inducing points for noise
         :param nh: number of inducing points for filter
-        :param k_len: length of kernel
-        :param k_wiggles: number of wiggles in kernel
+        :param tau_w: length of kernel window
+        :param tau_f: length scale of function prior
         :param causal: causal model
         :param causal_id: causal interdomain transformation
         :param noise_init: initialisation of noise
@@ -50,31 +50,25 @@ class CGPCM(Parametrisable):
         vars = {}
 
         # Config
-        k_stretch = 2
-        causal_extra_points = 3
+        k_stretch = 3
+        causal_extra_points = 2
 
-        # Compute effective kernel length to match equal prior power
+        # Acausal parameters
+        alpha = length_scale(tau_w)
+        gamma = length_scale(tau_f) - .5 * alpha
+        s2_f, vars['s2_f'] = var_pos(to_float((2 * alpha / np.pi) ** .5))
+
+        # Update in the causal case
         if causal:
-            k_len_eff = 2 * k_len
-        else:
-            k_len_eff = k_len
+            gamma += 3. * alpha / 8.
+            alpha /= 4.
 
-        # Hyperparameters for filter
-        alpha = length_scale(k_len_eff)
-        gamma = length_scale(float(k_len_eff) / k_wiggles)
         alpha, vars['alpha'] = var_pos(to_float(alpha))
         gamma, vars['gamma'] = var_pos(to_float(gamma))
 
-        # Scale to unity prior power
-        if causal:
-            s2_f = (8 * alpha / np.pi) ** .5
-        else:
-            s2_f = (2 * alpha / np.pi) ** .5
-        s2_f, vars['s2_f'] = var_pos(to_float(s2_f))
-
         # Hyperparameter and inducing points for noise
         if nx > 0:
-            omega = length_scale((max(e.x) - min(e.x)) / nx)
+            omega = .5 * length_scale((max(e.x) - min(e.x)) / nx)
[…]

### 2017-03-15 src/core/data.py

@@ -250,7 +250,7 @@ def load_gp_exp(sess, n=250, k_len=.1):
     return f, k
 
 
-def load_akm(sess, causal, n=250, nh=31, k_len=.1, k_wiggles=2, resample=0):
+def load_akm(sess, causal, n=250, nh=31, tau_w=.1, tau_f=.05, resample=0):
     """
     Sample from the AKM.
 
@@ -258,14 +258,14 @@ def load_akm(sess, causal, n=250, nh=31, k_len=.1, k_wiggles=2, resample=0):
     :param causal: causal AKM
     :param n: number of time points
     :param nh: number of inducing points for the filter
-    :param k_len: length of kernel
-    :param k_wiggles: number of wiggles in kernel
+    :param tau_w: length kernel window
+    :param tau_f: kernel length of function prior
     :param resample: number of times to resample function
     :return: data for function, kernel, and filter, and parameters of the AKM
     """
     # Config
-    frac_k_len_pos = .2
-    k_stretch = 8
+    frac_tau_f_pos = .1
+    k_stretch = 8  # Causal case will range from 0 to 6
     e = Data(np.linspace(0, 1, n), None)
 
     # Construct AKM
@@ -273,8 +273,8 @@ def load_akm(sess, causal, n=250, nh=31, k_len=.1, k_wiggles=2, resample=0):
                                 e=e,
                                 nx=0,
                                 nh=nh,
-                                k_len=k_len,
-                                k_wiggles=k_wiggles,
+                                tau_w=tau_w,
+                                tau_f=tau_f,
                                 causal=causal)
 
     # Sample
@@ -284,10 +284,10 @@ def load_akm(sess, causal, n=250, nh=31, k_len=.1, k_wiggles=2, resample=0):
 
     # Construct data
     f = akm.f()
-    tk = np.linspace(-k_stretch * akm.k_len,
-                     k_stretch * akm.k_len, 301)
+    tk = np.linspace(-k_stretch * tau_w,
+                     k_stretch * tau_w, 301)
     k = akm.k(tk)
-    i = util.nearest_index(tk, frac_k_len_pos * k_len)
+    i = util.nearest_index(tk, frac_tau_f_pos * tau_f)
     h = akm.h(tk, assert_positive_at_index=i)
 
     # Normalise

### 2017-03-15 src/core/exp.py

@@ -163,7 +163,8 @@ class TaskOptions(object):
 
     def _fn_opts_to_str(self, xs):
         def to_str(x): return ('y' if x else 'n') if type(x) == bool \
-                                                  else str(x)
+            else str(x)
+
         return ','.join(['{}={}'.format(x['name'], to_str(x['value']))
                          for x in xs])
 
@@ -193,6 +194,8 @@ def train(sess, task, debug_options):
     if 'kernel' in debug_options:
         out.state('plotting kernel and filter and then exiting')
         p = Plotter2D()
+        p.figure('Data')
+        p.plot(task.data['f'].x, task.data['f'].y)
         p.figure('Kernel')
         p.plot(task.data['k'].x, task.data['k'].y)
         p.figure('Filter')
@@ -249,7 +252,7 @@ def train(sess, task, debug_options):
     task.data['f_pred'] = mod.predict_f(task.data['f'].x)
     task.data['k_pred'] = mod.predict_k(task.data['k'].x)
     task.data['psd_pred'] = mod.predict_k(task.data['k'].x, psd=True)
-    pos_i = util.nearest_index(task.data['h'].x, .1 * mod.k_len)
+    pos_i = util.nearest_index(task.data['h'].x, .2 * mod.tau_w)
     task.data['h_pred'] = mod.predict_h(task.data['h'].x,
                                         assert_positive_at_index=pos_i)
     out.section_end()
@@ -364,7 +367,7 @@ def plot_compare(tasks, options):
     task2_colour = '#ca0020'
     marker_size = 2
     inducing_point_size = 2
-    psd_drop = 17.5
+    psd_drop = 100
 
     p = Plotter2D(figure_size=(24, 5) if 'big' in options else (12, 6),
                   font_size=12, figure_toolbar='none', grid_colour='none')
@@ -453,6 +456,7 @@ def plot_compare(tasks, options):
     # Build a filter to limit x axis
     x, y = task1.data['psd'].x, task1.data['psd'].y
     freq_max = max(np.abs(x[y >= max(y) - psd_drop]))
+    freq_max = 0.1
 
     def freq_filter(d):
         return d.filter(lambda x: x[np.logical_and(x >= 0, x <= freq_max)])[0]
@@ -470,8 +474,9 @@ def plot_compare(tasks, options):
            line_width=1)
     p.title('PSD of $f\,|\,h$')
[…]

### 2017-03-15 src/core/util.py

@@ -32,7 +32,7 @@ def length_scale(ls):
     :param ls: length scale
     :return: constant
     """
-    return .5 / ls ** 2
+    return (.5 * np.pi) * (.5 / ls ** 2)
 
 
 def is_inf(x):

### 2017-03-15 src/experiment_compare_series.sh

@@ -1,11 +1,13 @@
 #!/usr/bin/env bash
+
 task=$1
 seed=$2
-to_resample=$3
-opts=${@:4:${#}}
-for (( i=0; i<=$to_resample; i++ ))
-do
-    ./experiment_compare.sh $task seed $seed resample $i $opts
+from_resample=$3
+to_resample=$4
+opts=${@:5:${#}}
+
+for (( i = $from_resample; i <= $to_resample; i++ )); do
+    ./experiment_compare.sh $task $seed $i $opts
 done
 
 

### 2017-03-15 src/experiment_kernels.sh

@@ -1,0 +1,15 @@
+#!/usr/bin/env bash
+set -e
+
+task=$1
+from_seed=$2
+to_seed=$3
+step_seed=$4
+opts=${@:5:${#}}
+
+for (( i = $from_seed; i <= $to_seed; i += $step_seed )); do
+    ./train.py -t $task seed $i resample 0 small $opts -d kernel
+done
+
+
+

### 2017-03-15 src/experiment_toy.sh

@@ -1,0 +1,23 @@
+#!/usr/bin/env bash
+seed_causal=1025
+seed_acausal=1030
+from_resample=0
+to_resample=5
+
+for (( i = $from_resample; i <= $to_resample; i++ )); do
+    # Causal sample
+    prefix="toy seed $seed_causal resample $i causal"
+    ./train.py -t $prefix causal-model
+    ./train.py -t $prefix
+    ./plot.py full -t $prefix causal-model
+    ./plot.py full -t $prefix
+    ./plot.py compare -t $prefix causal-model -t $prefix
+
+    # Acausal sample
+    prefix="toy seed $seed_acausal resample $i noisy"
+    ./train.py -t $prefix causal-model
+    ./train.py -t $prefix
+    ./plot.py full -t $prefix causal-model
+    ./plot.py full -t $prefix
+    ./plot.py compare -t $prefix causal-model -t $prefix
+done

### 2017-03-15 src/experiment_train_plot.sh

@@ -7,6 +7,6 @@ resample=$3
 opts=${@:4:${#}}
 
 prefix="$task seed $seed resample $resample $opts"
-./train.py -t $prefix causal-model -t $prefix
-./plot.py compare -t $prefix causal-model -t $prefix
+./train.py -t $prefix
+./plot.py full -t $prefix -s
 

### 2017-03-15 src/tasks/toy.py

@@ -29,57 +29,46 @@ class Experiment(Task):
 
                           # Training options
                           iters_pre=0 if options['no-pre'] else 250,
-                          iters=25000 if options['extensive'] else 1000,
+                          iters=10000 if options['extensive'] else 2000,
                           iters_post=50 if options['post'] else 0,
-                          samps=200,
+                          samps=400,
 
                           # Sample options
                           causal=options['causal'],
                           causal_model=options['causal-model'],
                           resample=options['resample'],
-                          n=250 if options['small'] else 400,  # 500
-                          nx=100 if options['small'] else 100,  # 120
+                          n=150 if options['small'] else 400,
+                          nx=50 if options['small'] else 120,
                           nh=41,
-                          k_len=1.,
-                          k_wiggles=1.5,
-                          noise=1e-1 if options['noisy'] else 1e-4,
+                          noise=.5 if options['noisy'] else 1e-4,
                           noise_init=1e-4 if options['causal'] else 1e-2,
 
-                          data_scale=.8)
+                          tau_w=0.12 if options['causal'] else 0.04,
+                          tau_f=0.12 if options['causal'] else 0.04,
+                          data_scale=.75)
 
     def load(self, sess):
-        data_k_len = self.config.data_scale * self.config.k_len
-        data_wiggles = self.config.k_wiggles
-        learner_wiggles = self.config.k_wiggles / self.config.data_scale
-
-        if self.config.causal:
-            # Match effective kernel length
-            data_k_len /= 2.
-
-            # Match effective complexity of sample
-            data_wiggles /= 2
-
-
         # Load data
         f, k, h, psd = data.load_akm(sess=sess,
                                      n=self.config.n,
                                      nh=self.config.nh,
-                                     k_len=data_k_len,
-                                     k_wiggles=data_wiggles,
+                                     tau_w=self.config.tau_w \
+                                           * self.config.data_scale,
+                                     tau_f=self.config.tau_f \
+                                           * self.config.data_scale,
                                      causal=self.config.causal,
                                      resample=self.config.resample)
         e = f.make_noisy(self.config.noise)
         self._set_data(h=h, k=k, f=f, e=e, psd=psd)
-
-        # Too many wiggles in causal model case now
 
         # Construct model
         mod = VCGPCM.from_recipe(sess=sess,
                                  e=e,
                                  nx=self.config.nx,
                                  nh=self.config.nh,
-                                 k_len=self.config.k_len,
[…]

### 2017-03-15 src/train_plot.sh

@@ -1,5 +1,0 @@
-#!/usr/bin/env bash
-set -e
-./train.py -t $@
-./plot.py full -t $@ -s
-

### 2017-03-15 doc/paper.tex
First commit of source, still broken
Typos
Added draft of section on choice of hyperparameters

@@
 % Break in equations...
-\allowdisplaybreaks
+% \allowdisplaybreaks
 
@@
 \renewcommand{\max}{\operatorname{max}}
+\newcommand{\atan}{\operatorname{atan}}
 \newcommand{\argmax}{\operatorname{argmax}}
@@
 % Special math commands
-\renewcommand{\ss}[1]{_\mathit{#1}}       % Subscripts without spacing
+% \renewcommand{\ss}[1]{_\mathit{#1}}       % Subscripts without spacing
+\renewcommand{\ss}[1]{_\text{#1}}         % Subscripts in text
 \newcommand{\id}[1]{\, \mathrm{d} #1}     % Straight 'd' in integral
@@
 
+\subsection{Choice of Hyperparameters}
+Following \citet{Pavliotis:2012:Probabilistic_Perspective}, let the \textit{correlation time} $\tau$, or alternatively the \textit{length scale}, of a stationary process with kernel $k(r)$ be
+\begin{align*}
+    \tau = \frac{1}{k(0)}\int_0 k(r) \id{r}.
+\end{align*}
+
+Let $\tau\ss{ac}$ and $P\ss{ac}$ and $\tau\ss{c}$ and $P\ss{c}$ be respectively the correlation time and power of the process $f$ for the GPCM and CGPCM respectively.
+Then Appendix XX shows that $\tau_w=\sqrt{\pi/8 \alpha}$ and
+\begin{align*}
+    P\ss{ac} &= \sigma_{f,\text{ac}}^2\sqrt{\frac{\pi}{2\alpha\ss{ac}}},
+    &\tau\ss{ac} &= \sqrt{\frac{\pi}{2(\alpha\ss{ac} + 2 \gamma\ss{ac})}}  \\
+    P\ss{c} &= \sigma_{f,\text{c}}^2\sqrt{\frac{\pi}{8\alpha\ss{c}}},
+    &\tau\ss{c} &= \sqrt{\frac{2}{\pi(\alpha\ss{c} + 2 \gamma\ss{c})}} \\
+    &&&\phan{=}\atan\ll( \sqrt{\frac{\alpha\ss{c} + 2 \gamma\ss{c}}{\alpha\ss{c}}} \rr). \\
+\end{align*}
+It follows that $P\ss{c}=\frac{1}{2}P\ss{ac}$ and $\tau\ss{c}\approx \tau\ss{ac}$ if $\gamma\ss{c} \gg \alpha\ss{c}$, $\alpha\ss{c}=\alpha\ss{ac}$, $\gamma\ss{c}=\gamma\ss{ac}$, and $\sigma_{f,\text{c}}^2=\sigma_{f,\text{ac}}^2$.
+To allow for fair comparison, we require that $P\ss{ac}=P\ss{c}$ and $\tau\ss{ac}=\tau\ss{c}$.
+If we further choose $P\ss{ac}=P\ss{c}=1$ and $\sigma_{f,\text{ac}}^2=\sigma_{f,\text{c}}^2$, then we find that $\alpha\ss{c}=\frac{1}{4}\alpha\ss{ac}$, $\gamma\ss{c}=\frac{3}{8}\alpha\ss{ac}+ \gamma\ss{ac}$, and $\sigma_{f,\text{ac}}^2 = \sigma_{f,\text{c}}^2 = \sqrt{2\alpha\ss{ac}/\pi}$.
+
+To determine $\alpha\ss{ac}$ and $\gamma\ss{ac}$, we specify the correlation time $\tau_w$ of the window $w$ and the correlation time $\tau_f$ of the process $f$. Then Appendix XX shows that
+\begin{align*}
+    \alpha\ss{ac} &= \frac{\pi}{2} \cdot \frac{1}{2 \tau_w^2},
+    &\gamma\ss{ac} &= \frac{\pi}{2} \cdot \frac{1}{2 \tau^2_f} - \frac{1}{2} \alpha\ss{ac}.
+\end{align*}
+Since $\alpha\ss{ac} > 0$ and $\gamma\ss{ac} > 0$, it must hold that $\sqrt{2} \tau_w > \tau_f > 0$.
+
+Furthermore, for the GPCM, we let $t_u$ be evenly spaced from $-3\tau_w$ to $3 \tau_w$.
+For the CGPCM, we let $t_u$ be evenly spaced from zero to $6 \tau_w$, shifted two inter-point spacings $\Delta t_u$ to the left to account for derivatives of $h$ at zero.
+
+Finally, we let $t_z$ be evenly spaced on the domain of interest, and choose $\omega$ such that $s$'s correlation time $\tau_s$ equals twice $t_z$'s inter-point spacing $\Delta t_z$: $\omega = \pi / 8 \Delta t_z^2$.
+
 \section{Inference}
@@
 {\color{red}Red} denotes truth; {\color{darkgreen}green} denotes observations, the emperical autocorrelation, and the periodogram for respectively the function $f\cond h$, the kernel $k_{f\cond h}$, and the PSD of $f\cond h$; {\color{blue}blue} denotes quantities learned through the MF approximation; and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation.
-For any learned quantity it holds that lines correspond to means and areas correpond to 95\% central credible regions.}
+For any learned quantity it holds that lines correspond to means and areas correspond to 95\% central credible regions.}
     \label{fig:toy_acausal_sample}
@@
 {\color{red}Red} denotes truth; {\color{darkgreen}green} denotes observations, the emperical autocorrelation, and the periodogram for respectively the function $f\cond h$, the kernel $k_{f\cond h}$, and the PSD of $f\cond h$; {\color{blue}blue} denotes quantities learned through the MF approximation; and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation.
-For any learned quantity it holds that lines correspond to means and areas correpond to 95\% central credible regions.}
+For any learned quantity it holds that lines correspond to means and areas correspond to 95\% central credible regions.}
     \label{fig:toy_causal_sample}
@@
 {\color{red}Red} denotes truth, {\color{blue}blue} denotes quantities learned through the MF approximation, and {\color{mycyan}cyan} denotes quantities learned through the SMF approximation.
-For any learned quantity it holds that lines correspond to means and areas correpond to 95\% central credible regions.}
+For any learned quantity it holds that lines correspond to means and areas correspond to 95\% central credible regions.}
     \label{fig:hrir}

### 2017-03-15 src/core/data.py
Added HRIR experiment

@@ -201,17 +201,27 @@ def load_hrir(n=1000, h_wav_fn='data/KEMAR_R10e355a.wav'):
     :param h_wav_fn: HRIR WAV file
     :return: data for function, filter, and noise
     """
+    # Load data
     h = Data.from_wav(h_wav_fn)
     dt = h.x[1] - h.x[0]
-    t = np.arange(n) * dt
-    x = Data(t, np.random.randn(n))
+
+    # Take extra `len(h.x)` points to avoid transition effects
+    t = np.arange(n + len(h.x)) * dt
+    x = Data(t, np.random.randn(n + len(h.x)))
     x /= x.std  # Make sure noise is unity variance
-    f = Data(t, np.convolve(h.y, x.y)[:n])
+
+    # Convolve and take the right fragment
+    f = Data(t, np.convolve(h.y, x.y)).fragment(n, len(h.x))[0]
+
+    k = Data(np.linspace(-h.x[-1], h.x[-1],  2 * len(h.x) - 1),
+             np.convolve(h.y[::-1], h.y))
 
     # Normalise
-    h /= f.std
+    h /= h.energy ** .5
+    k /= max(k.y)
     f /= f.std
-    return f, h
+
+    return f, k, h
 
 
 def load_timit_tobar2015():

### 2017-03-15 src/core/exp.py

@@ -152,14 +152,19 @@ class TaskOptions(object):
     def __getitem__(self, name):
         return self._get_option(name)['value']
 
-    def fn(self, group_by=None):
+    def fn(self, group_by=None, prefix=''):
         group_names = [] if group_by is None else sorted(group_by)
         fn_names = sorted(set([x['name'] for x in self._options])
                           - set(group_names))
-        return '{}/{}'.format(self._fn_opts_to_str(map(self._get_option,
-                                                       group_names)),
-                              self._fn_opts_to_str(map(self._get_option,
-                                                       fn_names)))
+
+        if prefix:
+            prefix += '_'  # Add separator
+
+        group_part = self._fn_opts_to_str(map(self._get_option, group_names))
+        if group_part:
+            prefix += group_part + '/'
+
+        return prefix + self._fn_opts_to_str(map(self._get_option, fn_names))
 
     def _fn_opts_to_str(self, xs):
         def to_str(x): return ('y' if x else 'n') if type(x) == bool \
@@ -257,24 +262,25 @@ def train(sess, task, debug_options):
                                         assert_positive_at_index=pos_i)
     out.section_end()
 
-    # Train SMF
-    out.section('training SMF')
-    task.data['samples'] = mod.sample(iters=task.config.samps)
-    out.section_end()
+    if task.config.samps > 0:
+        # Train SMF
+        out.section('training SMF')
+        task.data['samples'] = mod.sample(iters=task.config.samps)
+        out.section_end()
 
-    # Predict SMF
-    out.section('predicting SMF')
-    task.data['f_pred_smf'] = mod.predict_f(task.data['f'].x,
-                                            samples_h=task.data['samples'])
-    task.data['k_pred_smf'] = mod.predict_k(task.data['k'].x,
-                                            samples_h=task.data['samples'])
-    task.data['psd_pred_smf'] = mod.predict_k(task.data['k'].x,
-                                              samples_h=task.data['samples'],
-                                              psd=True)
-    task.data['h_pred_smf'] = mod.predict_h(task.data['h'].x,
-                                            samples_h=task.data['samples'],
[…]

### 2017-03-15 src/experiment_compare.sh

@@ -1,12 +1,0 @@
-#!/usr/bin/env bash
-set -e
-
-task=$1
-seed=$2
-resample=$3
-opts=${@:4:${#}}
-
-prefix="$task seed $seed resample $resample $opts"
-./train.py -t $prefix causal-model -t $prefix
-./plot.py compare -t $prefix causal-model -t $prefix
-

### 2017-03-15 src/experiment_compare_series.sh

@@ -1,14 +1,0 @@
-#!/usr/bin/env bash
-
-task=$1
-seed=$2
-from_resample=$3
-to_resample=$4
-opts=${@:5:${#}}
-
-for (( i = $from_resample; i <= $to_resample; i++ )); do
-    ./experiment_compare.sh $task $seed $i $opts
-done
-
-
-

### 2017-03-15 src/experiment_kernels.sh

@@ -1,15 +1,0 @@
-#!/usr/bin/env bash
-set -e
-
-task=$1
-from_seed=$2
-to_seed=$3
-step_seed=$4
-opts=${@:5:${#}}
-
-for (( i = $from_seed; i <= $to_seed; i += $step_seed )); do
-    ./train.py -t $task seed $i resample 0 small $opts -d kernel
-done
-
-
-

### 2017-03-15 src/experiment_toy.sh

@@ -1,23 +1,22 @@
 #!/usr/bin/env bash
-seed_causal=1025
-seed_acausal=1030
+
 from_resample=0
 to_resample=5
 
 for (( i = $from_resample; i <= $to_resample; i++ )); do
     # Causal sample
-    prefix="toy seed $seed_causal resample $i causal"
-    ./train.py -t $prefix causal-model
-    ./train.py -t $prefix
-    ./plot.py full -t $prefix causal-model
-    ./plot.py full -t $prefix
-    ./plot.py compare -t $prefix causal-model -t $prefix
+    task="toy resample $i causal-sample"
+    ./train.py -t $task causal-model
+    ./train.py -t $task
+    ./plot.py full -t $task causal-model
+    ./plot.py full -t $task
+    ./plot.py compare -t $task causal-model -t $task
 
     # Acausal sample
-    prefix="toy seed $seed_acausal resample $i noisy"
-    ./train.py -t $prefix causal-model
-    ./train.py -t $prefix
-    ./plot.py full -t $prefix causal-model
-    ./plot.py full -t $prefix
-    ./plot.py compare -t $prefix causal-model -t $prefix
+    task="toy resample $i"
+    ./train.py -t $task causal-model
+    ./train.py -t $task
+    ./plot.py full -t $task causal-model
+    ./plot.py full -t $task
+    ./plot.py compare -t $task causal-model -t $task
 done

### 2017-03-15 src/experiment_train_plot.sh

@@ -1,12 +1,0 @@
-#!/usr/bin/env bash
-set -e
-
-task=$1
-seed=$2
-resample=$3
-opts=${@:4:${#}}
-
-prefix="$task seed $seed resample $resample $opts"
-./train.py -t $prefix
-./plot.py full -t $prefix -s
-

### 2017-03-15 src/tasks/hrir.py

@@ -1,0 +1,50 @@
+from core.exp import Task, TaskConfig, TaskOptions
+from core.cgpcm import VCGPCM
+import core.data as data
+
+
+class Experiment(Task):
+    """
+    HRIR.
+    """
+
+    def generate_config(self, args):
+        options = TaskOptions()
+        options.add_option('causal-model', 'use the causal model')
+        options.parse(args)
+
+        return TaskConfig(name='HRIR',
+                          seed=0,
+                          fn=options.fn(prefix='hrir'),
+
+                          # Training options
+                          iters_pre=50,
+                          iters=500,
+                          iters_post=0,
+                          samps=0,
+
+                          # Model options
+                          causal_model=options['causal-model'],
+                          n=500,
+                          nx=200,
+                          nh=101,
+                          noise_init=1e-4,
+                          tau_w=1e-3,
+                          tau_f=1e-4)
+
+    def load(self, sess):
+        # Load data
+        f, k, h = data.load_hrir(n=self.config.n)
+        e = f
+        self._set_data(h=h, f=f, e=f, k=k)
+
+        # Construct model
+        mod = VCGPCM.from_recipe(sess=sess,
+                                 e=e,
+                                 nx=self.config.nx,
+                                 nh=self.config.nh,
+                                 tau_w=self.config.tau_w,
+                                 tau_f=self.config.tau_f,
+                                 causal=self.config.causal_model,
+                                 noise_init=self.config.noise_init)
+        self._set_model(mod)

### 2017-03-15 src/tasks/toy.py

@@ -10,37 +10,31 @@ class Experiment(Task):
 
     def generate_config(self, args):
         options = TaskOptions()
-        options.add_option('causal', 'generate a causal sample')
+        options.add_option('causal-sample', 'generate a causal sample')
         options.add_option('causal-model', 'use the causal model')
-        options.add_option('post', 'use posttraining')
-        options.add_option('no-pre', 'no pretraining')
-        options.add_option('noisy', 'noisy sample')
-        options.add_option('small', 'small sample')
-        options.add_option('extensive', 'extensive training')
         options.add_value_option('resample', int,
                                  'number of times to resamples',
                                  required=True)
-        options.add_value_option('seed', int, 'seed', required=True)
         options.parse(args)
 
         return TaskConfig(name='Toy Experiment',
-                          seed=options['seed'],
-                          fn=options.fn(group_by=['causal', 'seed', 'small']),
+                          seed=1025 if options['causal'] else 1030,
+                          fn=options.fn(group_by=['causal'], prefix='toy'),
 
                           # Training options
-                          iters_pre=0 if options['no-pre'] else 250,
-                          iters=10000 if options['extensive'] else 2000,
-                          iters_post=50 if options['post'] else 0,
+                          iters_pre=0,
+                          iters=2000,
+                          iters_post=0,
                           samps=400,
 
                           # Sample options
                           causal=options['causal'],
                           causal_model=options['causal-model'],
                           resample=options['resample'],
-                          n=150 if options['small'] else 400,
-                          nx=50 if options['small'] else 120,
+                          n=400,
+                          nx=120,
                           nh=41,
-                          noise=.5 if options['noisy'] else 1e-4,
+                          noise=1e-4 if options['causal'] else .5,
                           noise_init=1e-4 if options['causal'] else 1e-2,
 
                           tau_w=0.12 if options['causal'] else 0.04,

### 2017-03-16 src/core/data.py
Refactored train.py and plot.py into core.py, and fixed up file name handling

@@ -193,7 +193,7 @@ class Data(object):
         return cls(t, y)
 
 
-def load_hrir(n=1000, h_wav_fn='data/KEMAR_R10e355a.wav'):
+def load_hrir(n=1000, h_wav_fn='data/KEMAR_R10e355a.wav', resample=0):
     """
     Load HRIR data.
 
@@ -207,7 +207,8 @@ def load_hrir(n=1000, h_wav_fn='data/KEMAR_R10e355a.wav'):
 
     # Take extra `len(h.x)` points to avoid transition effects
     t = np.arange(n + len(h.x)) * dt
-    x = Data(t, np.random.randn(n + len(h.x)))
+    for i in range(resample + 1):
+        x = Data(t, np.random.randn(n + len(h.x)))
     x /= x.std  # Make sure noise is unity variance
 
     # Convolve and take the right fragment
@@ -311,3 +312,38 @@ def load_akm(sess, causal, n=250, nh=31, tau_w=.1, tau_f=.05, resample=0):
     psd = Data(util.fft_freq(len(psd)), psd)
 
     return f, k, h, psd
+
+
+def load_seaice():
+    import pandas as pd
+
+    import datetime
+
+    def year_fraction(date):
+        start = datetime.date(date.year, 1, 1).toordinal()
+
+        year_length = datetime.date(date.year + 1, 1, 1).toordinal() - start
+
+        decimal_date = date.year + float(
+            date.toordinal() - start) / year_length
+
+        return decimal_date
+
+    ### Read in daily data
+
+    names = ['Year', 'Month', 'Day', 'Extent', 'Missing', 'Source Data']
+
+    df = pd.read_csv('/home/scott/DATA/N_seaice_extent_daily_v2.1.csv',
+                     skiprows=2, names=names)
+
+    dec_date = np.zeros(len(df))
+
+    for j, i in enumerate(df.index):
+        df1 = df[df.index == i]
+
+        dec_date[j] = year_fraction(
+            datetime.date(df1.Year, df1.Month, df1.Day))
+
+    df['decimal_date'] = dec_date
+
+

### 2017-03-16 src/core/exp.py

@@ -3,6 +3,7 @@ import numpy as np
 
 from par import Parametrisable
 from plot import Plotter2D
+from options import Options
 import out
 import learn
 import util
@@ -62,123 +63,11 @@ class Task(object):
         del self.mod
 
 
-class TaskOptions(object):
-    """
-    Options for task.
-    """
-
-    def __init__(self):
-        self._options = []
-
-    def add_option(self, name, desc='no description available'):
-        """
-        Add a boolean option.
-
-        :param name: name of options
-        :param desc: description of option
-        """
-        self._options.append({'has_value': False,
-                              'name': name.lower(),
-                              'value': False,
-                              'description': desc,
-                              'required': False})
-
-    def add_value_option(self, name, value_type,
-                         desc='no description available', required=False):
-        """
-        Add an option with a value.
-
-        :param name: name of option
-        :param value_type: type of option, should be callable
-        :param desc: description of option
-        :param required: option is required
-        """
-        self._options.append({'has_value': True,
-                              'name': name.lower(),
-                              'value_type': value_type,
-                              'value': None,
-                              'description': desc,
-                              'required': required})
-
-    def _get_option(self, name):
-        for option in self._options:
-            if name == option['name']:
-                return option
-        raise RuntimeError('option "{}" not found'.format(name))
-
-    def parse(self, args):
-        """
-        Parse arguments.
-
-        :param args: arguments
-        """
-        self._options = sorted(self._options, key=lambda x: x['name'])
-        self._parse_help(args)
-        self._parse_args(args)
-        self._parse_required()
-
-    def _parse_required(self):
-        missing = []
-        for option in self._options:
-            if option['required'] and option['value'] is None:
-                missing.append(option['name'])
-        if len(missing) == 1:
-            raise RuntimeError('missing option "{}"'.format(missing[0]))
-        elif len(missing) > 1:
-            missing_string = ', '.join(['"{}"'.format(x) for x in missing])
-            raise RuntimeError('missing options {}'.format(missing_string))
-
-    def _parse_args(self, args):
[…]

### 2017-03-16 src/core/sample.py

@@ -106,7 +106,7 @@ class ESS(object):
         :return: samples
         """
         samples = []
-        fetches_config = [{'name': 'log likelihood', 'modifier': '.2e'},
+        fetches_config = [{'name': 'log likelihood', 'modifier': '.0f'},
                           {'name': 'attempts', 'modifier': 'd'},
                           {'name': 'time/attempt',
                            'modifier': '.2f',

### 2017-03-16 src/core/util.py

@@ -3,6 +3,7 @@ import scipy.stats
 import numpy as np
 import tensorflow as tf
 import os
+import pickle
 
 from tfutil import shape
 
@@ -185,3 +186,20 @@ def mkdirs(path):
     dir_path = os.path.dirname(path)
     if not os.path.exists(dir_path):
         os.makedirs(dir_path)
+
+
+def map_pickle(transformation, fns):
+    """
+    Apply a function on the content of a pickled file or list of pickled files.
+    Ignores incorrect paths.
+
+    :param transformation: function
+    :param fns: path or list of paths
+    """
+    fns = [fns] if type(fns) == str else fns
+    for fn in fns:
+        if os.path.isfile(fn):
+            with open(fn) as f:
+                content = pickle.load(f)
+            with open(fn, 'w') as f:
+                pickle.dump(transformation(content), f)

### 2017-03-16 src/experiment_toy.sh

@@ -1,22 +1,23 @@
 #!/usr/bin/env bash
 
+opts=$@
 from_resample=0
-to_resample=5
+to_resample=100
 
 for (( i = $from_resample; i <= $to_resample; i++ )); do
     # Causal sample
-    task="toy resample $i causal-sample"
-    ./train.py -t $task causal-model
-    ./train.py -t $task
-    ./plot.py full -t $task causal-model
-    ./plot.py full -t $task
-    ./plot.py compare -t $task causal-model -t $task
+    ./core.py -t toy resample $i causal-sample causal-model \
+              -t toy resample $i causal-sample \
+              -p full index 0 \
+              -p full index 1 \
+              -p compare index1 0 index2 1 \
+              $opts
 
     # Acausal sample
-    task="toy resample $i"
-    ./train.py -t $task causal-model
-    ./train.py -t $task
-    ./plot.py full -t $task causal-model
-    ./plot.py full -t $task
-    ./plot.py compare -t $task causal-model -t $task
+    ./core.py -t toy resample $i causal-model \
+              -t toy resample $i \
+              -p full index 0 \
+              -p full index 1 \
+              -p compare index1 0 index2 1 \
+              $opts
 done

### 2017-03-16 src/plot.py

@@ -1,42 +1,0 @@
-#!/usr/bin/env python
-import importlib
-import pickle
-import argparse
-
-import core.out as out
-import core.util as util
-import core.exp as exp
-
-parser = argparse.ArgumentParser(description='Run a task.')
-parser.add_argument('plot', type=str, nargs='+',
-                    help='type of plot', default=[])
-parser.add_argument('-t', '--task', type=str, action='append', nargs='*',
-                    help='task to plot with options', default=[])
-parser.add_argument('-d', '--debug', nargs='*', help='debug options',
-                    default=[])
-parser.add_argument('-s', '--show', action='store_true', help='show plot')
-
-args = parser.parse_args()
-out.kv('plot type', args.plot[0])
-out.kv('plot options', ' '.join(args.plot[1:]))
-
-out.section('fetching tasks')
-tasks = []
-for task_args in args.task:
-    out.kv('specification', ' '.join(task_args))
-
-    # Fetch task
-    module = importlib.import_module('tasks.{}'.format(task_args[0]))
-    fn = module.Experiment(options=task_args[1:]).config.fn
-    with open('tasks/cache/{}.pickle'.format(fn)) as f:
-        tasks.append(pickle.load(f))
-out.section_end()
-
-out.section('plotting')
-p, fn = exp.plot(args.plot[0], tasks, options=args.plot[1:])
-fn = 'output/{}/{}.pdf'.format(args.plot[0], fn)
-util.mkdirs(fn)
-p.save(fn)
-if args.show:
-    p.show()
-out.section_end()

### 2017-03-16 src/tasks/hrir.py

@@ -1,4 +1,4 @@
-from core.exp import Task, TaskConfig, TaskOptions
+from core.exp import Task, TaskConfig, Options
 from core.cgpcm import VCGPCM
 import core.data as data
 
@@ -9,32 +9,36 @@ class Experiment(Task):
     """
 
     def generate_config(self, args):
-        options = TaskOptions()
+        options = Options('hrir')
         options.add_option('causal-model', 'use the causal model')
+        options.add_value_option('resample', value_type=int, default=0,
+                                 desc='number of times to resample')
         options.parse(args)
 
         return TaskConfig(name='HRIR',
                           seed=0,
-                          fn=options.fn(prefix='hrir'),
+                          fp=options.fp(),
 
                           # Training options
-                          iters_pre=50,
+                          iters_pre=200,
                           iters=500,
                           iters_post=0,
-                          samps=0,
+                          samps=200,
 
                           # Model options
                           causal_model=options['causal-model'],
-                          n=500,
-                          nx=200,
-                          nh=101,
+                          n=300,
+                          nx=150,
+                          nh=81,
                           noise_init=1e-4,
                           tau_w=1e-3,
-                          tau_f=1e-4)
+                          tau_f=1.5e-4,
+                          resample=options['resample'])
 
     def load(self, sess):
         # Load data
-        f, k, h = data.load_hrir(n=self.config.n)
+        f, k, h = data.load_hrir(n=self.config.n,
+                                 resample=self.config.resample)
         e = f
         self._set_data(h=h, f=f, e=f, k=k)
 

### 2017-03-16 src/tasks/toy.py

@@ -1,4 +1,4 @@
-from core.exp import Task, TaskConfig, TaskOptions
+from core.exp import Task, TaskConfig, Options
 from core.cgpcm import VCGPCM
 import core.data as data
 
@@ -9,7 +9,7 @@ class Experiment(Task):
     """
 
     def generate_config(self, args):
-        options = TaskOptions()
+        options = Options('toy')
         options.add_option('causal-sample', 'generate a causal sample')
         options.add_option('causal-model', 'use the causal model')
         options.add_value_option('resample', int,
@@ -18,27 +18,27 @@ class Experiment(Task):
         options.parse(args)
 
         return TaskConfig(name='Toy Experiment',
-                          seed=1025 if options['causal'] else 1030,
-                          fn=options.fn(group_by=['causal'], prefix='toy'),
+                          seed=1025 if options['causal-sample'] else 1030,
+                          fp=options.fp(groups=[['causal-sample']]),
 
                           # Training options
-                          iters_pre=0,
+                          iters_pre=400,
                           iters=2000,
                           iters_post=0,
                           samps=400,
 
                           # Sample options
-                          causal=options['causal'],
+                          causal=options['causal-sample'],
                           causal_model=options['causal-model'],
                           resample=options['resample'],
                           n=400,
                           nx=120,
                           nh=41,
-                          noise=1e-4 if options['causal'] else .5,
-                          noise_init=1e-4 if options['causal'] else 1e-2,
+                          noise=1e-4 if options['causal-sample'] else .5,
+                          noise_init=1e-4 if options['causal-sample'] else 1e-2,
 
-                          tau_w=0.12 if options['causal'] else 0.04,
-                          tau_f=0.12 if options['causal'] else 0.04,
+                          tau_w=0.12 if options['causal-sample'] else 0.04,
+                          tau_f=0.12 if options['causal-sample'] else 0.04,
                           data_scale=.75)
 
     def load(self, sess):

### 2017-03-16 src/train.py

@@ -1,49 +1,0 @@
-#!/usr/bin/env python
-import importlib
-import pickle
-import argparse
-
-from core.tfutil import *
-import core.util as util
-import core.out as out
-import core.exp as exp
-
-parser = argparse.ArgumentParser(description='Run a task.')
-parser.add_argument('-t', '--task', type=str, action='append', nargs='*',
-                    help='task to run with options', default=[])
-parser.add_argument('-d', '--debug', nargs='*', help='debug options',
-                    default=[])
-
-args = parser.parse_args()
-out.kv('number of tasks', len(args.task))
-
-for i, task_args in enumerate(args.task):
-    out.section('task')
-    out.kv('task number', '{}/{}'.format(i + 1, len(args.task)))
-    out.kv('specification', ' '.join(task_args))
-
-    # Fetch task
-    module = importlib.import_module('tasks.{}'.format(task_args[0]))
-    task = module.Experiment(options=task_args[1:])
-
-    # Initialise
-    out.kv('name', task.config.name)
-    tf.reset_default_graph()
-    with Session() as sess:
-        util.seed(task.config.seed)
-
-        # Train
-        out.section('training')
-        exp.train(sess, task, args.debug)
-        out.section_end()
-
-        # Save resulting task
-        out.section('saving')
-        task.make_pickleable(sess)
-        fn = 'tasks/cache/{}.pickle'.format(task.config.fn)
-        util.mkdirs(fn)
-        with open(fn, 'w') as f:
-            pickle.dump(task, f)
-        out.section_end()
-
-    out.section_end()

### 2017-03-16 src/core.py
Added missing files

@@ -1,0 +1,120 @@
+#!/usr/bin/env python
+import importlib
+import pickle
+import argparse
+
+from core.tfutil import *
+import core.util as util
+import core.out as out
+import core.exp as exp
+
+
+def plot(plot_specification, tasks, show):
+    """
+    Plot a list of tasks.
+
+    :param plot_specification: specification of plot
+    :param tasks: tasks
+    :param show: show plot
+    """
+    plot_type, plot_options = plot_specification[0], plot_specification[1:]
+    out.kv('plot type', plot_type)
+
+    out.section('fetching tasks')
+    fetched_tasks = []
+    for task_args in tasks:
+        out.kv('specification', ' '.join(task_args))
+
+        # Fetch task
+        module = importlib.import_module('tasks.{}'.format(task_args[0]))
+        fp = module.Experiment(options=task_args[1:]).config.fp
+        path = 'tasks/cache/' + str(fp) + '.pickle'
+        with open(path) as f:
+            fetched_tasks.append(pickle.load(f))
+    out.section_end()
+
+    out.section('plotting')
+    p, fp = exp.plot(plot_type, fetched_tasks, args=plot_options)
+    path = 'output/' + str(fp) + '.pdf'
+    util.mkdirs(path)
+    p.save(path)
+    if show:
+        p.show()
+    out.section_end()
+
+
+def learn(task_specification):
+    """
+    Learn a task.
+
+    :param task_specification: task specification
+    """
+    task_name, task_options = task_specification[0], task_specification[1:]
+
+    # Fetch task
+    module = importlib.import_module('tasks.{}'.format(task_name))
+    task = module.Experiment(options=task_options)
+
+    # Initialise
+    out.kv('name', task.config.name)
+    tf.reset_default_graph()
+    with Session() as sess:
+        util.seed(task.config.seed)
+
+        # Train
+        out.section('training')
+        exp.train(sess, task, args.debug)
+        out.section_end()
+
+        # Save resulting task
+        out.section('saving')
+        task.make_pickleable(sess)
+        path = 'tasks/cache/' + str(task.config.fp) + '.pickle'
+        util.mkdirs(path)
+        with open(path, 'w') as f:
+            pickle.dump(task, f)
+        out.section_end()
+
+
+if __name__ == '__main__':
+    parser = argparse.ArgumentParser(description='Run a task.')
+    parser.add_argument('-l', '--learn', type=int, action='store', nargs='*',
+                        help='learn all tasks, or specify indices of tasks to '
+                             'learn')
[…]

### 2017-03-16 src/core/options.py

@@ -1,0 +1,172 @@
+import os
+
+import out
+
+
+class Options(object):
+    """
+    Options parser.
+
+    :param name: name
+    """
+
+    def __init__(self, name):
+        self._name = name
+        self._options = []
+
+    def add_option(self, name, desc='no description available'):
+        """
+        Add a boolean option.
+
+        :param name: name of options
+        :param desc: description of option
+        """
+        self._options.append({'has_value': False,
+                              'name': name.lower(),
+                              'description': desc,
+                              'required': False,
+
+                              'value': False})
+
+    def add_value_option(self, name, value_type,
+                         desc='no description available', required=False,
+                         default=None):
+        """
+        Add an option with a value.
+
+        :param name: name of option
+        :param value_type: type of option, should be callable
+        :param desc: description of option
+        :param required: option is required
+        :param default: default value
+        """
+        self._options.append({'has_value': True,
+                              'name': name.lower(),
+                              'description': desc,
+                              'required': required,
+
+                              'value': default,
+                              'value_type': value_type,
+                              'specified': False})
+
+    def _get_option(self, name):
+        for option in self._options:
+            if name == option['name']:
+                return option
+        raise RuntimeError('option "{}" not found'.format(name))
+
+    def parse(self, args):
+        """
+        Parse arguments.
+
+        :param args: arguments
+        """
+        self._options = sorted(self._options, key=lambda x: x['name'])
+        self._parse_help(args)
+        self._parse_args(args)
+        self._parse_required()
+
+    def _parse_required(self):
+        missing = []
+        for option in self._options:
+            if option['required'] and not option['specified']:
+                missing.append(option['name'])
+        if len(missing) == 1:
+            raise RuntimeError('missing option "{}"'.format(missing[0]))
+        elif len(missing) > 1:
+            missing_string = ', '.join(['"{}"'.format(x) for x in missing])
+            raise RuntimeError('missing options {}'.format(missing_string))
+
+    def _parse_args(self, args):
+        it = iter(args)
+        for arg in it:
+            option = self._get_option(arg)
+            if option['has_value']:
+                option['value'] = option['value_type'](next(it))
+                option['specified'] = True
+            else:
+                option['value'] = True
+
+    def _parse_help(self, args):
[…]

### 2017-03-24 src/experiment_hrir_plot.sh
Updated HRIR experiment

@@ -1,0 +1,13 @@
+#!/usr/bin/env bash
+
+from_resample=100
+to_resample=120
+
+for (( i = $from_resample; i <= $to_resample; i++ )); do
+    ./core.py -t hrir resample $i causal-model \
+              -t hrir resample $i \
+              -p compare2 index1 0 index2 0 2side-h no-psd ms mf1 \
+              -p compare2 index1 1 index2 1 2side-h no-psd ms mf1 \
+              -p compare2 index1 0 index2 1 2side-h no-psd ms \
+              -r
+done

### 2017-03-24 src/experiment_toy_plot.sh

@@ -1,0 +1,23 @@
+#!/usr/bin/env bash
+
+from_resample=0
+to_resample=40
+
+for (( i = $from_resample; i <= $to_resample; i++ )); do
+    # Causal sample
+    ./core.py -t toy resample $i causal-sample causal-model \
+              -t toy resample $i causal-sample \
+              -p compare2 index1 0 index2 0 mf1 correct-h soft-correct-h \
+              -p compare2 index1 1 index2 1 mf1 correct-h soft-correct-h \
+              -p compare2 index1 0 index2 1 correct-h soft-correct-h \
+              -r
+
+
+    # Acausal sample
+    ./core.py -t toy resample $i causal-model \
+              -t toy resample $i \
+              -p compare2 index1 0 index2 0 mf1 correct-h soft-correct-h \
+              -p compare2 index1 1 index2 1 mf1 correct-h soft-correct-h \
+              -p compare2 index1 0 index2 1 correct-h soft-correct-h \
+              -r
+done

### 2017-03-24 src/tasks/hrir.py

@@ -24,18 +24,18 @@ class Experiment(Task):
 
                           # Training options
                           iters_pre=200,
-                          iters=2000,
+                          iters=200,
                           iters_post=0,
-                          samps=500,
+                          samps=200,
 
                           # Model options
                           causal_model=options['causal-model'],
-                          n=600,
-                          nx=300,
-                          nh=121,
-                          noise_init=2e-3,
+                          n=200,
+                          nx=100,
+                          nh=101,
+                          noise_init=1e-3,
                           tau_w=1e-3,
-                          tau_f=.2e-3,
+                          tau_f=.1e-3,
                           resample=options['resample'])
 
     def load(self, sess):
@@ -55,5 +55,3 @@ class Experiment(Task):
                                  causal=self.config.causal_model,
                                  noise_init=self.config.noise_init)
         self._set_model(mod)
-
-

### 2017-03-24 src/core.py
Changed filter estimating to minimal-phase form

@@ -114,7 +114,7 @@ if __name__ == '__main__':
     parser.add_argument('--stats', action='store_true',
                         help='show statistics of tasks')
     parser.add_argument('-r', '--remote', action='store_true',
-                        help='use remote cache')
+                        help='use remote task cache')
 
     args = parser.parse_args()
 
@@ -171,5 +171,5 @@ if __name__ == '__main__':
         out.section_end()
 
     # Showing plots
-    if len(args.plot) > 0 and args.show:
+    if args.plot and args.show:
         plt.show()

### 2017-03-24 src/core/cgpcm.py

@@ -50,11 +50,11 @@ class CGPCM(Parametrisable):
         vars = {}
 
         # Config
-        k_stretch = 3
+        tau_ws = 2
         causal_extra_points = 2
 
         # Acausal parameters
-        alpha = length_scale(tau_w)
+        alpha = 2 * length_scale(tau_w)
         gamma = length_scale(tau_f) - .5 * alpha
         s2_f, vars['s2_f'] = var_pos(to_float((2 * alpha / np.pi) ** .5))
 
@@ -82,14 +82,14 @@ class CGPCM(Parametrisable):
 
         # Inducing points for filter
         if causal:
-            th = np.linspace(0, 2 * k_stretch * tau_w, nh)
+            th = np.linspace(0, 2 * tau_ws * tau_w, nh)
             dth = th[1] - th[0]
             # Extra inducing points to account for derivatives at zero
             th -= dth * causal_extra_points
             th = constant(th)
         else:
-            th = constant(np.linspace(-k_stretch * tau_w,
-                                      k_stretch * tau_w, nh))
+            th = constant(np.linspace(-tau_ws * tau_w,
+                                      tau_ws * tau_w, nh))
 
         # Initial observation noise
         s2, vars['s2'] = var_pos(to_float(noise_init))
@@ -562,6 +562,7 @@ class VCGPCM(CGPCM):
         if is_numeric(samples_h):
             h = self.h.sample()
             samples_h = [self._run(h) for i in range(samples_h)]
+        n = len(samples_h)
 
         h = placeholder(shape(self.h.sample()))
         Kfu = self.kernel_h(t, self.th)
@@ -569,28 +570,37 @@ class VCGPCM(CGPCM):
         samples = map_progress(lambda x: self._run(mu, feed_dict={h: x}),
                                samples_h,
                                name='filter prediction using MC')
-        samples = np.concatenate(samples, axis=1)
 
+        # Check whether to correct the signs of the samples
         if correct_signs:
-            # Set samples to correct sign
+            samples = np.concatenate(samples, axis=1)  # Pack
             samples *= util.sign_smart(samples.T)[None, :]
+            samples = np.split(samples, n, 1)  # Unpack
+
+        def process(sample):
+            y = data.Data(t, sample)
+            if self.causal:
+                y = y.positive_part()
+            y = y.minimum_phase()
+
+            # Check whether to normalise predictions
+            if normalise:
+                y /= y.energy ** .5
+
+            return y.y[:, None]
+
[…]

### 2017-03-24 src/core/data.py

@@ -1,5 +1,6 @@
 import scipy.io as sio
 import scipy.io.wavfile as sio_wav
+import scipy.signal as signal
 
 from tf_util import *
 import util
@@ -16,14 +17,14 @@ class Data(object):
     """
 
     def __init__(self, x, y=None):
-        self.x = x
-        self.n = shape(x)[0]
+        self.x = np.squeeze(x)
+        self.n = len(self.x)
         if y is None:
             # Data is missing
             self.y = np.zeros(shape(x))
             self.y.fill(np.nan)
         else:
-            self.y = y
+            self.y = np.squeeze(y)
 
     def subsample(self, n):
         """
@@ -139,6 +140,28 @@ class Data(object):
         :return: new data
         """
         return Data(other_x, np.interp(other_x, self.x, self.y))
+
+    def zero_phase(self):
+        """
+        Transform signal to zero-phase form.
+
+        :return: zero-phase form
+        """
+        y = np.fft.ifft(np.abs(np.fft.fft(self.y)))
+        y = np.real(np.fft.fftshift(y))
+        x = self.dx * np.arange(self.n)
+        x -= x[self.n / 2]
+        return Data(x, y)
+
+    def minimum_phase(self):
+        """
+        Transform signal to minimum-phase form.
+
+        :return: minimum-phase form
+        """
+        log_mag = np.log(np.abs(np.fft.fft(self.y)))
+        y = np.real(np.fft.ifft(np.exp(signal.hilbert(log_mag).conj())))
+        return Data(self.dx * np.arange(len(y)), y)
 
     @property
     def dx(self):
@@ -290,7 +313,6 @@ def load_hrir(n=1000, h_wav_fn='data/KEMAR_R10e355a.wav', resample=0):
     """
     # Load data
     h = Data.from_wav(h_wav_fn)
-    h.x -= 1.05e-3  # Shift according to results of `hrir_shift.py`
 
     # Take `h.len` extra points to avoid transition effects
     t = np.arange(n + h.len) * h.dx

### 2017-03-24 src/core/experiment.py

@@ -115,9 +115,9 @@ class Task(object):
             return x[np.logical_and(x.x >= 0, x.x <= 2 * self.config.tau_w)]
 
         # Correct filters
-        correct_filter(self, 'h_pred', 'h')
+        correct_filter(self, 'h_pred', 'h_pred', 'h')
         if 'h_pred_smf' in self.data:
-            correct_filter(self, 'h_pred_smf', 'h')
+            correct_filter(self, 'h_pred_smf', 'h_pred_smf', 'h')
 
         report.update({'prediction': {'function': report_key('f'),
                                       'kernel': report_key('k',
@@ -245,13 +245,15 @@ def train(sess, task, debug_options):
         task.data['elbo_smf'] = mod.elbo_smf(samples)
 
 
-def correct_filter(task, key, reference_key):
+def correct_filter(task, key, out_key, reference_key, soft=False):
     """
     Correct the filter in a task.
 
     :param task: task
     :param key: key of predicted filter
+    :param out_key: key of corrected predicted filter
     :param reference_key: key of true filter
+    :param soft: soft correction: only adjust sign
     """
     mean, lower, upper, std = task.data[key]
     x_orig = mean.x
@@ -263,14 +265,17 @@ def correct_filter(task, key, reference_key):
     lower, upper = (lower, upper) if sign > 0 else (-upper, -lower)
 
     # Shift
-    delta = util.optimal_shift(mean, ref, bounds=(-10 * ref.dx,
-                                                  10 * ref.dx))
+    if not soft:
+        delta = util.optimal_shift(mean, ref, bounds=(-10 * ref.dx,
+                                                      10 * ref.dx))
+    else:
+        delta = 0
 
     # Save
-    task.data[key] = mean.shift(delta).at(x_orig), \
-                     lower.shift(delta).at(x_orig), \
-                     upper.shift(delta).at(x_orig), \
-                     std.shift(delta).at(x_orig)
+    task.data[out_key] = mean.shift(delta).at(x_orig), \
+                         lower.shift(delta).at(x_orig), \
+                         upper.shift(delta).at(x_orig), \
+                         std.shift(delta).at(x_orig)
 
 
 class TaskPlotter(object):
[…]

### 2017-03-24 src/core/plot.py

@@ -9,6 +9,7 @@ with warnings.catch_warnings():
     import matplotlib.pyplot as plt
 
 import util
+import data
 
 
 def to_rbga(color, alpha=1.0):
@@ -264,6 +265,11 @@ class Plotter2D(Plotter):
                    'markeredgecolor': 'marker_colour',
                    'markersize': 'marker_size',
                    'label': 'label'}
+
+        # Unpack if instance of `core.Data` is provided
+        if y is None and type(x) == data.Data:
+            x, y = x.x, x.y
+
         if y is None:
             p = plt.plot(self.transform_y(x),
                          *args, **self._map(mapping, kw_args))

### 2017-03-24 src/hrir_shift.py
Experimentation

@@ -1,40 +1,0 @@
-from core.tf_util import *
-import core.util as util
-import core.data as data
-import core.kernel as kernel
-import core.distribution as dist
-import core.learn as learn
-import core.out as out
-
-
-sess = Session()
-delta_t = tf.Variable(to_float(0))
-f, k, h = data.load_hrir()
-
-# Construct kernel
-tau_w = 1e-3
-tau_f = .2e-3
-alpha = util.length_scale(tau_w)
-gamma = util.length_scale(tau_f) - .5 * alpha
-alpha, alpha_var = var_pos(to_float(alpha))
-gamma, gamma_var = var_pos(to_float(gamma))
-deq = kernel.DEQ(s2=1., alpha=alpha, gamma=gamma)
-
-# Construct prior
-prior = dist.Normal(reg(deq(h.x + delta_t)))
-
-# Optimise log-likelihood
-ll = prior.log_pdf(h.y[None, :])
-initialise_uninitialised_variables(sess)
-learn.minimise_lbfgs(sess, -ll, iters=100,
-                     vars=[delta_t],
-                     fetches_config=[{'name': 'log-likelihood',
-                                      'tensor': ll,
-                                      'modifier': '.2e'}])
-
-# Report
-out.section('results')
-out.kv('alpha', sess.run(alpha), mod='.2e')
-out.kv('gamma', sess.run(gamma), mod='.2e')
-out.kv('shift', sess.run(1000 * delta_t), unit='ms', mod='.4f')
-out.section_end()

### 2017-03-24 src/phase_transform.py

@@ -1,0 +1,25 @@
+import numpy as np
+
+import core.plot as plot
+import core.data as data
+
+
+f, k, h = data.load_hrir(1000)
+
+print 'GO'
+h = h.minimum_phase()
+print 'Finished'
+
+p = plot.Plotter2D()
+p.subplot(2, 2, 1)
+p.title('Filter')
+p.plot(h)
+p.subplot(2, 2, 2)
+p.title('log $|F|$')
+p.plot(h.x, np.log(np.abs(np.fft.fft(h.y))))
+p.subplot(2, 2, 3)
+p.title('arg $F$')
+p.plot(h.x, np.angle(np.fft.fft(h.y)))
+
+
+p.show()

### 2017-03-24 src/report.py

@@ -121,10 +121,10 @@ def report_prediction(key):
                                  ['prediction', key, 'MLL', 'SMF'])}
 
 
-data_sets = {'hrir100': {'seeds': [102, 103, 104, 105, 106, 107, 108, 109, 112,
-                                   113, 114],
+data_sets = {'hrir100': {'seeds': [100, 101, 102, 103, 105, 107, 108, 109, 110,
+                                   111, 112, 113, 114, 116, 117, 120],
                          'base_path': 'output/stats/hrir/causal-model={},'
-                                      'post=n,resample={}.pickle'},
+                                      'resample={}.pickle'},
              'hrir200': {'seeds': [200, 201, 202, 203, 204, 206],
                          'base_path': 'output/stats/hrir/causal-model={},'
                                       'resample={}.pickle'}}

### 2017-03-26 src/core/exponentiated_quadratic.py
Highly performant BVN CDF implementation

@@ -546,8 +546,9 @@ class EQ(object):
         # Evaluate CDF for all `x1` and `x2`
         orig_shape = shape(mu1)
         num = reduce(operator.mul, orig_shape, 1)
-        xs = tf.reshape(tf.stack([x1, x2], axis=-1), [num, 2])
-        cdf_part = tf.reshape(bvn_cdf2(xs, rho), orig_shape)
+        x1 = tf.reshape(x1, [num])
+        x2 = tf.reshape(x2, [num])
+        cdf_part = tf.reshape(bvn_cdf(x1, x2, rho), orig_shape)
 
         # Compute exponentiated part
         quad_form = .5 * (ia11 * b1 ** 2 + ia22 * b2 ** 2 + 2 * ia12 * b1 * b2)

### 2017-03-26 src/core/tf_util_test.py

@@ -3,7 +3,7 @@ import unittest
 from tf_util import *
 
 
-class TestBVNCDF2(unittest.TestCase):
+class TestBvnCdf(unittest.TestCase):
     def setUp(self):
         self.sess = tf.Session()
         self.delta = 1e-8
@@ -14,28 +14,31 @@ class TestBVNCDF2(unittest.TestCase):
         self.div = to_float(5)
         self.rho = sum(self.v * self.random) / self.div
         self.x = constant(np.array([[0.1, 0.2], [0.2, 0.3]]))
-        self.y = bvn_cdf2(self.x, self.rho)
+        self.y = self.bvn_cdf(self.x, self.rho)
+
+    def bvn_cdf(self, x, rho):
+        return bvn_cdf(x[:, 0], x[:, 1], rho)
 
     def test_gradient_dy0_dx00(self):
         e = constant(np.array([[1., 0], [0, 0]])) * self.delta / 2.
-        y1 = bvn_cdf2(self.x - e, self.rho)
-        y2 = bvn_cdf2(self.x + e, self.rho)
+        y1 = self.bvn_cdf(self.x - e, self.rho)
+        y2 = self.bvn_cdf(self.x + e, self.rho)
         out = self.sess.run(y2[0] - y1[0]) / self.delta
         ref = self.sess.run(tf.gradients([self.y[0]], self.x)[0][0, 0])
         np.testing.assert_almost_equal(out, ref)
 
     def test_gradient_dy1_dx10(self):
         e = constant(np.array([[0, 0], [1., 0]])) * self.delta / 2.
-        y1 = bvn_cdf2(self.x - e, self.rho)
-        y2 = bvn_cdf2(self.x + e, self.rho)
+        y1 = self.bvn_cdf(self.x - e, self.rho)
+        y2 = self.bvn_cdf(self.x + e, self.rho)
         out = self.sess.run(y2[1] - y1[1]) / self.delta
         ref = self.sess.run(tf.gradients([self.y[1]], self.x)[0][1, 0])
         np.testing.assert_almost_equal(out, ref)
 
     def test_gradient_dy0_drho(self):
         e = to_float(1) * self.delta / 2
-        y1 = bvn_cdf2(self.x, self.rho - e)
-        y2 = bvn_cdf2(self.x, self.rho + e)
+        y1 = self.bvn_cdf(self.x, self.rho - e)
+        y2 = self.bvn_cdf(self.x, self.rho + e)
[…]
